### 2018-03-01 inst/doc/MSEtool.Rmd
gravity movement TMB almost working!

---
title: "MSEtool"
subtitle: "Development, Testing and Implementation of Management Procedures for Data-Rich Fisheries (v1.1)"
author: "Tom Carruthers (<t.carruthers@oceans.ubc.ca>) 
         Adrian Hordyk  (<a.hordyk@oceans.ubc.ca>),
         Quang Huynh (<q.huynh@oceans.ubc.ca>)"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
  rmarkdown::pdf_vignette:
    toc: true
    number_sections: true    
vignette: >
  %\VignetteIndexEntry{MSEtool}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<style type="text/css">

body{ /* Normal  */
   font-size: 12px;
}
td {  /* Table  */
   font-size: 8px;
}
h1 { /* Header 1 */
 font-size: 18px;
 color: DarkBlue;
}
h2 { /* Header 2 */
 font-size: 15px;
 color: DarkBlue;
}
h3 { /* Header 3 */
 font-size: 14px;
 color: DarkBlue;
}
code.r{ /* Code block */
  font-size: 10px;
}
pre { /* Code block */
  font-size: 10px
}
</style>

```{r set options, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
knitr::opts_chunk$set(dpi=85)
options(width = 650)
```

<br>

# Foreword

MSEtool development is funded by the Canadian Department of Fisheries and Oceans and benefits from ongoing collaborations with the Natural Resources Defense Council and number of Canadian government scientists including Robyn Forrest, Sean Anderson and Daniel Duplisea.

The package uses the Management Strategy Evaluation engine of [DLMtool](http://www.datalimitedtoolkit.org) to undertake closed-loop simulation testing of management procedures. The DLMtool package was developed at the University of British Columbia in collaboration with the Natural Resources Defense Council.

DLMtool development has been funded by the Gordon and Betty Moore Foundation, the Packard Foundation, U.S. National Oceanic and Atmospheric Administration, Fisheries and Oceans Canada, the Walton Foundation, Resources Legacy Fund, the Natural Resources Defense Council, the United Nations Food and Agricultural Organization and the California Department of Fish and Wildlife.

<br>

# Version Notes

The package is subject to ongoing testing. If you find a bug or a problem please send a report to <t.carruthers@oceans.ubc.ca> so that it can be fixed!

## New Additions to this Version (v1.1)

* The MCMC outputs of iSCAM models are now used to characterise operating model uncertainty

* iSCAM2DLM function added to rip from iSCAM assessments

* SS2DLM function updated for latest version of R4SS

## Coming soon

Data-rich MPs

* Statistical Catch at Age (SCA)
* Age-Structure Production Models (ASPM)
* State-Space Delay Difference models (DD_SS)
* Surplus Production models (SP)
* State-Space Surplus Production Models (SP_SS)
* Kalman filer assessment approximation

Visualization

* Standardized assessment reports

# Introduction

## Management Strategy Evaluation and DLMtool

<img src="DLMtool_logo.png" width="90">

Some text

## The operating model

## MSE design

## The MSE process

<img src="DLMtool_process.jpg" width="400">

<br>

#References

```{r, echo=FALSE, eval=FALSE}
sfStop()                            
```

<br>

### 2018-03-01 R/Movement_calcs.R

@@ -17,18 +17,10 @@
 #' @useDynLib MSEtool
 makemov<-function(fracs=rep(0.2,5),muprob=0.5,efracs=0.1,evisc=rep(0.1,5),OM=NULL){
 
-
- # library(DLMtool)
-#  fracs=rep(0.2,5)
-#  muprob=0.9
-#  efracs=0.1
-#  evisc=rep(0.1,5)
-#  OM=testOM
-
   nareas<-length(fracs)
   data <- list(model = "grav",fracs = fracs, muprob = muprob, nareas = nareas)
 
-  params <- list(logit_frac = rep(0,nareas-1), log_visc = 0)
+  params <- list(log_visc = 0,log_grav = rep(0,nareas-1))
   info <- list(data = data, params = params)
 
   obj <- MakeADFun(data = info$data, parameters = info$params, DLL = "MSEtool", silent = TRUE)

### 2018-03-01 inst/doc/MSE.R

@@ -1,8 +1,0 @@
-## ----set options, echo = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
-knitr::opts_chunk$set(dpi=85)
-options(width = 650)
-
-## ---- echo=FALSE, eval=FALSE--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-#  sfStop()
-

### 2018-03-01 inst/doc/MSEtool.R

@@ -1,0 +1,8 @@
+## ----set options, echo = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
+knitr::opts_chunk$set(dpi=85)
+options(width = 650)
+
+## ---- echo=FALSE, eval=FALSE--------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+#  sfStop()
+

### 2018-03-01 inst/include/grav.h

@@ -8,83 +8,98 @@
   DATA_SCALAR( muprob );
   DATA_INTEGER( nareas );
 
-  PARAMETER_VECTOR( log_grav );
   PARAMETER( log_visc );
+  PARAMETER_VECTOR( log_grav );
 
-  vector<Type> log_grav_p = log_grav; // this is a vector, not sure what will happen here
-  Type log_visc_p = log_visc;
-  Type gsum;
-  Type psum;
-  Type sigma;
-  Type obj;
+  Type psum = Type(0);
+  Type sigma = Type(0);
+  sigma=0.2;
 
   // -- Declarations
   matrix<Type> grav(nareas,nareas);
   matrix<Type> mov(nareas,nareas);
   matrix<Type> transN(nareas,nareas);
   vector<Type> idist(nareas);
+  vector<Type> tdist(nareas);
+  vector<Type> gsum(nareas);
+  vector<Type> Nsum(nareas);
+
+  // Zero inits
+  grav.setZero();
+  mov.setZero();
+  idist.setZero();
+  tdist.setZero();
+  gsum.setZero();
+  Nsum.setZero();
+
+  //transN.setZero();
 
   // Map out gravity terms accounting for viscosity, gravity of area 1 is fixed to zero
-  for(int af=1; af<nareas; af++){ // area from
-    grav(af,1) = 0;
-    for(int at=2; af<nareas; af++){ // area to
-      grav(af,at) = log_grav_p(at-1);
+  for(int af=0; af<nareas; af++){ // area from
+    //grav(af,0) = 0.0;
+    for(int at=1; at<nareas; at++){ // area to
+      grav(af,at) = log_grav(at-1);
     }
-    grav(af,af)+=grav(af,af)+log_visc_p; // add viscosity
+    grav(af,af)+=grav(af,af)+log_visc; // add viscosity
   }
 
   // Calculate logit fractions (movement to area from area)
-  for(int af=1; af<nareas; af++){
-    gsum=0.0;
-    for(int at=1; af<nareas; af++){
-      gsum+=exp(grav(af,at)); // sum up gravity terms
+  for(int af=0; af<nareas; af++){
+
+    for(int at=0; at<nareas; at++){
+      gsum(af)+=exp(grav(af,at)); // sum up gravity terms by from area
     }
-    for(int at=1; af<nareas; af++){
[…]

### 2018-05-23 inst/include/DD_SS.h
Update MSEtool userguide

@@ -15,12 +15,12 @@
   DATA_STRING(SR_type);
   DATA_VECTOR_INDICATOR(keep, C_hist);
 
-  PARAMETER( logit_UMSY );
-  PARAMETER( log_MSY );
-  PARAMETER( log_q );
-  PARAMETER( log_sigma );
-  PARAMETER( log_tau );
-  PARAMETER_VECTOR( log_rec_dev );
+  PARAMETER(logit_UMSY);
+  PARAMETER(log_MSY);
+  PARAMETER(log_q);
+  PARAMETER(log_sigma);
+  PARAMETER(log_tau);
+  PARAMETER_VECTOR(log_rec_dev);
 
   Type UMSY = 1/(1 + exp(-logit_UMSY));
   Type MSY = exp(log_MSY);
@@ -132,39 +132,39 @@
   Type B_BMSY_final = B(B.size()-1)/BMSY;
   Type B_B0_final = B(B.size()-1)/B0;
 
-  ADREPORT( UMSY );
-  ADREPORT( MSY );
-  ADREPORT( q );
-  ADREPORT( sigma );
-  ADREPORT( tau );
-  ADREPORT( U_UMSY_final );
-  ADREPORT( B_BMSY_final );
-  ADREPORT( B_B0_final );
-  REPORT( UMSY );
-  REPORT( MSY );
-  REPORT( q );
-  REPORT( sigma );
-  REPORT( tau );
-  REPORT( jnll_comp );
-  REPORT( jnll );
-  REPORT( Arec );
-  REPORT( Brec );
-  REPORT( Spr );
-  REPORT( Spr0 );
-  REPORT( DsprDu );
-  REPORT( Cpred );
-  REPORT( B );
-  REPORT( N );
-  REPORT( R );
-  REPORT( log_rec_dev );
-  REPORT( Rec_dev );
-  REPORT( U );
-  REPORT( h );
-  REPORT( BMSY );
-  REPORT( R0 );
-  REPORT( N0 );
-  REPORT( B0 );
-  REPORT( penalty );
+  ADREPORT(UMSY);
+  ADREPORT(MSY);
+  ADREPORT(q);
+  ADREPORT(sigma);
+  ADREPORT(tau);
+  ADREPORT(U_UMSY_final);
+  ADREPORT(B_BMSY_final);
+  ADREPORT(B_B0_final);
+  REPORT(UMSY);
+  REPORT(MSY);
[…]

### 2018-06-05 inst/doc/MSEtool.Rmd

@@
 ---
-title: "MSEtool"
+title: "User guide to MSEtool"
 subtitle: "Development, Testing and Implementation of Management Procedures for Data-Rich Fisheries (v1.1)"
@@
 vignette: >
-  %\VignetteIndexEntry{MSEtool}
+  %\VignetteIndexEntry{User guide to MSEtool}
   %\VignetteEngine{knitr::rmarkdown}
@@
 
-<br>
+<a href="http://www.datalimitedtoolkit.org"><img src="DLMtool_logo.png" width=250></a>
 
@@
 
+* Assessment models (`Assess` functions)
+    * Statistical Catch at Age (SCA and SCA2)
+    * Delay Difference (DD_TMB)
+    * State-Space Delay Difference (DD_SS)
+    * Surplus Production (SP)
+    * State-Space Surplus Production (SP_SS)
+
+* Visualization
+    * Standardized assessment reports
+    * Diagnostics for assessment models
+
+* Other features 
+    * Harvest control rules (`HCR` functions)
+    * `make_MP` function conveniently combines a __Assess__ and __HCR__ function to create a __MP__ function
+    * Movement matrices for parameterizing OM
     * The MCMC outputs of iSCAM models are now used to characterise operating model uncertainty
-
-* iSCAM2DLM function added to rip from iSCAM assessments
-
-* SS2DLM function updated for latest version of R4SS
+    * iSCAM2OM and iSCAM2Data functions rip from iSCAM assessments
+    * SS2OM and SS2Data functions updated for latest version of r4ss
 
@@
 
-Data-rich MPs
+* More data-rich assessment models
+    * Virtual Population Analysis (VPA)
+    * Age-Structure Production Models (ASPM)
+    * Kalman filter assessment approximation
 
-* Statistical Catch at Age (SCA)
-* Age-Structure Production Models (ASPM)
-* State-Space Delay Difference models (DD_SS)
-* Surplus Production models (SP)
-* State-Space Surplus Production Models (SP_SS)
-* Kalman filer assessment approximation
-
-Visualization
-
-* Standardized assessment reports
+* Stock projections for assessment models
 
@@
 
-## Management Strategy Evaluation and DLMtool
+The MSEtool package is an extension of the DLMtool R package for implementing management strategy evaluation (MSE) simulations.
+Whereas DLMtool is primarily designed for testing data-limited management procedures, MSEtool provides an object-oriented framework for testing data-rich management procedures (MPs), which typically involve combining an assessment model with a harvest control rule to provide a fishery management recommendation.
+MSEtool management procedures are designed to be used with the operating model of DLMtool through the `runMSE` function.
+The user guide to the DLMtool package is accessible via
 
-<img src="DLMtool_logo.png" width="90">
-
-Some text
-
-## The operating model
-
-## MSE design
-
-## The MSE process
-
-<img src="DLMtool_process.jpg" width="400">
-
-<br>
-
-#References
-
-```{r, echo=FALSE, eval=FALSE}
-sfStop()                            
+```{r, eval = FALSE}
+DLMtool::userguide()
 ```
 
-<br>
+The overall structure of a closed-loop simulation for a MSE in DLMtool can be described through the classes deployed in the package.
+An operating model (class __OM__ object) parameterizes the stock (biology), fleet, observation, and implementation dynamics of the MSE.
+To begin the MSE, the __OM__ object generates the data to be used for a management procedure and stores them in an object of class __Data__.
+A function of class __MP__ uses the __Data__ object and returns an object of class __Rec__ containing the management recommendations, e.g., effort and/or catch limits.
+From the management recommendations, the __OM__ object is updated and the process is successively repeated.
+The output for the closed-loop simulation is returned in an object of class __MSE__.
+
+<img src="DLMtool_design.png" width=400>
+
+_Figure 1: Design of DLMtool package for management strategy evaluations_
+
+## Overview of MSEtool
+
+To facilitate a standardized design for data-rich MPs, MSEtool builds out the __MP__ class with constituent functions of class __Assess__ and __HCR__.
+An __MP__ can be made from these functions to be tested in a MSE within the DLMtool framework.
+
+Additionally, __Assess__ functions can be used in an assessment-only context (external to management strategy evaluations).
+Output of an __Assess__ function can be used to provide standardized stock assessment reporting and diagnostics, such as Kobe plots, likelihood profiles for parameter estimates, and retrospective analyses of time series estimates of biomass, fishing mortality, recruitment, etc.
+
+<img src="MSEtool_design.png" width=690>
+
+_Figure 2: Design of MSEtool package for building management procedures from assessment models and harvest control rules for management strategy evaluations with the DLMtool package.
+Separate functions are also available for standardized reporting of output from assessment models in a stock assessment setting._
+
+MSEtool provides functions for converting the results of assessment models into DLMtool __OM__ and __Data__ objects.
+The SS2OM and SS2Data functions convert SS (Stock Synthesis) assessments, while iSCAM2OM and iSCAM2Data functions are provided for iSCAM assessments.
+
+Summary additional features created by Tom... for movement.
+
+# Assessment models
+
+Assessment models in MSEtool are functions of class __Assess__. 
+Currently, six assessment functions are available: __DD_TMB__ (Delay-Difference model and its state-space counterpart, __DD_SS__), __SP__ (Surplus Production model and its state-space counterpart, __SP_SS__), and __SCA__ (Statistical Catch at age model and another variant __SCA2__).
+These functions take a __Data__ object as an argument and return an object of class __Assessment__, for example,
+
+```{r, eval = FALSE}
+DD_result <- DD_TMB(1, sim_snapper, ...)
+```
+where `...` are model configuration arguments.
+
+The __Assessment__ object contains the raw output from TMB estimation model, optimization output from `nlminb`, and parameter estimates with covariance matrix (slots obj, opt, and SD, respectively).
+If applicable, time series estimates of total biomass, vulnerable biomass, spawning stock biomass, fishing mortality and/or exploitation rate, as well as their corresponding MSY and virgin reference points.
+The __Assess__ object also reports likelihood values and predicted values for the data used in the assessment model.
+
+Descriptions of the __Assess__ functions are provided in a separate Appendix vignette:
+
+```{r, eval = FALSE}
+vignette("Assessment", package = "MSEtool")
+```
+
+## Reporting and diagnostic functions
+
+Assessment results and summary, along with diagnostic procedures are available for each assessment model.
+Generic functions `summary` and `plot` are available for __Assessment__ objects.
+The `summary` function returns a list of estimated current stock status ($F/F_{MSY}$ , $B/B_{MSY}$ in the terminal year), input parameters to the model, derived quantities, and parameter estimates.
+The `plot` function generates time series figures of data, fitted data, and estimated quantities (e.g., biomass and recruitment).
+By default, these figures are save to disk and can be viewed with a HTML webpage generated by the function.
+
+```{r, eval = FALSE}
+summary(DD_result)
+plot(DD_result, save_figure = TRUE, save_dir = getwd()) # By default, also saves figures in the working directory and HTML browser for viewing
+```
+
+Two diagnostic functions are available: `retrospective` for retrospective analyses (re-running the model by sequentially removing terminal years of data) and `profile_likelihood` for profiling the likelihood function over a grid of parameters.
+
+```{r, eval = FALSE}
+retrospective(DD_result, 5) # Retrospective analysis going back 5 years from current year
+profile_likelihood(DD_result, UMSY = seq(0.05, 0.15, 0.01), MSY = seq(1, 3, 0.1)) # Joint profile over grid of UMSY and MSY
+```
+<img src="DD_profile.png" width=400>
+
+_Figure 3: Joint profile of UMSY and MSY from Delay-Difference assessment model (DD_TMB function). Contours indicate change in likelihood values from minimum (red point)._
+
+# Harvest control rules (HCRs)
+
+Harvest control rules in MSEtool are functions of class __HCR__, and are designed to provide different for a given assessment model.
+__HCR__ functions take an object of class __Assessment__ and return an object of class __Rec__.
+Currently, three __HCR__ functions are available as output controls (catch limits): HCR_MSY, HCR40_10, and HCR60_20.
+HCR_MSY prescribes the TAC to be the catch predicted with fishing at the estimated $F_{MSY}$ (the product of $F_{MSY}$ and estimated vulnerable biomass), while HCR40_10 and HCR60_20 are ramped control rules which reduce the $F_{MSY}$ catch limit based on spawning depletion.
+
+```{r, echo = FALSE, fig.width = 5, fig.height = 4}
+par(mar = c(5,5,3,1))
+Brel <- seq(0, 1, length.out = 200)
+plot(Brel, MSEtool::HCRlin(Brel, 0.1, 0.4), xlab = "Estimated SSB/SSB0", ylab = "TAC adjustment factor \n(proportion of FMSY catch)", main = "40-10 harvest control rule", type = "l", col = "blue")
+abline(v = c(0.1, 0.4), col = "red", lty = 2)
+```
+
+_Figure 4: The 40-10 harvest control rule prescribes an additional reduction in the TAC based on estimated spawning depletion._
+
+# Creating MPs from assessments and HCRs
+
+The `make_MP` function is a convenient way to stitch together a MP from an assessment model and HCR. The resulting function can then be passed to `DLMtool::runMSE()` for a management procedure.
+
+```{r, eval = FALSE}
+DD_MSY <- make_MP(DD_TMB, HCR_MSY, ...)
+DD_4010 <- make_MP(DD_TMB, HCR40_10)
+myMSE <- DLMtool::runMSE(OM = DLMtool::testOM, MPs = c("FMSYref", "DD_MSY", "DD_4010"))
+Tplot2(myMSE)
+```
+
+<img src="DD_MSE.png">
+
+_Figure 5. Trade-off plot of the MSE for the MPs that use the Delay-Difference model with two harvest control rules and the FMSY reference MP._
+
+Model configuration arguments can be passed through `...` when the MP is made.
+For example, the Ricker stock-recruitment relationship can be used in the Delay-Difference model instead of the default Beverton-Holt relationship with argument `SR = "Ricker"`:
+
+```{r, eval = FALSE}
+DD_MSY_Ricker <- make_MP(DD_TMB, HCR_MSY, SR = "Ricker")
+```
+
+This MP will call `DD_TMB(SR = "Ricker")`.
+
+## Diagnostic of assessment models in MSE
+
+By default, assessment model results are not saved during the MSE. However, this feature can be turned on with argument `diagnostic = TRUE` in `make_MP`:
+
+```{r, eval = FALSE}
+DD_MSY <- make_MP(DD_TMB, HCR_MSY, diagnostic = TRUE)
+```
+
+For each MP, two diagnostic objects are saved and reported in the `DLMtool::DLMenv` environment: the system time needed to run each assessment call and the __Assessment__ objects from each assessment call.
+The timing information may be useful to estimate total run time for additional MSE simulations or identify iterations in which the model has trouble with convergence.
+The __Assessment__ objects are used to obtain convergence information, as well as retrospective estimates of biomass and mortality over successive applications.
+
+Currently, this feature is memory-intensive and may cause performance issues or crashes.
+If the MSE is crashing with large number of simulations, a solution may be to run `DLMtool::runMSE()` in several batches with different seed numbers.
+MSE performance over all batches can be calculated by stitching the class __MSE__ objects together with `DLMtool::joinMSE()`.
+
+The `diagnosic_AM` function plots the timing and convergence information, and `retrospective_AM` plots the assessment model's retrospective patterns (for a single simulation in the MSE).
+
+```{r, eval = FALSE}
+DD_MSY <- make_MP(DD_TMB, HCR_MSY, diagnostic = TRUE)
+myMSE <- DLMtool::runMSE(OM = DLMtool::testOM, MPs = c("FMSYref", "DD_MSY", "DD_4010"))
+diagnostic_AM(myMSE)
+retrospective_AM(myMSE, sim = 1, MP = "DD_MSY")
+```
+<img src="DD_diag_AM.png">
+
+_Figure 6.
+Convergence and runtime diagnostics for DD_MSY MP during the MSE.
+In MSEtool, convergence is defined by a positive-definite Hessian matrix.
+In the case of non-convergence, the previous management recommendation is used._
+
+<img src="DD_retro_AM.png">
+
+_Figure 7.
+Retrospective analysis of the Delay-Difference model in the DD_MSY MP during the MSE for simulation #1.
+Operating model (true) values of spawning biomass (SSB), relative spawning biomass (SSB_SSBMSY), spawning biomass depletion (SSB_SSB0), fishing mortality (F), relative F (F_FMSY), and vulnerable bimoass (VB) are in dark black.
+Colored lines indicate model estimates sequentially over the projection years.
+Note that this assessment model assumes the vulnerable biomass to be the spawning biomass._
+
+## Obtaining management recommendations
+
+Like all other output MPs, `DLMtool::TAC()` can be used to obtain a catch limit:
+```{r, eval = FALSE}
+DD_MSY <- make_MP(DD_TMB, HCR_MSY)
+snapper_TAC <- TAC(sim_snapper, MPs = c("DD_MSY", "DD_4010"))
+plot(snapper_TAC)
+```
+<img src="DD_TAC.png">
+
+_Figure 8. Stochastic TAC recommendations from the Delay-Difference model with two harvest control rules. DD_4010 prescribes zero catch because the estimated depletion is less than 0.1._

### 2018-07-17 inst/doc/MSEtool.R
Update vignettes for v 1.0.0

@@ -7,23 +7,23 @@ options(width = 650)
 #  DLMtool::userguide()
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-#  DD_result <- DD_TMB(1, sim_snapper, ...)
+#  DD_result <- DD_TMB(Data = Red_snapper, ...)
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 #  vignette("Assessment", package = "MSEtool")
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 #  summary(DD_result)
-#  plot(DD_result, save_figure = TRUE, save_dir = getwd()) # By default, also saves figures in the working directory and HTML browser for viewing
+#  plot(DD_result, save_figure = TRUE, save_dir = getwd()) # By default, also saves figures in the working directory and HTML browser for future viewing
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-#  retrospective(DD_result, 5) # Retrospective analysis going back 5 years from current year
-#  profile_likelihood(DD_result, UMSY = seq(0.05, 0.15, 0.01), MSY = seq(1, 3, 0.1)) # Joint profile over grid of UMSY and MSY
[…]

### 2018-07-25 inst/doc/MSEtool.Rmd
Update vignettes

@@
 title: "User guide to MSEtool"
-subtitle: "Development, Testing and Implementation of Management Procedures for Data-Rich Fisheries (v1.1)"
+subtitle: "Development, Testing and Implementation of Management Procedures for Data-Rich Fisheries"
 author: "Tom Carruthers (<t.carruthers@oceans.ubc.ca>) 
@@
   rmarkdown::html_vignette:
-    toc: true
-    number_sections: true
-  rmarkdown::pdf_vignette:
     toc: true
@@
 
-The package uses the Management Strategy Evaluation engine of [DLMtool](http://www.datalimitedtoolkit.org) to undertake closed-loop simulation testing of management procedures.
-The DLMtool package was developed at the University of British Columbia in collaboration with the Natural Resources Defense Council.
-
+The [DLMtool](http://www.datalimitedtoolkit.org) package was developed at the University of British Columbia in collaboration with the Natural Resources Defense Council.
 DLMtool development has been funded by the Gordon and Betty Moore Foundation, the Packard Foundation, U.S.
@@
 
-## New Additions to this Version (v1.1)
+## New additions to this version (1.0.0)
 
-* Assessment models (`Assess` functions)
-    * Statistical Catch at Age (SCA and SCA2)
-    * Delay Difference (DD_TMB)
-    * State-Space Delay Difference (DD_SS)
-    * Surplus Production (SP)
-    * State-Space Surplus Production (SP_SS)
+* Assessment models (__Assess__ class functions)
+    * Statistical Catch at Age (`SCA` and `SCA2`)
+    * Delay Difference (`DD_TMB`)
+    * State-Space Delay Difference (`DD_SS`)
+    * Surplus Production (`SP`)
+    * State-Space Surplus Production (`SP_SS`)
 
-* Visualization
-    * Standardized assessment reports
-    * Diagnostics for assessment models
+* Visualization functions
+    * Standardized assessment reports (`plot` and `summary`)
+    * Diagnostics for assessment models (`profile_likelihood` and `retrospective`)
+    * Diagnostics for assessment models within closed-loop simulation (AM, assessment in MSE) (`prelim_AM`, `diagnostic_AM`, and `retrospective_AM`)
 
 * Other features 
-    * Harvest control rules (`HCR` functions)
+    * Harvest control rules (__HCR__ class functions: `HCR_MSY`, `HCR40_10`, and `HCR60_20`)
     * `make_MP` function conveniently combines a __Assess__ and __HCR__ function to create a __MP__ function
-    * Movement matrices for parameterizing OM
-    * The MCMC outputs of iSCAM models are now used to characterise operating model uncertainty
-    * iSCAM2OM and iSCAM2Data functions rip from iSCAM assessments
-    * SS2OM and SS2Data functions updated for latest version of r4ss
+    * `simmov` function creates movement matrices for parameterizing multiple area operating models (age-independent)
+    * `iSCAM2OM` and `iSCAM2Data` functions import from iSCAM assessments. The MCMC outputs of iSCAM models are now used to characterise operating model uncertainty
+    * `SS2OM` and `SS2Data` functions import Stock Synthesis assessments using the r4ss package version 1.24.0 (latest CRAN version)
 
@@
 
+* Age-varying movement matrices for `simmov` to condition operating models
+
 # Introduction 
-
-The MSEtool package is an extension of the DLMtool R package for implementing management strategy evaluation (MSE) simulations.
-Whereas DLMtool is primarily designed for testing data-limited management procedures, MSEtool provides an object-oriented framework for testing data-rich management procedures (MPs), which typically involve combining an assessment model with a harvest control rule to provide a fishery management recommendation.
+The MSEtool package is an extension of the DLMtool R package for implementing closed-loop simulation testing of management procedures, a process often referred to as management strategy evaluation (MSE).
+Whereas DLMtool is primarily designed for testing data-limited management procedures, MSEtool provides an object-oriented framework for creating data-rich management procedures (MPs), which typically involve combining an assessment model with a harvest control rule to provide a fishery management recommendation.
 MSEtool management procedures are designed to be used with the operating model of DLMtool through the `runMSE` function.
@@
 To begin the MSE, the __OM__ object generates the data to be used for a management procedure and stores them in an object of class __Data__.
-A function of class __MP__ uses the __Data__ object and returns an object of class __Rec__ containing the management recommendations, e.g., effort and/or catch limits.
-From the management recommendations, the __OM__ object is updated and the process is successively repeated.
-The output for the closed-loop simulation is returned in an object of class __MSE__.
+A function of class __MP__ uses the __Data__ object and returns an object of class __Rec__ containing the management recommendations, e.g., some combination of spatial closures and size, effort and catch limits.
+From the management recommendations, the __OM__ object is updated and the process is successively repeated as the operating model progresses over time.
+The output for the closed-loop simulation is returned in an object of class __MSE__, from which summary statistics can be calculated.
 
@@
 
+### Additional features
 MSEtool provides functions for converting the results of assessment models into DLMtool __OM__ and __Data__ objects.
-The SS2OM and SS2Data functions convert SS (Stock Synthesis) assessments, while iSCAM2OM and iSCAM2Data functions are provided for iSCAM assessments.
+The `SS2OM` and `SS2Data` functions convert SS (Stock Synthesis) assessments, while `iSCAM2OM` and `iSCAM2Data` functions are provided for iSCAM assessments.
 
-Summary additional features created by Tom... for movement.
+To support multiple-area operating models, the `simmov` function estimates the movement matrix based on the prescribed number of areas, biomass distribution among areas in unfished conditions, and the probability of staying in each area in the next time step.
 
@@
 Assessment models in MSEtool are functions of class __Assess__. 
-Currently, six assessment functions are available: __DD_TMB__ (Delay-Difference model and its state-space counterpart, __DD_SS__), __SP__ (Surplus Production model and its state-space counterpart, __SP_SS__), and __SCA__ (Statistical Catch at age model and another variant __SCA2__).
+Currently, six assessment functions are available: `DD_TMB` (Delay-Difference model and its state-space counterpart, `DD_SS`), `SP` (Surplus Production model and its state-space counterpart, `SP_SS`), and `SCA` (Statistical Catch at age model and another variant `SCA2`).
 These functions take a __Data__ object as an argument and return an object of class __Assessment__, for example,
@@
 ```{r, eval = FALSE}
-DD_result <- DD_TMB(1, sim_snapper, ...)
+DD_result <- DD_TMB(Data = Red_snapper, ...)
 ```
@@
 
-The __Assessment__ object contains the raw output from TMB estimation model, optimization output from `nlminb`, and parameter estimates with covariance matrix (slots obj, opt, and SD, respectively).
+The __Assessment__ object contains the raw output from TMB estimation model, optimization output from `nlminb`, and parameter estimates with covariance matrix.
 If applicable, time series estimates of total biomass, vulnerable biomass, spawning stock biomass, fishing mortality and/or exploitation rate, as well as their corresponding MSY and virgin reference points.
-The __Assess__ object also reports likelihood values and predicted values for the data used in the assessment model.
+The __Assessment__ object also reports likelihood values and predicted values for the data used in the assessment model.
+The object documentation for the __Assessment__ object can be obtained through:
+
+```{r, eval = FALSE}
+class?Assessment
+```
 
@@
 summary(DD_result)
-plot(DD_result, save_figure = TRUE, save_dir = getwd()) # By default, also saves figures in the working directory and HTML browser for viewing
+plot(DD_result, save_figure = TRUE, save_dir = getwd()) # By default, also saves figures in the working directory and HTML browser for future viewing
 ```
@@
 ```{r, eval = FALSE}
-retrospective(DD_result, 5) # Retrospective analysis going back 5 years from current year
-profile_likelihood(DD_result, UMSY = seq(0.05, 0.15, 0.01), MSY = seq(1, 3, 0.1)) # Joint profile over grid of UMSY and MSY
+retrospective(DD_result, nyr = 5) # Retrospective analysis going back 5 years from current year
+profile_likelihood(DD_result, R0 = seq(0.75, 1.25, 0.025), h = seq(0.95, 1, 2.5e-3)) # Joint profile over grid of R0 and steepness
 ```
@@
 
-_Figure 3: Joint profile of UMSY and MSY from Delay-Difference assessment model (DD_TMB function). Contours indicate change in likelihood values from minimum (red point)._
+_Figure 3: Joint profile of R0 and steepness from Delay-Difference assessment model (DD_TMB function). Contours indicate change in likelihood values from minimum (red point)._
 
@@
 Currently, three __HCR__ functions are available as output controls (catch limits): HCR_MSY, HCR40_10, and HCR60_20.
-HCR_MSY prescribes the TAC to be the catch predicted with fishing at the estimated $F_{MSY}$ (the product of $F_{MSY}$ and estimated vulnerable biomass), while HCR40_10 and HCR60_20 are ramped control rules which reduce the $F_{MSY}$ catch limit based on spawning depletion.
+HCR_MSY prescribes the TAC to be the catch predicted with fishing at the estimated $F_{MSY}$, i.e., the product of $F_{MSY}$ and estimated vulnerable biomass, while HCR40_10 and HCR60_20 are ramped control rules which reduce the $F_{MSY}$ catch limit based on spawning depletion $SSB/SSB_0$.
 
@@
 Brel <- seq(0, 1, length.out = 200)
-plot(Brel, MSEtool::HCRlin(Brel, 0.1, 0.4), xlab = "Estimated SSB/SSB0", ylab = "TAC adjustment factor \n(proportion of FMSY catch)", main = "40-10 harvest control rule", type = "l", col = "blue")
+plot(Brel, MSEtool::HCRlin(Brel, 0.1, 0.4), xlab = expression(Estimated~~SSB/SSB[0]), ylab = "TAC adjustment factor \n(proportion of FMSY catch)", main = "40-10 harvest control rule", type = "l", col = "blue")
 abline(v = c(0.1, 0.4), col = "red", lty = 2)
@@
 
-<img src="DD_MSE.png">
+<img src="DD_Tplot2.png">
 
@@
 
-Model configuration arguments can be passed through `...` when the MP is made.
-For example, the Ricker stock-recruitment relationship can be used in the Delay-Difference model instead of the default Beverton-Holt relationship with argument `SR = "Ricker"`:
+Model configuration arguments can be passed through `...` in `make_MP` when the MP is made.
+For example, the Ricker stock-recruitment relationship can be used in the Delay-Difference model instead of the default Beverton-Holt relationship and steepness can be fixed instead of esimated by passing argument `SR = "Ricker"` and `fix_h = TRUE`, respectively:
 
 ```{r, eval = FALSE}
-DD_MSY_Ricker <- make_MP(DD_TMB, HCR_MSY, SR = "Ricker")
+DD_MSY_Ricker_fixsteep <- make_MP(DD_TMB, HCR_MSY, SR = "Ricker", fix_h = TRUE)
 ```
@@
 
-By default, assessment model results are not saved during the MSE. However, this feature can be turned on with argument `diagnostic = TRUE` in `make_MP`:
+Three functions are designed to evaluate the performance of __Assess__ models in MSE: `prelim_AM`, `diagnostic_AM`, and `retrospective_AM` (where _AM_ stands for "assessment in MSE").
+
+`prelim_AM` is designed to evaluate the configuration of the assessment model before running the closed-loop simulation.
+Given an operating model, this function generates data from the historical period of the MSE, applies the assessment model to those data, then returns the convergence rate of the model.
+A list of __Assessment__ objects are returned and can be used for troubleshooting.
+Additional arguments are passed to the __Assess__ function via `...`:
 
 ```{r, eval = FALSE}
-DD_MSY <- make_MP(DD_TMB, HCR_MSY, diagnostic = TRUE)
+prelim_AM(DLMtool::testOM, DD_TMB, ...)
+#> Generating Hist object from OM object via runMSE:
+#> Loading operating model
+#> Optimizing for user-specified movement
+#> Optimizing for user-specified depletion
+#> Calculating historical stock and fishing dynamics
+#> Calculating MSY reference points
+#> Calculating B-low reference points
+#> Calculating reference yield - best fixed F strategy
+#> Returning historical simulations
+#> Running DD_TMB with 48 simulations for DLMtool::testOM.
+#> Assessments complete.
+#> Total time to run 48 assessments: 1.2 seconds
+#> 2 of 48 simulations (4.2%) did not converge.
+#> See simulation number: 11 21
 ```
 
-For each MP, two diagnostic objects are saved and reported in the `DLMtool::DLMenv` environment: the system time needed to run each assessment call and the __Assessment__ objects from each assessment call.
-The timing information may be useful to estimate total run time for additional MSE simulations or identify iterations in which the model has trouble with convergence.
-The __Assessment__ objects are used to obtain convergence information, as well as retrospective estimates of biomass and mortality over successive applications.
+By default, assessment model results are not saved during the MSE (called by `DLMtool::runMSE`). However, this feature can be turned on with argument `diagnostic` in `make_MP`:
 
-Currently, this feature is memory-intensive and may cause performance issues or crashes.
-If the MSE is crashing with large number of simulations, a solution may be to run `DLMtool::runMSE()` in several batches with different seed numbers.
+```{r, eval = FALSE}
+DD_MSY <- make_MP(DD_TMB, HCR_MSY, diagnostic = "full")
+```
+
+For each MSEtool MP, two diagnostic objects can be saved and reported in the `DLMtool::DLMenv` environment: convergence diagnostics and the __Assessment__ objects from each assessment call.
+The first can be obtained with `diagnostic = "min"`, while both are returned with `diagnostic = "full"`.
+
+MSEtool MPs may be memory-intensive and cause performance issues or crashes.
+With large number of simulations, a solution may be to run `DLMtool::runMSE()` in several batches with different seed numbers.
 MSE performance over all batches can be calculated by stitching the class __MSE__ objects together with `DLMtool::joinMSE()`.
 
-The `diagnosic_AM` function plots the timing and convergence information, and `retrospective_AM` plots the assessment model's retrospective patterns (for a single simulation in the MSE).
+The `diagnosic_AM` function plots the convergence information, and `retrospective_AM` plots the assessment model's retrospective patterns (for a single simulation in the MSE).
 
 ```{r, eval = FALSE}
-DD_MSY <- make_MP(DD_TMB, HCR_MSY, diagnostic = TRUE)
+DD_MSY <- make_MP(DD_TMB, HCR_MSY, diagnostic = "full")
+DD_4010 <- make_MP(DD_TMB, HCR40_10)
 myMSE <- DLMtool::runMSE(OM = DLMtool::testOM, MPs = c("FMSYref", "DD_MSY", "DD_4010"))
@@
 ```
-<img src="DD_diag_AM.png">
+<img src="DD_diagnostic_AM.png">
 
 _Figure 6.
-Convergence and runtime diagnostics for DD_MSY MP during the MSE.
-In MSEtool, convergence is defined by a positive-definite Hessian matrix.
+Convergence diagnostics (return by TMB objects and nlminb function) for DD_MSY MP during the MSE.
+Here, the assessment model was applied beginning in year 50 of the MSE and repeated every 4th year.
+In MSEtool, convergence is defined by a positive-definite Hessian matrix (top middle figure).
 In the case of non-convergence, the previous management recommendation is used._
 
-<img src="DD_retro_AM.png">
+<img src="DD_retrospective_AM.png">
 
@@
 
+Once it is determined that the MSEtool assessment models are performing adequately (no major convergence issues), then the performance of the MPs that use these models can be evaluated with the DLMtool package.
+
 ## Obtaining management recommendations
@@
 DD_MSY <- make_MP(DD_TMB, HCR_MSY)
-snapper_TAC <- TAC(sim_snapper, MPs = c("DD_MSY", "DD_4010"))
+DD_4010 <- make_MP(DD_TMB, HCR40_10)
+snapper_TAC <- TAC(Red_snapper, MPs = c("DD_MSY", "DD_4010"))
 plot(snapper_TAC)
@@
 
-_Figure 8. Stochastic TAC recommendations from the Delay-Difference model with two harvest control rules. DD_4010 prescribes zero catch because the estimated depletion is less than 0.1._
+_Figure 8.
+Stochastic TAC recommendations from the Delay-Difference model with two harvest control rules.
+When the 40-10 harvest control rule is applied, the prescribed TAC is close to zero because the estimated terminal year biomass is 0.13 in the assessment model._

### 2018-07-25 R/diagnostic_AM.R

@@ -35,12 +35,13 @@ prelim_AM <- function(x, Assess, ncpus = 1, ...) {
       }
 
   sfInit(parallel = ncpus > 1, cpus = ncpus)
+  sfLibrary("MSEtool", character.only = TRUE)
   nsim <- nrow(Data@Cat)
   message(paste0("Running ", deparse(substitute(Assess)), " with ", nsim, " simulations for ", deparse(substitute(x)), "."))
   dots <- list(...)
   if(length(dots) > 0) message(paste0("\nAdditional arguments to be provided to ", deparse(substitute(x)), ":\n", paste(names(dots), collapse = "\n")))
   Assess <- match.fun(Assess)
-  if(!inherits(Assess, "Assess")) stop(paste(deparse(substitute(x))), "does not appear to be an Assess function.")
+  if(!inherits(Assess, "Assess")) stop(paste(deparse(substitute(Assess))), "does not appear to be an Assess function.")
 
   if(sfParallel()) sfExport(list = c("Assess", "Data"))
   timing <- proc.time()

### 2018-07-25 inst/doc/MSEtool.R

@@ -10,6 +10,9 @@ options(width = 650)
 #  DD_result <- DD_TMB(Data = Red_snapper, ...)
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
+#  class?Assessment
+
+## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 #  vignette("Assessment", package = "MSEtool")
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
@@ -33,19 +36,38 @@ abline(v = c(0.1, 0.4), col = "red", lty = 2)
 #  Tplot2(myMSE)
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-#  DD_MSY_Ricker <- make_MP(DD_TMB, HCR_MSY, SR = "Ricker")
+#  DD_MSY_Ricker_fixsteep <- make_MP(DD_TMB, HCR_MSY, SR = "Ricker", fix_h = TRUE)
 
[…]

### 2018-08-03 inst/doc/MSEtool.Rmd
Update vignette intro

@@
 
-# Foreword
-
-MSEtool development is funded by the Canadian Department of Fisheries and Oceans and benefits from ongoing collaborations with the Natural Resources Defense Council and number of Canadian government scientists including Robyn Forrest, Sean Anderson and Daniel Duplisea.
-
-The [DLMtool](http://www.datalimitedtoolkit.org) package was developed at the University of British Columbia in collaboration with the Natural Resources Defense Council.
-DLMtool development has been funded by the Gordon and Betty Moore Foundation, the Packard Foundation, U.S.
-National Oceanic and Atmospheric Administration, Fisheries and Oceans Canada, the Walton Foundation, Resources Legacy Fund, the Natural Resources Defense Council, the United Nations Food and Agricultural Organization and the California Department of Fish and Wildlife.
-
-<a href="http://www.datalimitedtoolkit.org"><img src="DLMtool_logo.png" width=250></a>
-
 # Version Notes
@@
     * Virtual Population Analysis (VPA)
-    * Age-Structure Production Models (ASPM)
     * Kalman filter assessment approximation
@@
 
+* Multi-stock and multi-fleet operating models
+
 # Introduction 
-The MSEtool package is an extension of the DLMtool R package for implementing closed-loop simulation testing of management procedures, a process often referred to as management strategy evaluation (MSE).
+MSEtool is an extension of the DLMtool package for implementing closed-loop simulation testing of management procedures, a process often referred to as management strategy evaluation (MSE).
+The components of the operating model for conducting MSEs for data-rich stocks is wholly contained within DLMtool, but MSEtool provides the tools for implementing additional features to facilitate such MSEs.
 Whereas DLMtool is primarily designed for testing data-limited management procedures, MSEtool provides an object-oriented framework for creating data-rich management procedures (MPs), which typically involve combining an assessment model with a harvest control rule to provide a fishery management recommendation.
-MSEtool management procedures are designed to be used with the operating model of DLMtool through the `runMSE` function.
+Management procedures created by MSEtool are designed to be used with the operating model of DLMtool through the `runMSE` function.
+
 The user guide to the DLMtool package is accessible via
@@
 
-To facilitate a standardized design for data-rich MPs, MSEtool builds out the __MP__ class with constituent functions of class __Assess__ and __HCR__.
-An __MP__ can be made from these functions to be tested in a MSE within the DLMtool framework.
-
 Additionally, __Assess__ functions can be used in an assessment-only context (external to management strategy evaluations).
@@
 _Figure 2: Design of MSEtool package for building management procedures from assessment models and harvest control rules for management strategy evaluations with the DLMtool package.
+Assess and HCR functions and the Assessment object are unique to MSEtool, but work within the framework of DLMtool objects.
 Separate functions are also available for standardized reporting of output from assessment models in a stock assessment setting._
@@
 When the 40-10 harvest control rule is applied, the prescribed TAC is close to zero because the estimated terminal year biomass is 0.13 in the assessment model._
+
+# Acknowledgements
+
+MSEtool development is funded by the Canadian Department of Fisheries and Oceans and benefits from ongoing collaborations with the Natural Resources Defense Council and number of Canadian government scientists including Robyn Forrest, Sean Anderson and Daniel Duplisea.
+
+The [DLMtool](http://www.datalimitedtoolkit.org) package was developed at the University of British Columbia in collaboration with the Natural Resources Defense Council.
+DLMtool development has been funded by the Gordon and Betty Moore Foundation, the Packard Foundation, U.S.
+National Oceanic and Atmospheric Administration, Fisheries and Oceans Canada, the Walton Foundation, Resources Legacy Fund, the Natural Resources Defense Council, the United Nations Food and Agricultural Organization and the California Department of Fish and Wildlife.
+
+<a href="http://www.datalimitedtoolkit.org"><img src="DLMtool_logo.png" width=250></a>

### 2018-08-13 inst/doc/MSEtool.Rmd
Use tempdir() instead of getwd()

@@
 subtitle: "Development, Testing and Implementation of Management Procedures for Data-Rich Fisheries"
-author: "Tom Carruthers (<t.carruthers@oceans.ubc.ca>) 
-         Adrian Hordyk  (<a.hordyk@oceans.ubc.ca>),
-         Quang Huynh (<q.huynh@oceans.ubc.ca>)"
+author: "Quang Huynh (<q.huynh@oceans.ubc.ca>), Tom Carruthers (<t.carruthers@oceans.ubc.ca>), 
+         Adrian Hordyk (<a.hordyk@oceans.ubc.ca>)"
 date: "`r Sys.Date()`"
@@
 
+To facilitate a standardized design for data-rich MPs, MSEtool creates new classes to build out the __MP__ class.
+MSEtool uses functions of class __Assess__ and __HCR__ for an assessment model and harvest control rule, respectively.
+An __MP__ class function can be made from these functions to be tested in a MSE within the DLMtool framework.
+
 Additionally, __Assess__ functions can be used in an assessment-only context (external to management strategy evaluations).
@@
 summary(DD_result)
-plot(DD_result, save_figure = TRUE, save_dir = getwd()) # By default, also saves figures in the working directory and HTML browser for future viewing
+plot(DD_result, save_figure = TRUE, save_dir = tempdir()) # By default, also saves figures in a temporary directory for viewing through a HTML browser.
+The directory can be changed to a user's filespace of choice.
 ```
@@
 
-This MP will call `DD_TMB(SR = "Ricker")`.
+This MP will call `DD_TMB(..., SR = "Ricker", fix_h = TRUE)`.
 

### 2018-08-13 R/report_core.R

@@ -1,28 +1,34 @@
 
 
-#' Perform profile likelihood
+#' Profile likelihood of assessment models
 #'
-#' Performs a profile likelihood.
+#' Profile the likelihood for leading parameters of assessment models.
 #'
-#' @param Assessment An S4 object of class Assessment.
+#' @param Assessment An S4 object of class \linkS4class{Assessment}.
 #' @param figure Indicates whether a figure will be plotted.
 #' @param save_figure Indicates whether figures will be saved to directory.
 #' @param save_dir The directory to which figures will be saved. By default: \code{getwd()}
 #' @param ... A sequence of values of the parameter(s) for the profile.
 #' See details for name of arguments to be passed on.
-#' @details For \code{DD_TMB}, sequence of values for \code{UMSY} and \code{MSY} are needed.
+#' @details For the following assessment models, the required sequence of values are:
+#' \itemize{
+#' \item \code{DD_TMB} and \code{DD_SS}: \code{R0} and \code{h}
+#' \item \code{SP} and \code{SP_SS}: \code{UMSY} and \code{MSY}
+#' \item \code{SCA}: \code{R0} and \code{h}
+#' \item \code{SCA2}: \code{meanR}
+#' }
 #' @author Q. Huynh
-#' @return A data frame of negative log-likelihood values from profile and, optionally,
+#' @return A data frame of negative log-likelihood values from the profile and, optionally,
 #' a figure of the likelihood surface.
 #' @examples
 #' \donttest{
 #' output <- DD_TMB(Data = DLMtool::Red_snapper)
 #' pro <- profile_likelihood(output, R0 = seq(0.75, 1.25, 0.025), h = seq(0.9, 0.99, 0.01),
 #' save_figure = FALSE)
-#' }
 #'
 #' # Ensure your grid is of proper resolution. A grid that is too coarse will distort the shape of
 #' # the likelihood surface.
+#' }
 #' @export
 profile_likelihood <- function(Assessment, figure = TRUE, save_figure = TRUE,
                                save_dir = tempdir(), ...) {
@@ -40,11 +46,12 @@ profile_likelihood <- function(Assessment, figure = TRUE, save_figure = TRUE,
 }
 
 
-#' Perform retrospective analysis
+#' Retrospective analysis of assessment models
 #'
-#' Perform a retrospective analysis.
+#' Perform a retrospective analysis, successive removals of most recent years of data to evaluate resulting
+#' parameter estimates.
 #'
-#' @param Assessment An S4 object of Assessment output..
[…]

### 2018-08-13 inst/doc/MSEtool.R

@@ -17,7 +17,7 @@ options(width = 650)
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 #  summary(DD_result)
-#  plot(DD_result, save_figure = TRUE, save_dir = getwd()) # By default, also saves figures in the working directory and HTML browser for future viewing
+#  plot(DD_result, save_figure = TRUE, save_dir = tempdir()) # By default, also saves figures in a temporary directory for viewing through a HTML browser. The directory can be changed to a user's filespace of choice.
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 #  retrospective(DD_result, nyr = 5) # Retrospective analysis going back 5 years from current year

### 2018-08-13 inst/doc/Assessment.Rmd

---
title: "Appendix: Description of the Assessment Models in MSEtool"
author: "Tom Carruthers (<t.carruthers@oceans.ubc.ca>) 
         Adrian Hordyk  (<a.hordyk@oceans.ubc.ca>),
         Quang Huynh (<q.huynh@oceans.ubc.ca>)"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Appendix: Description of the Assessment Models in MSEtool}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---
<script type="text/x-mathjax-config">
  MathJax.Hub.Config({ TeX: { equationNumbers: {autoNumber: "all"} } });
</script>

<style type="text/css">

body{ /* Normal  */
   font-size: 12px;
}
td {  /* Table  */
   font-size: 8px;
}
h1 { /* Header 1 */
 font-size: 18px;
 color: DarkBlue;
}
h2 { /* Header 2 */
 font-size: 15px;
 color: DarkBlue;
}
h3 { /* Header 3 */
 font-size: 14px;
 color: DarkBlue;
}
code.r{ /* Code block */
  font-size: 10px;
}
pre { /* Code block */
  font-size: 10px
}
</style>

```{r set options, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
knitr::opts_chunk$set(dpi=85)
options(width = 650)
```

<br>

# Introduction

In MSEtool, assessment models are of class `Assess`. This appendix provides a brief description and references for the `Assess` objects. Further details regarding parameterization, e.g., fixing parameters, and tuning, e.g., adjusting start parameters, are provided in the function documentation.

## Note

For LaTeX equation rendering, it is recommended that this vignette be viewed in a HTML browser. This can be done with the `browseVignettes` function in R:

```{r, eval = FALSE}
browseVignettes("MSEtool")
```

# Delay Difference (DD_TMB) Model

There has been a rich history of development for the delay difference model for catch and index data. For the formulation used in MSEtool, the most relevant citations are Chapter 9 of Hilborn and Walters (1992) and Carruthers et al. (2012).

## Growth
Growth in weight-at-age $w_a$ follows the recursive Ford-Brody equation:
$$w_a = \rho + \tilde\alpha w_{a-1}.$$
We can obtain $\tilde\alpha$ and $\rho$ for the delay difference model if weight is also described by the equation 
$$w_a = W_{\infty}(1 - \exp[K\{a-a_0\}])^b.$$ 
Parameter $\tilde\alpha$ is solved in the limiting case where $w_a = w_{a-1} = W_{\infty}$ as $t \rightarrow \infty$,
$$\tilde\alpha = W_{\infty}(1 - \rho). $$
Substitution of equation 3 into equation 1 solves for $\rho$, 
$$\rho = \dfrac{w_a - W_{\infty}}{w_{a-1} - W_{\infty}}.$$
In MSEtool, $a = k+2$ is arbitrarily chosen to calculate $\rho$, where $k$ is the age of knife-edge selectivity. From catch and effort data alone, the age corresponding to the length of 50% maturity is chosen for $k$.

## Dynamics equations
The population biomass $B_t$ and abundance $N_t$ in year $t$ is given by
$$ \begin{align}
B_t &= s_{t-1}(\tilde{\alpha} N_{t-1} + \rho B_{t-1}) + w_k R_t\\
N_t &= s_{t-1} N_{t-1} + R_t,
\end{align}$$
where $R_t$ is the recruitment (defined in the next section) at age $k$ and $w_k$ is the weight of recruits. Survival $s_t$ is defined as
$$ s_t = (1 - u_t)\exp(-M), $$
where $u_t$ is the annual harvest rate and $M$ is the instantaneous natural mortality rate. By conditioning the model on effort (as the ratio of the catch and index), the harvest rate is
$$ u_t = 1 - \exp(-qf_t), $$
where $q$ is the estimated catchability coefficient and $f_t$ is the effort in year $t$.

The predicted catch is 
$$\hat{C}_t = \hat{u}_t \hat{B}_t ,$$
where the circumflex $^$ denotes the model estimate.

The log-likelihood $L$ of the observed catch $C_t$ over years $t = 1, ..., T$, assuming a lognormal distribution, is
$$L(C_t)\propto -T \log\sigma - 0.5\sum_{t=1}^{T}\left(\dfrac{\log(C_t/\hat{C}_t)}{\sigma}\right)^2,$$
where $\sigma$ is the standard deviation of the lognormal distribution.

## Stock-recruit parameters

### Beverton-Holt relationship
Assuming a Beverton-Holt stock recruit relationship and spawning occurring after fishing in each annual time step, then recruitment (at age $k$) in year $t$ is:
$$ R_t = \dfrac{\alpha B_{t-k}(1 - u_{t-k})}{1 + \beta B_{t-k}(1 - u_{t-k})},$$
where $B_{t-k}(1-u_{t-k})$ is the population biomass after fishing in year $t-k$ and
$$ \begin{align}
\alpha &= \dfrac{4hR_0}{(1-h)B_0}\\
\beta &= \dfrac{5h-1}{(1-h)B_0},
\end{align}$$

Virgin recruitment $R_0$ and steepness $h$ are estimated parameters, with virgin biomass $B_0$ calculated as
$$B_0 =  R_0 \phi_0.$$
The virgin biomass per recruit $\phi_0$ is
$$\phi_0 = \dfrac{\tilde{\alpha} \exp(-M) + w_k (1 - \exp(-M))}{1 - \rho \exp(-M)}$$
and is obtained by solving the equilibrium equation for biomass, $B_0 = \exp(M)(\tilde{\alpha}N_0 + \rho B_0) + w_k R_0$, is solved for $B_0/R_0$, with $N_0 = R_0/(1−\exp(-M))$.

### Ricker equation
Assuming a Ricker stock-recruit relationship, the recruitment is
$$ R_t = \alpha B_{t-k}(1 - u_{t-k})\exp[-\beta B_{t-k}(1 - u_{t-k})],$$
where
$$ \begin{align}
\alpha &= \dfrac{(5h)^{1.25} R_0}{B_0}\\
\beta &= \dfrac{5}{4B_0}\log(5h),
\end{align}$$

and $B_0$ is calculated as in equation 14.

## State-space version (DD_SS)
In the state-space version, annual recruitment deviates from the stock-recruit relationship are estimated. The recruitment in year $t$ is
$$ R_t = \dfrac{\alpha B_{t-k}}{1 + \beta B_{t-k}} \exp(\delta_t - 0.5 \tau^2)$$
or
$$ R_t = \alpha B_{t-k}\exp(-\beta B_{t-k})\exp(\delta_t - 0.5 \tau^2),$$
where $\delta_t \sim N(0, \tau^2)$ are recruitment deviations in lognormal space and $\tau$ is the standard deviation of the recruitment deviations.

The log-likelihood of the log-recruitment deviations $\delta_t$ over years $t=k,\ldots...,T$ is
$$L(\delta_t) \propto -(T-k) \log\tau - 0.5\sum_{t=k}^{T}\left(\dfrac{\log(\delta_t)}{\tau}\right)^2.$$

# Surplus-production (SP) model
## Dynamics equations
The surplus production model uses the Fletcher (1978) formulation. The biomass $B_t$ in year $t$ is 
$$B_t = B_{t-1} + P_{t-1} - C_{t-1},$$
where $C_t$ is the observed catch and $P_t$ is the surplus production given by:
$$P_t = \gamma MSY \left(\dfrac{B_t}{K}-\left[\dfrac{B_t}{K}\right]^n\right), $$
where $K$ is the carrying capacity, $MSY$ is the estimated maximum sustainable yield, and $n$ is the parameter that controls shape of the production curve, and
$\gamma$ is
$$\gamma = \dfrac{1}{n-1}n^{n/(n-1)}.$$

By conditioning the model on observed catch, the predicted index $\hat{I}_t$ is
$$\hat{I}_t = \hat{q} \hat{B}_t $$
and the harvest rate is
$$\hat{u}_t = \dfrac{C_t}{\hat{B}_t}.$$
The log-likelihood $L$ of the observed index $I_t$ over years $t = 1, ..., T$, assuming a lognormal distribution, is
$$L(I_t)\propto -T \log\sigma - 0.5\sum_{t=1}^{T}\left(\dfrac{\log(I_t/\hat{I}_t)}{\sigma}\right)^2,$$

## Derivation of SP model parameters from MSY and UMSY estimates

From estimates of leading parameters $u_{MSY}$ and $MSY$, the biomass $B_{MSY}$ at $MSY$ is
$$B_{MSY} = \dfrac{MSY}{u_{MSY}},$$
the carrying capacity $K$ is
$$K = n^{1/(n-1)} B_{MSY} ,$$
and the intrinsic rate of population increase $r$ is
$$ r = n u_{MSY}.$$
The production parameter $n$ is typically fixed and the model has a symmetric productive curve ($B_{MSY}/K = 0.5$) when $n = 2$.

## State-space version (SP_SS)
In the state-state version, annual biomass deviates are estimated as random effects. Similar to Meyer and Millar (1999), the biomass $B_t$ in year $t$ is 
$$B_t = (B_{t-1} + P_{t-1} - C_{t-1})\exp(\delta_t - 0.5 \tau^2),$$
where $\delta_t \sim N(0, \tau^2)$ are biomass deviations in lognormal space and $\tau$ is the standard deviation of the biomass deviations.

The log-likelihood of the deviations $\delta_t$ is
$$L(\delta_t) \propto -T \log\tau - 0.5\sum_{t=1}^{T}\left(\dfrac{\log(\delta_t)}{\tau}\right)^2.$$
# Statistical catch-at-age model (SCA and SCA2) 
## Dynamics equations
The statistical catch-at-age model used in MSEtool assumes pulse fishing at the beginning of each annual time-step, following the parameterization of Forrest et al. (2008). The model uses a time series of total catch (in weight), index, and catch-at-age observations, as well as information on weight, maturity, natural mortality at age.

The population abundance $N_{a,t}$ of age $a$ in year $t$ is
$$ N_{a,t} = 
\begin{cases} 
      R_t & a = 1\\
      N_{a-1,t-1} (1 - v_{a-1} u_{t-1}) \exp(-M_{a-1}) & a = 2, \ldots, A-1\\
      N_{a-1,t-1} (1 - v_{a-1} u_{t-1}) \exp(-M_{a-1}) + N_{a,t-1} (1 - v_a u_{t-1}) \exp(-M_a) & a = A 
\end{cases},
$$
where $R_t$ is the recruitment (age-1), $v_a$ is the vulnerability at age $a$, $u_t$ is the annual harvest rate, $M_a$ is the instantaneous natural mortality rate at age $a$, and $A$ is the maximum age in the model as a plus-group accumulator age.

Assuming logistic vulnerability, the vulnerability is:
$$v_a = \left[1 + \exp\left(-\log(19) \dfrac{a - a_{50}}{a_{95} - a_{50}}\right)\right]^{-1}, $$
where $a_{50}$ and $a_{95}$ are the estimated ages of 50% and 95% vulnerability, respectively.

Assuming dome vulnerability, a double-normal formulation is used:
$$ v_a = 
\begin{cases} 
      f(a; \mu_{asc}, \sigma_{asc}) & a \le \mu_{asc}\\
      1 & \mu_{asc} \lt a \le \mu_{des}\\
      f(a; \mu_{des}, \sigma_{des}) & a \gt \mu_{des}
\end{cases},
$$
where $f(a; \mu, \sigma)$ is the normal probability density function scaled to one at $\mu$. The $\mu$ and $\sigma$ for both the ascending and descending limbs of the double-normal equation are estimated parameters.

The vulnerable biomass $VB_t$ in year $t$ is
$$VB_t = \sum_{a=1}^A v_a w_a N_{a,t},$$
where weight-at-age $w_a$ is given by 
$$w_a = W_{\infty}(1 - \exp[K\{a-a_0\}])^b.$$

The mature spawning biomass $E_t$ is given by
$$E_t = \sum_{a=1}^A m_a w_a N_{a,t},$$
where maturity at age $m_a$ is
$$m_a = \left[1 + \exp\left(-\log(19) \dfrac{a - \tilde{a}_{50}}{\tilde{a}_{95} - \tilde{a}_{50}}\right)\right]^{-1}, $$
where $\tilde{a}_{50}$ and $\tilde{a}_{95}$ are the ages of 50% and 95% maturity, respectively.

By conditioning the model on catch in weight $Y_t$, the estimated annual harvest rate $\hat{u}_t$ is
$$\hat{u}_t = Y_t / \widehat{VB}_t .$$
The estimated catch at age $\hat{C}_{a,t}$ is
$$\hat{C}_{a,t} = \hat{v}_a \hat{u}_t \hat{N}_{a,t}.$$
The log-likelihood $L$ of the observed catch at age $C_{a,t}$ over years $t = 1, ..., T$, assuming a multinomial distribution, is
$$L(C_{a,t})\propto \sum_{t=1}^T \left(N_t \sum_{a=1}^A p_{a,t} \log(\hat{p}_{a,t}) \right),$$
where $N_t$ is the sample size of catch-at-age observations in year $t$, $p_{a,t}$ and $\hat{p}_{a,t}$ are the observed and predicted
proportions of catch-at-age.

The estimated index $\hat{I}_t$, assuming that it is an index for total biomass, is
$$ \hat{I}_t = \hat{q} \hat{B}_t,$$
where $B_t = \sum_{a=1}^A w_a N_{a,t}$. A function argument allows for the user to specify that the observed index is for vulnerable or spawning biomass.

The log-likelihood $L$ of the observed index $I_t$ over years $t = 1, ..., T$, assuming a lognormal distribution, is
$$L(I_t)\propto -T \log\sigma - 0.5\sum_{t=1}^{T}\left(\dfrac{\log(I_t/\hat{I}_t)}{\sigma}\right)^2.$$

## Estimation of recruitment and reference points

There are two variants of the statistical catch-at-age model for estimation of recruitment and reference points. In function `SCA`, productivity parameters $R_0$ and $h$ are estimated in the assessment model. Annual recruitment is estimated as deviations from the resulting stock-recruitment relationship within the model. MSY reference points are derived from the estimates of $R_0$ and $h$.

In `SCA2`, no stock-recruit relationship is assumed in the assessment model, i.e., annual recruitment is estimated as deviations from the mean recruitment over the observed time series, similar to Cadigan (2016). After the assessment, a stock-recruit function can be fitted post-hoc to the recruitment and spawning stock biomass estimates from the assessment model to obtain MSY reference points.

### Stock-recruit function in assessment model (SCA)

#### Beverton-Holt stock-recruit function

Recruitment $R_t$ in year $t$ is
$$ R_t = \dfrac{\alpha E_{t-1}}{1 + \beta E_{t-1}} \exp(\delta_t - 0.5 \tau^2),$$
where $\delta_t \sim N(0, \tau^2)$ are recruitment deviations from the stock-recruit relationship in lognormal space and $\tau$ is the standard deviation of the recruitment deviations. Parameters $\alpha$ and $\beta$ are defined as
$$ \begin{align}
\alpha &= \dfrac{4hR_0}{(1-h)B_0}\\
\beta &= \dfrac{5h-1}{(1-h)B_0},
\end{align}$$
where $B_0 = R_0 \phi_0$. The biomass per recruit $\phi_0$ is calculated as $\phi_0 = \sum_{a=1}^A m_a w_a l_a$, where
$$
l_a = 
\begin{cases} 
      1 & a = 1\\
      l_{a-1} \exp(-M_{a-1}) & a = 2, \ldots, A-1\\
      \dfrac{l_{a-1} \exp(-M_{a-1})}{1 - \exp(-M_a)} & a = A \\
\end{cases}. 
$$

#### Ricker stock-recruit function

Recruitment $R_t$ in year $t$ is
$$ R_t = \alpha E_{t-1} \exp(-\beta E_{t-1}) \exp(\delta_t - 0.5 \tau^2),$$
where
$$ \begin{align}
\alpha &= \dfrac{(5h)^{1.25} R_0}{B_0}\\
\beta &= \dfrac{5}{4B_0}\log(5h).
\end{align}$$

### No stock-recruit function in assessment model (SCA2)

Recruitment $R_t$ in year $t$ is
$$R_t = \bar{R} \exp(\delta_t - 0.5 \tau^2), $$
where $\delta_t \sim N(0, \tau^2)$ are recruitment deviations in lognormal space from the estimated mean recruitment $\bar{R}$ and $\tau$ is the standard deviation of the recruitment deviations. Typically, $\tau$ is set to 1 so that recruitment is estimated almost as free parameters (Cadigan, 2016).

### Likelihood of recruitment deviations
The log-likelihood of the log-recruitment deviations $\delta_t$ over years $t=1,\ldots...,T$ is
$$L(\delta_t) \propto -T \log\tau - 0.5\sum_{t=1}^{T}\left(\dfrac{\log(\delta_t)}{\tau}\right)^2.$$

# References

Cadigan, N.G. 2016. A state-space stock assessment model for northern cod, including under-reported catches and variable natural mortality rates. Canadian Journal of Fisheries and Aquatic Science 72:296-308.

Carruthers, T, Walters, C.J,, and McAllister, M.K. 2012. Evaluating methods that classify fisheries stock status using only fisheries catch data. Fisheries Research 119-120:66-79.

Fletcher, R. I. 1978. On the restructuring of the Pella-Tomlinson system. Fishery Bulletin 76:515:521.

Forrest, R.E., Martell, S.J.D., Melnychuk, M.C., and Walters, C.J. 2008. An age-structured model with leading management parameters, incorporating age-specific selectivity and maturity. Canadian Journal of Fisheries and Aquatic Science 65:286-296.

Hilborn, R., and Walters, C., 1992. Quantitative Fisheries Stock Assessment: Choice, Dynamics and Uncertainty. Chapman and Hall, New York.

### 2018-08-14 inst/doc/MSEtool.Rmd

@@
 options(width = 650)
+library(MSEtool)
 ```
@@
 
-# Creating MPs from assessments and HCRs
+# Data-rich management procedures (MPs)
 
-The `make_MP` function is a convenient way to stitch together a MP from an assessment model and HCR. The resulting function can then be passed to `DLMtool::runMSE()` for a management procedure.
+## Creating MPs from assessment models and HCRs
+
+MSEtool comes with a set of pre-built data-rich MPs:
+```{r}
+avail("MP", all_avail = FALSE)
+```
+These MPs use an assessment model and a harvest control rule to come up with a management recommendation (TAC).
+
+Additionally, the `make_MP` function is a convenient way to stitch together additional MPs from an assessment model and HCR.
+This is handy if an assessment model with a different configuration than the default is needed or a different harvest control rule is desired.
+The resulting function can then be passed to `DLMtool::runMSE()` as a management procedure.
 
@@
 
-Once it is determined that the MSEtool assessment models are performing adequately (no major convergence issues), then the performance of the MPs that use these models can be evaluated with the DLMtool package.
+Once it is determined that the MSEtool assessment models are performing adequately (no major convergence issues), then the performance of the MPs that use these models can be evaluated within the DLMtool package.
 
@@
 Stochastic TAC recommendations from the Delay-Difference model with two harvest control rules.
-When the 40-10 harvest control rule is applied, the prescribed TAC is close to zero because the estimated terminal year biomass is 0.13 in the assessment model._
+When the 40-10 harvest control rule is applied, the prescribed TAC is close to zero because the assessment model estimated the terminal year biomass depletion to be 0.13 (close to the limit reference point)._
 

### 2018-08-23 R/SCA.R
Re-parameterize dome with better starting values, new argument for lognormal age comps

@@ -10,7 +10,8 @@
 #' @param SR Stock-recruit function (either \code{"BH"} for Beverton-Holt or \code{"Ricker"}).
 #' @param vulnerability Whether estimated vulnerability is \code{"logistic"} or \code{"dome"} (double-normal).
 #' See details for parameterization.
-#' @param CAA_multiplier Numeric for data weighting of catch-at-age matrix. See details.
+#' @param CAA_dist Whether a multinomial or lognormal distribution is used for likelihood of the catch-at-age matrix. See details.
+#' @param CAA_multiplier Numeric for data weighting of catch-at-age matrix if \code{CAA_hist = "multinomial"}. Otherwise ignored. See details.
 #' @param I_type Whether the index surveys population biomass (B; this is the default in the DLMtool operating model),
 #' vulnerable biomass (VB), or spawning stock biomass (SSB).
 #' @param rescale A multiplicative factor that rescales the catch in the assessment model, which
@@ -65,6 +66,9 @@
 #' to that number. If \code{CAA_multiplier <= 1}, then all the annual samples sizes
 #' will be re-scaled by that number. By default, sample sizes are capped at 50.
 #'
+#' Alternatively, a lognormal distribution with inverse proportion variance can be used for the catch at age (Punt and Kennedy, 1994, as
+#' cited by Maunder 2011).
+#'
 #' Vulnerability can be specified to be either logistic or dome. If logistic, then the model
 #' vector \code{vul_par} is of length 2:
 #' \itemize{
@@ -72,15 +76,18 @@
 #' \item \code{vul_par[2]}: the age of 95\% vulnerability (a95) as an offset, i.e., \code{a95 = a50 + exp(vul_par[2])}.
 #' }
 #'
-#' With dome vulnerability, a double normal parameterization is used, where \code{vul_par}
+#' With dome vulnerability, a double Gaussian parameterization is used, where \code{vul_par}
 #' is an estimated vector of length 4:
 #' \itemize{
-#' \item \code{vul_par[1]}: \code{log(sd_asc)}, where sd_asc is the standard deviation of the normal distribution function for the ascending limb
[…]

### 2018-08-23 R/SCA2.R

@@ -4,13 +4,14 @@
 #' @useDynLib MSEtool
 #' @export
 SCA2 <- function(x = 1, Data, SR = c("BH", "Ricker"), vulnerability = c("logistic", "dome"),
-                 CAA_multiplier = 50, I_type = c("B", "VB", "SSB"), rescale = "mean1",
+                 CAA_dist = c("multinomial", "lognormal"), CAA_multiplier = 50, I_type = c("B", "VB", "SSB"), rescale = "mean1",
                  start = NULL, fix_h = FALSE, fix_U_equilibrium = TRUE, fix_sigma = FALSE, fix_tau = TRUE,
                  common_dev = "comp50", integrate = FALSE, silent = TRUE, opt_hess = FALSE, n_restart = ifelse(opt_hess, 0, 1),
                  control = list(iter.max = 5e3, eval.max = 1e4), inner.control = list(),  ...) {
   dependencies <- "Data@Cat, Data@Ind, Data@Mort, Data@L50, Data@L95, Data@CAA, Data@vbK, Data@vbLinf, Data@vbt0, Data@wla, Data@wlb, Data@MaxAge"
   dots <- list(...)
   vulnerability <- match.arg(vulnerability)
+  CAA_dist <- match.arg(CAA_dist)
   SR <- match.arg(SR)
   I_type <- match.arg(I_type)
   if(any(names(dots) == "yind")) {
@@ -41,7 +42,7 @@ SCA2 <- function(x = 1, Data, SR = c("BH", "Ricker"), vulnerability = c("logisti
   t0 <- Data@vbt0[x]
   La <- Linf * (1 - exp(-K * (c(1:max_age) - t0)))
   Wa <- a * La ^ b
-  A50 <- min(0.5 * max_age, iVB(t0, K, Linf, Data@L50[x]))
+  A0 <- min(0.5 * max_age, iVB(t0, K, Linf, Data@L50[x]))
   A95 <- max(A50+0.5, iVB(t0, K, Linf, Data@L95[x]))
   mat_age <- 1/(1 + exp(-log(19) * (c(1:max_age) - A50)/(A95 - A50)))
   LH <- list(LAA = La, WAA = Wa, Linf = Linf, K = K, t0 = t0, a = a, b = b, A50 = A50, A95 = A95)
@@ -51,7 +52,7 @@ SCA2 <- function(x = 1, Data, SR = c("BH", "Ricker"), vulnerability = c("logisti
   if(rescale == "mean1") rescale <- 1/mean(C_hist)
[…]

### 2018-08-23 inst/include/SCA.h

@@ -14,6 +14,7 @@
   DATA_STRING(vul_type);  // String indicating whether logistic or dome vul is used
   DATA_STRING(I_type);    // String whether index surveys B, VB, or SSB
   DATA_STRING(SR_type);   // String indicating whether Beverton-Holt or Ricker stock-recruit is used
+  DATA_STRING(CAA_dist);  // String indicating whether CAA is multinomial or lognormal
   DATA_VECTOR(est_early_rec_dev);
   DATA_VECTOR(est_rec_dev); // Indicator of whether rec_dev is estimated in model or fixed at zero
 
@@ -173,11 +174,16 @@
     if(C_hist(y) > 0) {
       vector<Type> loglike_CAAobs(max_age);
       vector<Type> loglike_CAApred(max_age);
-      for(int a=0;a<max_age;a++) {
-        loglike_CAAobs(a) = (CAA_hist(y,a) + 1e-8) * CAA_n(y);
-        loglike_CAApred(a) = CAApred(y,a)/CN(y);
+      for(int a=0;a<max_age;a++) loglike_CAApred(a) = CAApred(y,a)/CN(y);
+      if(!R_IsNA(asDouble(CAA_n(y)))) {
+        if(CAA_dist == "multinomial") {
+          for(int a=0;a<max_age;a++) loglike_CAAobs(a) = (CAA_hist(y,a) + 1e-8) * CAA_n(y);
+          nll_comp(1) -= dmultinom(loglike_CAAobs, loglike_CAApred, true);
+        } else {
+          for(int a=0;a<max_age;a++) loglike_CAAobs(a) = CAA_hist(y,a);
+          nll_comp(1) -= dlnorm_comp(loglike_CAAobs, loglike_CAApred);
+        }
       }
-      if(!R_IsNA(asDouble(CAA_n(y)))) nll_comp(1) -= dmultinom(loglike_CAAobs, loglike_CAApred, true);
     }
     if(!R_IsNA(asDouble(est_rec_dev(y)))) nll_comp(2) -= dnorm(log_rec_dev(y), Type(0), tau, true);
   }

### 2018-08-23 inst/include/SCA2.h

@@ -13,6 +13,7 @@
   DATA_VECTOR(mat);       // Maturity-at-age at the beginning of the year
   DATA_STRING(vul_type);  // String indicating whether logistic or dome vul is used
   DATA_STRING(I_type);    // String whether index surveys B, VB, or SSB
+  DATA_STRING(CAA_dist);  // String indicating whether CAA is multinomial or lognormal
   DATA_VECTOR(est_early_rec_dev);
   DATA_VECTOR(est_rec_dev); // Indicator of whether rec_dev is estimated in model or fixed at zero
 
@@ -125,13 +126,18 @@
     if(C_hist(y) > 0) {
       vector<Type> loglike_CAAobs(max_age);
       vector<Type> loglike_CAApred(max_age);
-      for(int a=0;a<max_age;a++) {
-        loglike_CAAobs(a) = (CAA_hist(y,a) + 1e-8) * CAA_n(y);
-        loglike_CAApred(a) = CAApred(y,a)/CN(y);
+      for(int a=0;a<max_age;a++) loglike_CAApred(a) = CAApred(y,a)/CN(y);
+      if(!R_IsNA(asDouble(CAA_n(y)))) {
+        if(CAA_dist == "multinomial") {
+          for(int a=0;a<max_age;a++) loglike_CAAobs(a) = (CAA_hist(y,a) + 1e-8) * CAA_n(y);
+          nll_comp(1) -= dmultinom(loglike_CAAobs, loglike_CAApred, true);
+        } else {
+          for(int a=0;a<max_age;a++) loglike_CAAobs(a) = CAA_hist(y,a);
+          nll_comp(1) -= dlnorm_comp(loglike_CAAobs, loglike_CAApred);
+        }
       }
-      if(!R_IsNA(asDouble(CAA_n(y)))) nll_comp(1) -= dmultinom(loglike_CAAobs, loglike_CAApred, true);
     }
-	  if(!R_IsNA(asDouble(est_rec_dev(y)))) nll_comp(2) -= dnorm(log_rec_dev(y), Type(0), tau, true);
+    if(!R_IsNA(asDouble(est_rec_dev(y)))) nll_comp(2) -= dnorm(log_rec_dev(y), Type(0), tau, true);
   }
   for(int a=0;a<max_age-1;a++) {
     if(!R_IsNA(asDouble(est_early_rec_dev(a)))) nll_comp(2) -= dnorm(log_early_rec_dev(a), Type(0), tau, true);

### 2018-08-23 inst/include/functions.h

@@ -130,37 +130,62 @@ Type sum_VBPR(vector<Type> NPR, vector<Type> weight, vector<Type> vul) {
 template<class Type>
 vector<Type> calc_logistic_vul(vector<Type> vul_par, int max_age) {
   vector<Type> vul(max_age);
-  Type vul_50 = vul_par(0);
-  Type vul_95 = vul_50 + exp(vul_par(1));
+  Type a_50 = vul_par(0);
+  Type a_95 = a_50 + exp(vul_par(1));
 
 	for(int a=0;a<max_age;a++) {
 	  Type aa = a;
 	  aa += 1;
-	  vul(a) = pow(1 + exp(-log(Type(19.0)) * (aa - vul_50)/(vul_95 - vul_50)), -1);
+	  vul(a) = pow(1 + exp(-log(Type(19.0)) * (aa - a_50)/(a_95 - a_50)), -1);
   }
 	return vul;
 }
 
 template<class Type>
+Type dnorm_vul(Type x, Type mu, Type sd) {
+  Type res = -0.5;
+  res *= pow(x - mu, 2);
+  res /= pow(sd, 2);
+  return exp(res);
+}
+
+template<class Type>
 vector<Type> calc_dome_vul(vector<Type> vul_par, int max_age) {
   vector<Type> vul(max_age);
-  Type vul_sd_asc = exp(vul_par(0));
-  Type vul_mu_asc = vul_par(1);
-  Type vul_mu_des = vul_mu_asc + exp(vul_par(2));
-  Type vul_sd_des = exp(vul_par(3));
 
-  Type denom_asc = dnorm(vul_mu_asc, vul_mu_asc, vul_sd_asc, false);
-  Type denom_des = dnorm(vul_mu_des, vul_mu_des, vul_sd_des, false);
+  Type a_50 = vul_par(0);
+  Type a_full = a_50 + exp(vul_par(1));
+  Type a_full2 = a_full + exp(vul_par(2));
+  Type vul_max = invlogit(vul_par(3));
+
+  Type var_asc = pow(a_50 - a_full, 2);
+  var_asc /= log(Type(4));
+
+  Type maxage = max_age;
+  Type var_des = pow(maxage - a_full2, 2);
+  var_des /= -2 * log(vul_max);
+
+  Type sd_asc = pow(var_asc, 0.5);
+  Type sd_des = pow(var_des, 0.5);
 
[…]

### 2018-09-05 inst/doc/MSEtool.R
Update vignette for v-1.1

@@ -61,11 +61,9 @@ avail("MP", all_avail = FALSE)
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
 #  DD_MSY <- make_MP(DD_TMB, HCR_MSY, diagnostic = "full")
+#  myMSE <- runMSE(..., MPs = "DD_MSY", PPD = TRUE)
 
 ## ---- eval = FALSE------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------
-#  DD_MSY <- make_MP(DD_TMB, HCR_MSY, diagnostic = "full")
-#  DD_4010 <- make_MP(DD_TMB, HCR40_10)
-#  myMSE <- DLMtool::runMSE(OM = DLMtool::testOM, MPs = c("FMSYref", "DD_MSY", "DD_4010"))
 #  diagnostic_AM(myMSE)
 #  retrospective_AM(myMSE, sim = 1, MP = "DD_MSY")
 

### 2019-02-27 R/MOM_object.R
Only hermaphroditism left - error checked against runMSE

@@ -39,6 +39,8 @@
 #' @slot CatchFrac A list nstock long, of matrices nsim x nfleet representing the fraction of current catches of the various fleets to each stock (each matrix is nsim by nfleet long and rows sum to 1 for each stock)
 #' @slot Allocation  A list nstock long, of vector allocations by fleet (default is NULL and allocation is according to last historical year).
 #' @slot Efactor A list nstock long, of current effort factors by fleet (default is 1 - same as current effort)
+#' @slot Complexes A list of stock complexes. Each position is a vector of stock numbers (as they appear in StockPars) for which data should be aggregated and TAC recommendations split among stocks according to vulnerable biomass
+#' @slot SexPars A list of slots that control sex-specific dynamics
 #' @slot Rel A list of biological / ecological relationships among stocks
 #' over-ridden if an MP of class 'MP_F" is supplied that is a multi-fleet MP.
 #' @author T. Carruthers and A. Hordyk
@@ -51,12 +53,13 @@ setClass("MOM", representation(Name = "character", Agency="character",
                               nsim="numeric", proyears="numeric",
                               interval='numeric', pstar='numeric', maxF='numeric', reps='numeric',
                               cpars="list", seed="numeric", Source="character",Stocks='list',Fleets='list',Obs='list',Imps='list',
-                              CatchFrac='list',Allocation='list',Efactor='list',Rel='list'))
+                              CatchFrac='list',Allocation='list',Efactor='list',Complexes='list',SexPars='list',Rel='list'))
 
 
 # initialize MOM
 setMethod("initialize", "MOM", function(.Object, Stocks=NULL, Fleets=NULL, Obs=NULL, Imps=NULL, CatchFrac=NULL, cpars=NULL,
-                                       interval=4, pstar=0.5, maxF=0.8, reps=1, nsim=48, proyears=50, Source=NULL, Allocation=NULL,Efactor=NULL, Rel=NULL) {
+                                       interval=4, pstar=0.5, maxF=0.8, reps=1, nsim=48, proyears=50, Source=NULL,
+                                       Allocation=NULL,Efactor=NULL, Complexes=NULL, SexPars=NULL, Rel=NULL) {
   # .Object}) ; .Object<-new('MOM')
 
[…]

### 2019-02-27 R/multiMSE.R

@@ -135,9 +135,18 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
   Stocks<-MOM@Stocks
   Fleets<-MOM@Fleets
   Rel<-MOM@Rel
+  SexPars<-MOM@SexPars
+  Complexes<-MOM@Complexes
   CatchFrac<-MOM@CatchFrac
   np<-length(Stocks)
   nf<-length(Fleets[[1]])
+
+  if(np==1&nf==1){
+    message("You have specified only a single stock and fleet. You should really be using the function runMSE()")
+  }else if(np>1 & length(MOM@Rel)==0 & length(MOM@SexPars)==0){
+    message("You have specified more than one stock but no MICE relationships (slot MOM@Rel) or sex-specific relationships (slot MOM@SexPars) among these. As they are independent, consider doing MSE for one stock at a time for computational efficiency.")
+  }
+
   maxF<-MOM@maxF
   Snames<-SIL(Stocks,"Name")
   Fnames<-matrix(make.unique(SIL(Fleets,"Name")),nrow=nf)
@@ -337,6 +346,27 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
 
   } # end of np
 
+
+  # if SexPars Sex specific exceptions (recalculation of SSB0, aR, bR, SSBpR)
+  if(length(SexPars)>0){
+
+    SSB0s<-matrix(NIL(StockPars,"SSB0"),nrow=nsim) # sim, p
+
+    for(p in 1:np){
+
+      SSB0<-apply(matrix(rep(SexPars$SSBfrom[p,],each=nsim),nrow=nsim)*SSB0s,1,sum)
+      StockPars[[p]]$SSBpR<-array(SSB0/StockPars[[p]]$R0,c(nsim,nareas)) # !!!!!!!!!!! SSBpR hardwired to be the same among areas !!!!
+
+      idist<-StockPars[[p]]$R0a/apply(StockPars[[p]]$R0a,1,sum)
+      SSB0a<-SSB0*idist
+
+      StockPars[[p]]$bR <- matrix(log(5 * StockPars[[p]]$hs)/(0.8 * SSB0a), nrow=nsim)  # Ricker SR params
[…]

### 2019-02-27 R/multiMSE_internal.R

@@ -22,9 +22,10 @@
 #' @param CatchFrac A list of stock-specific fleet fractions of current catch list[[stock]][nsim, nf]
 #' @param bounds Bounds for total q estimation
 #' @param Rel A list of inter-stock relationships see slot Rel of MOM object class
+#' @param SexPars A list of sex-specific relationships (SSBfrom, stock_age)
 #' @author T.Carruthers
 #' @keywords internal
-HistMICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears, N, VF, FretA, maxF=0.9, MPA,Rel,qs,qfrac){
+HistMICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears, N, VF, FretA, maxF=0.9, MPA,Rel,SexPars,qs,qfrac){
 
   Nx<-array(N[x,,,,],dim(N)[2:5])
   VFx<-array(VF[x,,,,],dim(VF)[2:5])
@@ -44,13 +45,13 @@ HistMICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears, N, VF,
 
   distx<-Asizex<-SSBpRx<-R0ax<-aRx<-bRx<-array(NA,c(np,nareas))
   Perrx<-array(NA,c(np,nyears+maxage))
-  movx<-array(NA,c(np,maxage,nareas,nareas))
+  movx<-array(NA,c(np,maxage,nareas,nareas,nyears))
 
   for(p in 1:np){
     distx[p,]<-StockPars[[p]]$R0a[x,]/sum(StockPars[[p]]$R0a[x,])
     Perrx[p,]<-StockPars[[p]]$Perr_y[x,1:(nyears+maxage)]
     Asizex[p,]<-StockPars[[p]]$Asize[x,]
-    movx[p,,,]<-StockPars[[p]]$mov[x,,,]
+    movx[p,,,,]<-StockPars[[p]]$mov[x,,,,1:nyears]
     SSBpRx[p,]<-StockPars[[p]]$SSBpR[x,]
     R0ax[p,]<-StockPars[[p]]$R0a[x,]
     aRx[p,]<-StockPars[[p]]$aR[x,]
[…]

### 2019-02-27 R/multi_q_estimation.R

@@ -22,10 +22,11 @@
 #' @param bounds Bounds for total q estimation
 #' @param tol A numeric value that is the fraction of machine tolerance (once reduction in objective function steps below this, optimization ends)
 #' @param Rel A list of inter-stock relationships see slot Rel of MOM object class
+#' @param SexPars A list of sex-specific dynamics SSBfrom stock_age
 #' @author T.Carruthers
 #' @keywords internal
 #' @export
-getq_multi_MICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears, N, VF, FretA, maxF=0.9, MPA,CatchFrac, bounds= c(1e-05, 15),tol=1E-3,Rel){
+getq_multi_MICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears, N, VF, FretA, maxF=0.9, MPA,CatchFrac, bounds= c(1e-05, 15),tol=1E-6,Rel,SexPars){
 
   Nx<-array(N[x,,,,],dim(N)[2:5])
   VFx<-array(VF[x,,,,],dim(VF)[2:5])
@@ -46,12 +47,12 @@ getq_multi_MICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears,
 
   distx<-Asizex<-SSBpRx<-R0ax<-aRx<-bRx<-array(NA,c(np,nareas))
   Perrx<-array(NA,c(np,nyears+maxage))
-  movx<-array(NA,c(np,maxage,nareas,nareas))
+  movx<-array(NA,c(np,maxage,nareas,nareas,nyears))
 
   for(p in 1:np){
     distx[p,]<-StockPars[[p]]$R0a[x,]/sum(StockPars[[p]]$R0a[x,])
     Perrx[p,]<-StockPars[[p]]$Perr_y[x,1:(nyears+maxage)]
-    movx[p,,,]<-StockPars[[p]]$mov[x,,,]
+    movx[p,,,,]<-StockPars[[p]]$mov[x,,,,1:nyears]
     SSBpRx[p,]<-StockPars[[p]]$SSBpR[x,]
     R0ax[p,]<-StockPars[[p]]$R0a[x,]
[…]

### 2019-02-27 R/popdyn_MICE.R

@@ -33,10 +33,13 @@
 #' @param Perrx Matrix [stock, year] process error - the lognormal factor for recruitment strength
 #' @param SRrelx Integer vector [stock] the form of the stock recruitment relationship (1 = Beverton-Holt, 2= Ricker)
 #' @param Rel A list of inter-stock relationships see slot Rel of MOM object class
+#' @param SexPars A list of sex-specific relationships (SSBfrom, stock_age)
 #' @author T.Carruthers
 #' @keywords internal
 #' @export
-popdynMICE<-function(qsx,qfracx,np,nf,nyears,nareas,maxage,Nx,VFx,FretAx,Effind,movx,Spat_targ,M_ageArrayx,Mat_agex,Asizex,Kx,Linfx,t0x,Mx,R0x,R0ax,SSBpRx,hsx,aRx, bRx,ax,bx,Perrx,SRrelx,Rel){
+popdynMICE<-function(qsx,qfracx,np,nf,nyears,nareas,maxage,Nx,VFx,FretAx,Effind,movx,Spat_targ,
+                     M_ageArrayx,Mat_agex,Asizex,Kx,Linfx,t0x,Mx,R0x,R0ax,SSBpRx,hsx,aRx,
+                     bRx,ax,bx,Perrx,SRrelx,Rel, SexPars){
 
   Bx<-SSNx<-SSBx<-VBx<-Zx<-array(NA,dim(Nx))
   VBfx<-array(NA,c(np,nf,maxage,nyears,nareas)) # initial year calculation
@@ -90,12 +93,12 @@ popdynMICE<-function(qsx,qfracx,np,nf,nyears,nareas,maxage,Nx,VFx,FretAx,Effind,
                        FMretx=FMretx,
                        FMx=FMx,
                        PerrYrp=Perrx[,y+maxage-2], hsx=hsy[,y-1], aRx=aRx, bRx=bRx,
-                       movx=movx, Spat_targ=Spat_targ, SRrelx=SRrelx,
+                       movy=array(movx[,,,,y-1],c(np,maxage,nareas,nareas)), Spat_targ=Spat_targ, SRrelx=SRrelx,
                        M_agecur=array(M_ageArrayx[,,y-1],dim(M_ageArrayx)[1:2]),
[…]

### 2019-03-01 R/MMSE_Object.R
Documenting / testing multiMSE

@@ -48,8 +48,10 @@
 #' @slot CAA Catch at age in last projection year. An array with dimensions: nsim, nMPs, nages. Non-negative real numbers
 #' @slot CAL Catch at length in last projection year. An array with dimensions: nsim, nMPs, nCALbins. Non-negative real numbers
 #' @slot CALbins Mid-points of the catch-at-length bins. Vector of length nCALbins. Positive real numbers.
+#' @slot MSY_P Array of projected MSY by year with dimensions: nsim, nstock, nMP, proyears.
+#' @slot FMSY_P Array of projected FMSY by year with dimensions: nsim, nstock, nMP, proyears.
+#' @slot SSBMSY_P Array of projected Spawning Stock Biomass at MSY by year with dimensions: nsim, nstock, nMP, proyears.
 #' @slot Misc Miscellanenous output such as posterior predictive data
-#'
 #' @author T. Carruthers
 #' @keywords classes
 setClass("MMSE", representation(Name = "character", nyears = "numeric",
@@ -60,14 +62,15 @@ setClass("MMSE", representation(Name = "character", nyears = "numeric",
                                B = "array", SSB="array", VB="array", FM = "array", C = "array",
                                TAC = "array", SSB_hist = "array",
                                CB_hist = "array", FM_hist = "array", Effort = "array", PAA= "array", CAA= "array",
-                               CAL= "list", CALbins="list", Misc="list"))
+                               CAL= "list", CALbins="list", MSY_P="array", FMSY_P="array", SSBMSY_P="array", Misc="list"))
 
 
 setMethod("initialize", "MMSE", function(.Object, Name, nyears, proyears,
                                         nMPs, MPs, MPcond, MPrefs,
                                         nsim, nstocks, nfleets, Snames, Fnames, Stocks, Fleets, Obss, Imps,
                                         OM, Obs, B_BMSY, F_FMSY, B, SSB, VB, FM, C, TAC,
-                                        SSB_hist, CB_hist, FM_hist, Effort = array(), PAA,  CAA, CAL, CALbins, Misc) {
+                                        SSB_hist, CB_hist, FM_hist, Effort = array(), PAA,  CAA, CAL, CALbins,
+                                        MSY_P, FMSY_P, SSBMSY_P, Misc) {
   .Object@Name <- Name
   .Object@nyears <- nyears
   .Object@proyears <- proyears

### 2019-03-01 R/MMSE_plots.R

@@ -1,0 +1,129 @@
+
+plotquant<-function(x,p=c(0.05,0.25,0.75,0.95),yrs,qcol,lcol,addline=T,ablines=NA){
+  ny<-length(yrs)
+  qs<-apply(x,2,quantile,p=p[c(1,4)])
+  qsi<-apply(x,2,quantile,p=p[2:3])
+  polygon(c(yrs,yrs[ny:1]),c(qs[1,],qs[2,ny:1]),border=NA,col='#b3ecff')
+  
+  polygon(c(yrs,yrs[ny:1]),c(qsi[1,],qsi[2,ny:1]),border=NA,col=qcol)
+  if(!is.na(ablines[1]))abline(h=ablines,col='#99999980')
+  
+  if(addline)for(i in 1:2)lines(yrs,x[i,],col=lcol,lty=i)
+  lines(yrs,apply(x,2,quantile,p=0.5),lwd=2,col="white")
+}
+
+Pplot3<-function(MSEobj,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",curyr=2018,quants=c(0.05,0.25,0.75,0.95),MPcols='black',addline=F,maxrow=NA){
+  
+  if(is.na(maxcol))maxcol=ceiling(length(MSEobj@MPs)/0.5) # defaults to portrait 1:2
+  MPs<-MSEobj@MPs
+  nMPs<-length(MPs)
+  yrs<-curyr+(1:MSEobj@proyears)
+  
+  MPcols[MPcols=="green"]<-"darkgreen"
+  
+  plots<-split(1:nMPs, ceiling(seq_along(1:nMPs)/maxcol))
+  
+  nr<-length(plots)*2
+  if(!is.na(maxrow))nr=max(nr,maxrow)
+  nc<-maxcol
+  
+  mat<-array(0,c(nc,nr*1.5))
+  ind<-floor(0.5+(1:nr)*1.5)
+  mat[,ind]<-1:(nr*nc)
+  mat<-t(mat)
+  ht<-rep(0.2,nr*1.5)
+  ht[ind]<-1
+  layout(mat,heights=ht)
+  par(mai=c(0.3,0.3,0.01,0.01),omi=c(0.5,0.5,0.05,0.05))
+  
+  B_BMSY<-MSEobj@B_BMSY
[…]

### 2019-03-01 R/multiMSE.R

@@ -375,7 +375,7 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
   #  qs <- array(0,dim(nsim,nf)) # no fishing
   #} else {
 
-  if(!silent) message("Optimizing for user-specified depletion (takes approximately nstocks/3 x nfleets/3 minutes..)")  # Print a progress update
+  if(!silent) message("Optimizing for user-specified depletion (takes approximately [(nstocks x nfleets)/(9 x number of cores in cluster)] minutes per simulation)")  # Print a progress update
 
   bounds <- c(0.0001, 15) # q bounds for optimizer
 
@@ -1072,26 +1072,10 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
   MSY_P <- FMSY_P <- SSBMSY_P <- array(NA, dim=c(nsim,np, nMP, proyears))
 
   for(p in 1:np){
-    MSY_P <- array(StockPars[[p]]$MSY, dim=c(nsim,np, nMP, proyears))
-    FMSY_P <- array(StockPars[[p]]$FMSY, dim=c(nsim,np, nMP, proyears))
-    SSBMSY_P <- array(StockPars[[p]]$SSBMSY, dim=c(nsim,np, nMP, proyears))
+    MSY_P[,p,,] <- StockPars[[p]]$MSY
+    FMSY_P[,p,,] <- StockPars[[p]]$FMSY
+    SSBMSY_P[,p,,] <- StockPars[[p]]$SSBMSY
   }
-
-
-  #if (AnnualMSY) {
-
-   # if(!silent) message("Calculating MSY reference points for each projection year")
-    #for (y in 1:proyears) {
-     # if(!silent) cat('.')
-      #if (!silent) flush.console()
-      #MSYrefsYr <- sapply(1:nsim, optMSY_eq, M_ageArray, Wt_age, Mat_age, V, maxage, R0, SRrel, hs, yr=nyears+y)
-      #MSY_P[,,y] <- MSYrefsYr[1, ]
-      #FMSY_P[,,y] <- MSYrefsYr[2,]
-      #SSBMSY_P[,,y] <- MSYrefsYr[3,]
-
-    #}
-    #if(!silent) cat("\n")
-  #}
 
   interval<-MOM@interval
[…]

### 2019-03-01 R/MMSE_plots.R
MPs for complexes added and some plotting for MOM and MMSE objects

@@ -4,29 +4,29 @@ plotquant<-function(x,p=c(0.05,0.25,0.75,0.95),yrs,qcol,lcol,addline=T,ablines=N
   qs<-apply(x,2,quantile,p=p[c(1,4)])
   qsi<-apply(x,2,quantile,p=p[2:3])
   polygon(c(yrs,yrs[ny:1]),c(qs[1,],qs[2,ny:1]),border=NA,col='#b3ecff')
-  
+
   polygon(c(yrs,yrs[ny:1]),c(qsi[1,],qsi[2,ny:1]),border=NA,col=qcol)
   if(!is.na(ablines[1]))abline(h=ablines,col='#99999980')
-  
+
   if(addline)for(i in 1:2)lines(yrs,x[i,],col=lcol,lty=i)
   lines(yrs,apply(x,2,quantile,p=0.5),lwd=2,col="white")
 }
 
 Pplot3<-function(MSEobj,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",curyr=2018,quants=c(0.05,0.25,0.75,0.95),MPcols='black',addline=F,maxrow=NA){
-  
+
   if(is.na(maxcol))maxcol=ceiling(length(MSEobj@MPs)/0.5) # defaults to portrait 1:2
   MPs<-MSEobj@MPs
   nMPs<-length(MPs)
   yrs<-curyr+(1:MSEobj@proyears)
-  
+
   MPcols[MPcols=="green"]<-"darkgreen"
-  
+
   plots<-split(1:nMPs, ceiling(seq_along(1:nMPs)/maxcol))
-  
+
   nr<-length(plots)*2
   if(!is.na(maxrow))nr=max(nr,maxrow)
   nc<-maxcol
-  
+
   mat<-array(0,c(nc,nr*1.5))
   ind<-floor(0.5+(1:nr)*1.5)
   mat[,ind]<-1:(nr*nc)
@@ -35,19 +35,19 @@ Pplot3<-function(MSEobj,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",cur
   ht[ind]<-1
   layout(mat,heights=ht)
   par(mai=c(0.3,0.3,0.01,0.01),omi=c(0.5,0.5,0.05,0.05))
-  
+
   B_BMSY<-MSEobj@B_BMSY
[…]

### 2019-03-01 R/MOM_plots.R

@@ -1,0 +1,275 @@
+# =======================================================================
+# === MOM plotting functions ============================================
+# =======================================================================
+
+
+
+#' Standard plot for an object of class MOM
+#'
+#' Plot the stocks, fleets, catch fractions and relationships in multi operating model object
+#'
+#' @param MOM Object of class MOM. A Multi-OM object created by new('MOM', ...)
+#' @param silent Logical. Do you wish to see print outs / warnings?
+#' @param maxsims Integer. What are the maximum number of individual simulations you wish to plot?
+#' @param imp Logical. Whether implementation error should be removed
+#' @param proc Logical. Whether process error should be removed (e.g. interannual variability in natural mortality rate)
+#' @param grad Logical. Whether gradients (consistent temporal changes in parameters) should be removed
+#' @param silent Logical. Should changes be printed to the console?
+#' @author T.Carruthers
+#' @export
+plot.MOM<-function(MOM,silent=T,maxsims=6){
+
+  #MOM<-match.arg(MOM)
+  #if(is.null(MOM))MOM=MOM
+
+  Rel<-MOM@Rel
+  Stocks<-MOM@Stocks
+  Fleets<-MOM@Fleets
+  Obs<-MOM@Obs
+  Imps<-MOM@Imps
+  cpars<-MOM@cpars
+  nsim<-MOM@nsim
+  maxsims<-min(MOM@nsim,maxsims)
+  nsim<-200 # for plotting purposes
+  nyears <- MOM@Fleets[[1]][[1]]@nyears  # number of historical years
+  proyears<-MOM@proyears
+  allyears<-nyears+proyears
+
+  np<-length(MOM@Stocks)
+  nf<-length(Fleets[[1]])
+  Snams<-SIL(MOM@Stocks,"Name")
+  Fnames<-t(matrix(SIL(MOM@Fleets,"Name"),nrow=nf))
+
+
+  # --- Plot MICE Rels --------------------------------------------------------------------------------------------
+
+  if(length(Rel)>0){
+
[…]

### 2019-03-01 R/multiMSE.R

@@ -138,6 +138,7 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
   SexPars<-MOM@SexPars
   Complexes<-MOM@Complexes
   CatchFrac<-MOM@CatchFrac
+
   np<-length(Stocks)
   nf<-length(Fleets[[1]])
 
@@ -270,6 +271,7 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
     Vraw<-array(NIL(listy=FleetPars[[p]],namey="V"),c(nsim,maxage,allyears,nf))
     Vind<-as.matrix(expand.grid(1:nsim,p,1:nf,1:maxage,1:allyears))
     VF[Vind]<-Vraw[Vind[,c(1,4,5,3)]]
+
     if(nf==1){
       V<-VF[,p,1,,] #<-SOL(FleetPars[[p]],"V")
     }else{
@@ -281,6 +283,7 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
       #V<-nlz(V,c(1,3),"max") # currently assume unfished vulnerability is equally weighted among fleets
       # V includes discards
     }
+
     VBiomass[SPAYR] <- Biomass[SPAYR] * V[SAY]  # Calculate vunerable biomass
 
     Fretraw<-array(NIL(listy=FleetPars[[p]],namey="retA"),c(nsim,maxage,allyears,nf))
@@ -586,7 +589,7 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
   # Check that depletion is correct
   if (checks) {
     if (prod(round(Depletion,2)/ round(D_specified,2)) != 1) {
-      print(cbind(round(Depletion,4), round(D_specified,4)))
+      print(cbind(round(Depletion,4),rep(NaN,nsim), round(D_specified,4)))
       warning("Possible problem in depletion calculations")
     }
   }
@@ -991,8 +994,8 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
 
   # Detecting MP specification  -----------------------------------------------------------------------------------------------------------------------------------
 
-  if(identical(ldim(MPs),MSEtool:::ldim(Fleets))){
[…]

### 2019-03-01 R/multiMSE_internal.R

@@ -250,24 +250,32 @@ multiData<-function(MSElist,StockPars,p,mm,nf){
 #' @param np The number of stocks
 #' @param mm Integer the MP number
 #' @param nf The number of fleets
+#' @param realVB A matrix of real vulnerable biomass [nsim,year,np]
 #' @author T. Carruthers
 #' @export
-multiDataS<-function(MSElist,StockPars,np,mm,nf){
+multiDataS<-function(MSElist,StockPars,np,mm,nf,realVB){
+
+  nsim<-dim(MSElist[[1]][[1]][[1]]@Cat)[1]
+  nyears<-dim(MSElist[[1]][[1]][[1]]@Cat)[2]
+  na<-dim(MSElist[[1]][[1]][[1]]@CAA)[3]
+  nl<-dim(MSElist[[1]][[1]][[1]]@CAL)[3]
+  ni<-np*nf
+
+  if(realVB[1,1,1]==0)realVB[,,1]<-realVB[,,2] # impute vulnerable biomass for year 1 if missing (a negligible issue to be fixed in popdyn_MICE)
 
   DBF<-list()
+  realVBi<-array(NA,c(nsim,nyears,nf*np))
+
   for(f in 1:nf){
     for(p in 1:np){
       i<-p+(np*(f-1))
       DBF[[i]]<-MSElist[[p]][[f]][[mm]]
+      realVBi[,,i]<-realVB[,,p]
     }
   }
-  ni<-np*nf
+
 
   Dataout<-DBF[[1]]
-  nsim<-dim(Dataout@Cat)[1]
-  nyears<-dim(Dataout@Cat)[2]
-  na<-dim(Dataout@CAA)[3]
-  nl<-dim(Dataout@CAL)[3]
 
   Cat<-array(SIL(DBF,"Cat"),c(nsim,nyears,ni))
   CAA<-array(SIL(DBF,"CAA"),c(nsim,nyears,na,ni))
@@ -288,48 +296,59 @@ multiDataS<-function(MSElist,StockPars,np,mm,nf){
   Dataout@LFS<-apply(LFS*Cat[,nyears,],1,sum)/apply(Cat[,nyears,],1,sum)
   Dataout@LFC<-apply(LFC*Cat[,nyears,],1,sum)/apply(Cat[,nyears,],1,sum)
 
[…]

### 2019-03-05 R/MMSE_plots.R
Added plot.MMSE plot.MOM and applyMMP
Only thing left is to add SexPars$Herm to the forward projection in popdynMICE

@@ -1,4 +1,14 @@
-
+#' A fairly tidy time-series quantile plot
+#'
+#' @param x Matrix. A time series quantity [simulation, year]
+#' @param p Numeric vector. The percentiles that are plotted (LB2,LB1,UB1,UB2). LB2 and UB2 are the outer percentiles, LB1 and UB1 are the inner percentiles.
+#' @param yrs Numeric vector. The years corresponding to the indexing of x
+#' @param qcol Character, color. The color of the inner percentile range
+#' @param lcol Character, color. The color of the outer percentile range.
+#' @param addline Logical. Should two individual simulations be added to the percentile plots?
+#' @param ablines Numeric vector. Horizontal lines to be added to the plot.
+#' @author T.Carruthers
+#' @export
 plotquant<-function(x,p=c(0.05,0.25,0.75,0.95),yrs,qcol,lcol,addline=T,ablines=NA){
   ny<-length(yrs)
   qs<-apply(x,2,quantile,p=p[c(1,4)])
@@ -12,123 +22,160 @@ plotquant<-function(x,p=c(0.05,0.25,0.75,0.95),yrs,qcol,lcol,addline=T,ablines=N
   lines(yrs,apply(x,2,quantile,p=0.5),lwd=2,col="white")
 }
 
-Pplot3<-function(MSEobj,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",curyr=2018,quants=c(0.05,0.25,0.75,0.95),MPcols='black',addline=F,maxrow=NA){
-
-  if(is.na(maxcol))maxcol=ceiling(length(MSEobj@MPs)/0.5) # defaults to portrait 1:2
-  MPs<-MSEobj@MPs
-  nMPs<-length(MPs)
-  yrs<-curyr+(1:MSEobj@proyears)
-
-  MPcols[MPcols=="green"]<-"darkgreen"
-
-  plots<-split(1:nMPs, ceiling(seq_along(1:nMPs)/maxcol))
-
-  nr<-length(plots)*2
-  if(!is.na(maxrow))nr=max(nr,maxrow)
-  nc<-maxcol
-
-  mat<-array(0,c(nc,nr*1.5))
-  ind<-floor(0.5+(1:nr)*1.5)
-  mat[,ind]<-1:(nr*nc)
[…]

### 2019-03-05 R/MOM_plots.R

@@ -19,9 +19,6 @@
 #' @export
 plot.MOM<-function(MOM,silent=T,maxsims=6){
 
-  #MOM<-match.arg(MOM)
-  #if(is.null(MOM))MOM=MOM
-
   Rel<-MOM@Rel
   Stocks<-MOM@Stocks
   Fleets<-MOM@Fleets

### 2019-03-05 R/applyMMP.R

@@ -1,0 +1,120 @@
+
+
+#' Get part of an MP specific data-list
+#'
+#' @param MSElist A hierarchical list [Stock][Fleet][MP]
+#' @param mm integer the MP number
+#'
+#' @return a sublist of MSElist for a specific MP
+#' @export
+#'
+getDataList<-function(MSElist,mm){
+  DataList<-new('list')
+  for(ss in 1:length(MSElist)){
+    DataList[[ss]]<-new('list')
+    for(ff in 1:length(MSElist[[ss]]))DataList[[ss]][[ff]]<-MSElist[[ss]][[ff]][[mm]]
+  }
+  DataList
+}
+
+
+#' Apply multi Management Procedures (class MMP) to a hierarchical list of Data class objects
+#'
+#' @param Data List, a hierarchical list of Data objects (Fleets nested in Stocks)
+#' @param MPstorun Name(s) of the MPs to run
+#' @param reps Number of samples
+#' @param nsims Optional. Number of simulations.
+#' @param silent Logical. Should messages be suppressed?
+#'
+#' @return A hierarchical list of management recommendations (object class Rec), Fleets nested in Stocks
+#' @export
+#'
+applyMMP <- function(DataList, MP = NA, reps = 1, nsims=NA, silent=FALSE) {
+
+  if (is.na(nsims)) nsims <- length(DataList[[1]][[1]]@Mort)
+  nMPs <- length(MPs)
+  if (.hasSlot(DataList[[1]][[1]], "nareas")) {
+    nareas <- DataList[[1]][[1]]@nareas
+  } else {
+    nareas <- 2
+  }
+
+  DataList_F <-DataList  # formatted data list
+
+  for(ss in 1:length(DataList)){
+    for(ff in 1:length(DataList[[ss]])){
+      DataList_F[[ss]][[ff]] <- updateMSE(DataList[[ss]][[ff]])
+    }
+  }
+
+  temp <- lapply(1:nsims, MP, DataList = DataList, reps = reps)
+  recList<-
+
+  #if (!silent && any(apply(is.na(recList$TAC), 2, sum) > rep(0.5 * reps, nsims)))
+  #  message("Method ", MPs[mp], " produced greater than 50% NA values")
+
+  #Data@MPs <- MPs
+  CombineMMP(temp,nareas)
+
+}
+
+
+
+#' Create a blank MP recommendations object (class Rec) of the right dimensions
+#'
+#' @param nsim Integer. The number of stocks
[…]

### 2019-03-05 R/multiMSE.R

@@ -77,7 +77,7 @@ multiMSE <- function(MOM, MPs = list(c("AvC","DCAC"),c("FMSYref","curE")),
   if (!all(is.na(MPvec))) {
     for (mm in MPvec) {
       chkMP <- try(get(mm), silent=TRUE)
-      if (class(chkMP) != 'MP') stop(mm, " is not a valid MP", call.=FALSE)
+      if (!(class(chkMP) %in% c('MP','MMP'))) stop(mm, " is not a valid MP", call.=FALSE)
     }
   }
 
@@ -240,16 +240,18 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
   MPA<-array(1,c(np,nf, nyears+proyears,nareas))
   Agearray <- array(rep(1:maxage, each = nsim), dim = c(nsim, maxage))  # Age array
 
-  # Survival array with M-at-age
+  # Hermaphroditism
+  HermFrac<-expandHerm(SexPars$Herm,maxage=StockPars[[1]]$maxage,np=np,nsim=nsim) # [sim, stock, maxage] Defaults to all 1s if length(SexPars$Herm)==0
 
   for(p in 1:np){
 
     surv <- matrix(1, nsim, maxage)
     surv[, 2:maxage] <- t(exp(-apply(StockPars[[p]]$M_ageArray[,,1], 1, cumsum)))[, 1:(maxage-1)]  # Survival array
 
-    Nfrac <- surv * StockPars[[p]]$Mat_age[,,1]  # predicted Numbers of mature ages in first year
+    Nfrac <- surv * StockPars[[p]]$Mat_age[,,1] * HermFrac[,p,]  # predicted Numbers of mature ages in first year
 
     SPAYR <- as.matrix(expand.grid(1:nareas, 1, 1:maxage, p, 1:nsim)[5:1])  # Set up some array indexes sim (S) age (A) year (Y) region/area (R)
+    SPA<-SPAYR[,1:3]
     SAY <- SPAYR[, c(1,3,4)]
     SAR <- SPAYR[, c(1,3,5)]
     SA <- Sa <- SPAYR[, c(1,3)]
@@ -263,7 +265,7 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
[…]

### 2019-03-05 R/multiMSE_internal.R

@@ -354,4 +354,23 @@ multiDataS<-function(MSElist,StockPars,np,mm,nf,realVB){
 }
 
 
+#' Expand the Herm list in SexPars to a matrix of fractions at age
+#'
+#' @param Herm A list of Hermaphroditic fractions at age (starting age class 1)
+#' @param maxage The maximum age of stocks being simulated
+#' @param np The total number of stocks being simulated
+#' @author T. Carruthers
+#' @export
+expandHerm<-function(Herm,maxage,np,nsim){
+  HermFrac<-array(1,c(nsim,np,maxage))
+  if(length(Herm)>0){
+    ps<-matrix(as.numeric(sapply(names(Herm),function(x)strsplit(x,"_")[[1]][2:3])),nrow=length(Herm),byrow=T)
+    for(i in 1:length(Herm)){
+      HermFrac[,ps[i,1],1:ncol(Herm[[1]])]<-Herm[[1]]
+      HermFrac[,ps[i,2],]<-0
+      HermFrac[,ps[i,2],1:ncol(Herm[[1]])]<-1-Herm[[1]]
+    }
+  }
+  HermFrac
+}
 

### 2019-03-06 R/multiMSE.R
Adding Hermaphroditism exception to popdynMICE

@@ -240,7 +240,7 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
   MPA<-array(1,c(np,nf, nyears+proyears,nareas))
   Agearray <- array(rep(1:maxage, each = nsim), dim = c(nsim, maxage))  # Age array
 
-  # Hermaphroditism
+  # Hermaphroditism (this is the fraction to be kept (after sex change)) E.g. protygynous (Female to male) is H_1_2 where 1 is female 2 is male
   HermFrac<-expandHerm(SexPars$Herm,maxage=StockPars[[1]]$maxage,np=np,nsim=nsim) # [sim, stock, maxage] Defaults to all 1s if length(SexPars$Herm)==0
 
   for(p in 1:np){
@@ -262,7 +262,7 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
 
     # <<<< Previously an if statement for existence of initdist, see runMSE.R of DLMtool/R>>>>>>>>
 
-    R0a <- matrix(StockPars[[p]]$R0, nrow=nsim, ncol=nareas, byrow=FALSE) * StockPars[[p]]$initdist[,1,]  # !!!! INITDIST OF AGE 1. Unfished recruitment by area
+    R0a <- matrix(StockPars[[p]]$R0, nrow=nsim, ncol=nareas, byrow=FALSE) * StockPars[[p]]$initdist[,1,] #*HermFrac[,p,1]  # !!!! INITDIST OF AGE 1. Unfished recruitment by area
 
     SSN[SPAYR] <- Nfrac[SA] * StockPars[[p]]$R0[S] * StockPars[[p]]$initdist[SAR]  # Calculate initial spawning stock numbers
     N[SPAYR] <- StockPars[[p]]$R0[S] * surv[SA] *HermFrac[SPA] * StockPars[[p]]$initdist[SAR]  # Calculate initial stock numbers
@@ -351,12 +351,13 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
 
   } # end of np
 
-  # GOT TO HERE IN YOUR CHECKING
 
   # if SexPars Sex specific exceptions (recalculation of SSB0, aR, bR, SSBpR)
[…]

### 2019-03-06 R/multi_q_estimation.R

@@ -91,13 +91,13 @@ getq_multi_MICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears,
              depc=depc,CFc=CFc,mode='opt',np=np,nf=nf,nyears=nyears,nareas=nareas,maxage=maxage,Nx=Nx,VFx=VFx,FretAx=FretAx,
              Effind=Effind,distx=distx,movx=movx,Spat_targ=Spat_targ,M_ageArrayx=M_ageArrayx,Mat_agex=Mat_agex,Asizex=Asizex,Kx=Kx,
              Linfx=Linfx,t0x=t0x,Mx=Mx,R0x=R0x,R0ax=R0ax,SSBpRx=SSBpRx,SSB0x=SSB0x,hsx=hsx,ax=ax,bx=bx,aRx=aRx,bRx=bRx,Perrx=Perrx,SRrelx=SRrelx,
-             Rel=Rel,SexPars=SexPars,
+             Rel=Rel,SexPars=SexPars,x=x,
              control=list(trace=1,factr=factr))
 
   out<-qestMICE(par=opt$par, depc=depc,CFc=CFc,mode='calc',np=np,nf=nf,nyears=nyears,nareas=nareas,maxage=maxage,Nx=Nx,VFx=VFx,FretAx=FretAx,
           Effind=Effind,distx=distx,movx=movx,Spat_targ=Spat_targ,M_ageArrayx=M_ageArrayx,Mat_agex=Mat_agex,Asizex=Asizex,Kx=Kx,
           Linfx=Linfx,t0x=t0x,Mx=Mx,R0x=R0x,R0ax=R0ax,SSBpRx=SSBpRx,SSB0x=SSB0x,hsx=hsx,aRx=aRx,bRx=bRx,ax=ax,bx=bx,Perrx=Perrx,SRrelx=SRrelx,
-          Rel=Rel,SexPars=SexPars)
+          Rel=Rel,SexPars=SexPars,x=x)
 
   return(out)
 
@@ -143,11 +143,12 @@ getq_multi_MICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears,
 #' @param SRrelx Integer vector [stock] the form of the stock recruitment relationship (1 = Beverton-Holt, 2= Ricker)
 #' @param Rel A list of inter-stock relationships see slot Rel of MOM object class
[…]

### 2019-03-06 R/popdyn_MICE.R

@@ -34,12 +34,13 @@
 #' @param SRrelx Integer vector [stock] the form of the stock recruitment relationship (1 = Beverton-Holt, 2= Ricker)
 #' @param Rel A list of inter-stock relationships see slot Rel of MOM object class
 #' @param SexPars A list of sex-specific relationships (SSBfrom, stock_age)
+#' @param x Integer the simulation number
 #' @author T.Carruthers
 #' @keywords internal
 #' @export
 popdynMICE<-function(qsx,qfracx,np,nf,nyears,nareas,maxage,Nx,VFx,FretAx,Effind,movx,Spat_targ,
                      M_ageArrayx,Mat_agex,Asizex,Kx,Linfx,t0x,Mx,R0x,R0ax,SSBpRx,hsx,aRx,
-                     bRx,ax,bx,Perrx,SRrelx,Rel, SexPars){
+                     bRx,ax,bx,Perrx,SRrelx,Rel,SexPars,x){
 
   Bx<-SSNx<-SSBx<-VBx<-Zx<-array(NA,dim(Nx))
   VBfx<-array(NA,c(np,nf,maxage,nyears,nareas)) # initial year calculation
@@ -73,6 +74,7 @@ popdynMICE<-function(qsx,qfracx,np,nf,nyears,nareas,maxage,Nx,VFx,FretAx,Effind,
     Fdist[Find]<-VBcur[Find]^Spat_targ[Find[,1:2]]
     #VBagg<-apply(Fdist,1:3,sum)
     Fdist[Find]<-Fdist[Find]/apply(Fdist,1:3,sum)[Find[,1:3]]
+    Fdist[is.na(Fdist)]<-0 # This is an NA catch for hermaphroditism
     Ecur[]<-Effind[,,y-1] #matrix(Effind[,,y-1],nrow=np,ncol=nf)
     Vcur[]<-VFx[,,,y-1]#array(VFx[,,,y-1],c(np,nf,maxage))
     FMx[Find]<-qsx[Find[,1]]*qfracx[Find[,1:2]]*Ecur[Find[,1:2]]*Fdist[Find]*Vcur[Find[,1:3]]/Asizex[Find[,c(1,4)]]
@@ -98,9 +100,9 @@ popdynMICE<-function(qsx,qfracx,np,nf,nyears,nareas,maxage,Nx,VFx,FretAx,Effind,
[…]

### 2019-03-06 R/MMSE_plots.R
New hermaphroditic code

@@ -11,15 +11,15 @@
 #' @export
 plotquant<-function(x,p=c(0.05,0.25,0.75,0.95),yrs,qcol,lcol,addline=T,ablines=NA){
   ny<-length(yrs)
-  qs<-apply(x,2,quantile,p=p[c(1,4)])
-  qsi<-apply(x,2,quantile,p=p[2:3])
+  qs<-apply(x,2,quantile,p=p[c(1,4)],na.rm=T)
+  qsi<-apply(x,2,quantile,p=p[2:3],na.rm=T)
   polygon(c(yrs,yrs[ny:1]),c(qs[1,],qs[2,ny:1]),border=NA,col='#b3ecff')
 
   polygon(c(yrs,yrs[ny:1]),c(qsi[1,],qsi[2,ny:1]),border=NA,col=qcol)
   if(!is.na(ablines[1]))abline(h=ablines,col='#99999980')
 
   if(addline)for(i in 1:2)lines(yrs,x[i,],col=lcol,lty=i)
-  lines(yrs,apply(x,2,quantile,p=0.5),lwd=2,col="white")
+  lines(yrs,apply(x,2,quantile,p=0.5,na.rm=T),lwd=2,col="white")
 }
 
 
@@ -84,7 +84,7 @@ plot.MMSE<-function(MMSE,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",qu
 
   F_FMSY<-MMSE@F_FMSY
   F_FMSYsum<-apply(F_FMSY,c(1,2,3,5),sum)
-  Flims<- c(0,quantile(F_FMSYsum,0.95))
+  Flims<- c(0,quantile(F_FMSYsum,0.95,na.rm=T))
 
   for(pp in 1:length(plots)){
 

### 2019-03-06 R/multiMSE.R

@@ -355,13 +355,24 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
   # if SexPars Sex specific exceptions (recalculation of SSB0, aR, bR, SSBpR)
   if(length(SexPars)>0){
 
+    message("You have specified sex-specific dynamics, unfished spawning biomass and specified stock depletion will be mirrored across sex types according to SexPars$SSBfrom")
+
     SSB0s<-matrix(NIL(StockPars,"SSB0"),nrow=nsim) # sim, p
     sexmatches<-sapply(1:nrow(SexPars$SSBfrom),function(x,mat)paste(mat[x,],collapse="_"), mat=SexPars$SSBfrom)
-    depcopy<-match(sexmatches,sexmatches)
+    parcopy<-match(sexmatches,sexmatches)
+    StockPars_t<-StockPars # need to store a temporary object for copying to/from
 
     for(p in 1:np){
 
       SSB0<-apply(matrix(rep(SexPars$SSBfrom[p,],each=nsim),nrow=nsim)*SSB0s,1,sum)
+      StockPars[[p]]$SSB0<-SSB0
+
+      # copied parameters
+      StockPars[[p]]$D<-StockPars_t[[parcopy[p]]]$D
+      StockPars[[p]]$hs<-StockPars_t[[parcopy[p]]]$hs
+      StockPars[[p]]$R0<-StockPars_t[[parcopy[p]]]$R0
+      StockPars[[p]]$R0a<-StockPars_t[[parcopy[p]]]$R0a
+
       StockPars[[p]]$SSBpR<-array(SSB0/StockPars[[p]]$R0,c(nsim,nareas)) # !!!!!!!!!!! SSBpR hardwired to be the same among areas !!!!
 
       idist<-StockPars[[p]]$R0a/apply(StockPars[[p]]$R0a,1,sum)
@@ -370,8 +381,11 @@ multiMSE_int <- function(MOM, MPs=list(c("AvC","DCAC"),c("FMSYref","curE")),
       StockPars[[p]]$bR <- matrix(log(5 * StockPars[[p]]$hs)/(0.8 * SSB0a), nrow=nsim)  # Ricker SR params
       StockPars[[p]]$aR <- matrix(exp(StockPars[[p]]$bR * SSB0a)/StockPars[[p]]$SSBpR, nrow=nsim)  # Ricker SR params
 
-      StockPars[[p]]$SSB0<-SSB0
[…]

### 2019-03-06 R/multiMSE_internal.R

@@ -72,7 +72,7 @@ HistMICE<-function(x,StockPars, FleetPars, np,nf, nareas, maxage, nyears, N, VF,
   qsx<-qs[x,]
   qfracx<-matrix(qfrac[x,,],c(np,nf))
 
-  popdynMICE(qsx=qsx,qfracx=qfracx,np,nf,nyears,nareas,maxage,Nx,VFx,FretAx,Effind,movx,Spat_targ,M_ageArrayx,Mat_agex,Asizex,Kx,Linfx,t0x,Mx,R0x,R0ax,SSBpRx,hsx,aRx, bRx,ax,bx,Perrx,SRrelx,Rel,SexPars)
+  popdynMICE(qsx=qsx,qfracx=qfracx,np,nf,nyears,nareas,maxage,Nx,VFx,FretAx,Effind,movx,Spat_targ,M_ageArrayx,Mat_agex,Asizex,Kx,Linfx,t0x,Mx,R0x,R0ax,SSBpRx,hsx,aRx, bRx,ax,bx,Perrx,SRrelx,Rel,SexPars,x)
 
 }
 

### 2019-03-06 R/multi_q_estimation.R

@@ -168,6 +168,15 @@ qestMICE<-function(par,depc,CFc,mode='opt',np,nf,nyears,nareas,maxage,Nx,VFx,Fre
 
   SSBest<-apply(HistVars$SSBx,c(1,3),sum)
   deppred<-SSBest[,nyears]/SSB0x
+
+  if(length(SexPars)>0){ # you need to make depletion just one variable for all components of a sex-specific model
+
+    sexmatches<-sapply(1:nrow(SexPars$SSBfrom),function(x,mat)paste(mat[x,],collapse="_"), mat=SexPars$SSBfrom)
+    parcopy<-match(sexmatches,sexmatches)
+    deppred<-deppred[parcopy]
+
+  }
+
   Cpred0<-array(NA,c(np,nf,maxage,nareas))
   Cind<-TEG(dim(Cpred0))
   Find<-cbind(Cind[,1:3],nyears,Cind[,4]) # p f age y area

### 2019-03-06 R/popdyn_MICE.R

@@ -174,18 +174,7 @@ popdynOneMICE<-function(np,nf,nareas, maxage, Ncur, Vcur, FMretx, FMx, PerrYrp, 
                         SRrelx,M_agecur,Mat_agecur,Asizex,
                         Kx,Linfx,t0x,Mx,R0x,R0ax,SSBpRx,ax,bx,Rel,SexPars,x){
 
-  if(length(SexPars$Herm)>0){ # Hermaphroditic mode
-    Ncur[is.na(Ncur)]<-0 # catch for NAs
-    for(i in 1:length(SexPars$Herm)){
-      ps<-as.numeric(strsplit(names(SexPars$Herm)[i],"_")[[1]][2:3])
-      pfrom<-ps[1]
-      pto<-ps[2]
-      frac<-rep(1,maxage)
-      frac[1:length(SexPars$Herm[[i]][x,])]<-SexPars$Herm[[i]][x,]
-      Ncur[pfrom,,]<-Ncur[pfrom,,]+frac*Ncur[pto,,] #)Nnext[pto,,])
-      Ncur[pto,,]<-(1-frac)*Ncur[pto,,]
-    }
-  }
+
 
   # Initial Bcur calc (before any weight at age recalculation change)
   # Bcalc ---------------------------------------------------------------------------
@@ -259,9 +248,25 @@ popdynOneMICE<-function(np,nf,nareas, maxage, Ncur, Vcur, FMretx, FMx, PerrYrp, 
                             R0a=R0ax[p,], SSBpR=SSBpRx[p,], aR=aRx[p,], bR=bRx[p,],
                             mov=movy[p,,,], SRrel=SRrelx[p])
     Nnext[p,,]<-NextYrN
+
+
   }
 
-  Nnext[is.na(Nnext)]<-0 # catch for NAs
+  if(length(SexPars$Herm)>0){ # Hermaphroditic mode
+    Nnext[is.na(Nnext)]<-0 # catch for NAs
+    for(i in 1:length(SexPars$Herm)){
+      ps<-as.numeric(strsplit(names(SexPars$Herm)[i],"_")[[1]][2:3])
+      pfrom<-ps[2]
+      pto<-ps[1]
+      frac<-rep(1,maxage)
+      frac[1:length(SexPars$Herm[[i]][x,])]<-SexPars$Herm[[i]][x,]
+      h_rate<-hrate(frac)
[…]

### 2019-03-08 R/MMSE_plots.R
multMSE 10/10
Updated and preliminary debugging checks for all modes of multiMSE

All combinations of:
one stock multi fleet / one fleet multi stock / multi fleet mulit stock
no sex specific / with sex -specific/ with sex-specific hermaphroditism
MMP / by stock / by fleet / complex
no allocations ./ Allocation / Efactor

@@ -39,6 +39,7 @@ plotquant<-function(x,p=c(0.05,0.25,0.75,0.95),yrs,qcol,lcol,addline=T,ablines=N
 #' @author T.Carruthers
 #' @export
 plot.MMSE<-function(MMSE,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",quants=c(0.05,0.25,0.75,0.95),curyr=2018,addline=F){
+  # maxcol=6;qcol=rgb(0.4,0.8,0.95); lcol= "dodgerblue4"; quants=c(0.05,0.25,0.75,0.95); curyr=2018; addline=F
 
 
   if(is.na(maxcol))maxcol=ceiling(length(MMSE@MPs)/0.5) # defaults to portrait 1:2
@@ -83,7 +84,7 @@ plot.MMSE<-function(MMSE,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",qu
   # --- F projection -----------------------------------------------------------
 
   F_FMSY<-MMSE@F_FMSY
-  F_FMSYsum<-apply(F_FMSY,c(1,2,3,5),sum,na.rm=T)
+  F_FMSYsum<-apply(F_FMSY,c(1,2,4,5),sum,na.rm=T)
   Flims<- c(0,quantile(F_FMSYsum,0.95,na.rm=T))
 
   for(pp in 1:length(plots)){
@@ -111,7 +112,7 @@ plot.MMSE<-function(MMSE,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",qu
   # --- Total yield projection -----------------------------------------------------
 
   Yd<-MMSE@C#MMSE@OM$RefY
-  Ydsum<-apply(Yd, c(1,2,3,5),sum,na.rm=T)
+  Ydsum<-apply(Yd, c(1,2,4,5),sum,na.rm=T)
   Ydsum<-Ydsum/array(rep(Ydsum[,,,1],MMSE@proyears),dim(Ydsum))
   Ylims<- c(0,quantile(Ydsum,0.95,na.rm=T))
 
@@ -153,7 +154,7 @@ plot.MMSE<-function(MMSE,maxcol=6,qcol=rgb(0.4,0.8,0.95), lcol= "dodgerblue4",qu
 
     for(MP in toplot){
       for(ss in 1:ns){
[…]

### 2019-03-08 R/MOM_object.R

@@ -37,7 +37,7 @@
 #' @slot Obs    Hierarchical List of Observation model objects Level 1 is stock, level 2 is fleet
 #' @slot Imps   Hierarchical List of Implementation model objects Level 1 is stock, level 2 is fleet
 #' @slot CatchFrac A list nstock long, of matrices nsim x nfleet representing the fraction of current catches of the various fleets to each stock (each matrix is nsim by nfleet long and rows sum to 1 for each stock)
-#' @slot Allocation  A list nstock long, of vector allocations by fleet (default is NULL and allocation is according to last historical year).
+#' @slot Allocation  A list nstock long, of matrices nsim x nfleet representing the fraction of future TACs of the various fleets to each stock (each matrix is nsim by nfleet long and rows sum to 1 for each stock).
 #' @slot Efactor A list nstock long, of current effort factors by fleet (default is 1 - same as current effort)
 #' @slot Complexes A list of stock complexes. Each position is a vector of stock numbers (as they appear in StockPars) for which data should be aggregated and TAC recommendations split among stocks according to vulnerable biomass
 #' @slot SexPars A list of slots that control sex-specific dynamics

### 2019-03-08 R/multiMSE_internal.R

@@ -374,3 +374,17 @@ expandHerm<-function(Herm,maxage,np,nsim){
   HermFrac
 }
 
+
+#' Fill any NAs arising from MPCalcs (hermaphroditism mode)
+#'
+#' @param MPCalcs A list of arrays arising fromt the DLMtool function CalcMPDynamics()
+#' @author T. Carruthers
+#' @export
+MPCalcsNAs<-function(MPCalcs){
+
+  for(i in 1:length(MPCalcs))MPCalcs[[i]][is.na(MPCalcs[[i]])]<-0
+  MPCalcs
+
+}
+
+

### 2019-03-08 R/popdyn_MICE.R

@@ -174,8 +174,6 @@ popdynOneMICE<-function(np,nf,nareas, maxage, Ncur, Vcur, FMretx, FMx, PerrYrp, 
                         SRrelx,M_agecur,Mat_agecur,Asizex,
                         Kx,Linfx,t0x,Mx,R0x,R0ax,SSBpRx,ax,bx,Rel,SexPars,x){
 
-
-
   # Initial Bcur calc (before any weight at age recalculation change)
   # Bcalc ---------------------------------------------------------------------------
   Bcur<-SSBcur<-SSNcur<-array(NA,dim(Ncur))
@@ -236,7 +234,7 @@ popdynOneMICE<-function(np,nf,nareas, maxage, Ncur, Vcur, FMretx, FMx, PerrYrp, 
   Nnext<-array(NA,c(np,maxage,nareas))
 
   # just for vulnerable biomass calculation
-  SumF<-apply(FMx,c(1,3:4),sum) # sum over fleets stock,age,area
+  SumF<-apply(FMx,c(1,3:4),sum,na.rm=T) # sum over fleets stock,age,area
   MaxF<-apply(SumF,c(1,3),max)     # get max F
   Selx<-array(NA,dim(SumF))
   Selx[Nind]<-SumF[Nind]/MaxF[Nind[,c(1,3)]]

### 2019-03-21 R/Misc_exported.R
Create folder for Asesssment figures

@@ -51,6 +51,6 @@ getclass <- function(x, classy) any(inherits(get(x), classy))
 #' @seealso \link[DLMtool]{userguide}
 #' @export
 userguide <- function() {
-  message("For the DLMtool user guide, type in 'DLMtool::userguide()' to the console.")
+  message("For the DLMtool user guide, type in \"DLMtool::userguide()\" to the console.")
   browseVignettes("MSEtool")
 }

### 2019-03-27 R/define_S4_class.R
Clean up Assessment vignette

@@ -133,8 +133,6 @@ setMethod("summary", signature(object = "Assessment"), function(object) {
 #' @title Assessment object
 #' @description Produces HTML file with figures of parameter estimates and output from an \linkS4class{Assessment} object.
 #'
-#' @aliases plot,Assessment-method
-#'
 #' @param x An object of class \linkS4class{Assessment}
 #' @param save_figure Indicates whether figures will be saved to disk. A corresponding
 #' html report will be produced.

### 2019-04-12 R/SCA.R
Fix VPA bugs and add to vignette

@@ -518,7 +518,7 @@ SCA_refpt_calc <- function(E, R, weight, mat, M, vul, SR, fix_h, h) {
     if(SR == "BH") Rpred <<- (0.8 * R0 * h * E)/(0.2 * EPR0 * R0 *(1-h)+(h-0.2)*E)
     if(SR == "Ricker") Rpred <<- E/EPR0 * (5*h)^(1.25 * (1 - E/E0))
     sigmaR <<- sqrt(sum((log(R/Rpred))^2)/length(Rpred))
-    nLL <- -sum(dnorm(log(R/Rpred), 0.5 * sigmaR^2, sigmaR, log = TRUE))
+    nLL <- -sum(dnorm(log(R/Rpred), -0.5 * sigmaR^2, sigmaR, log = TRUE))
     return(nLL)
   }
 

### 2019-04-12 R/SCA2.R

@@ -240,7 +240,7 @@ SCA2 <- function(x = 1, Data, SR = c("BH", "Ricker"), vulnerability = c("logisti
     SE_Dev <- c(rev(SE_Early), SE_Main)
 
     Assessment@FMSY <- report$FMSY
-    Assessment@MSY < report$MSY
+    Assessment@MSY <- report$MSY
     Assessment@BMSY <- report$BMSY
     Assessment@SSBMSY <- report$EMSY
     Assessment@VBMSY <- report$VBMSY

### 2019-04-12 R/VPA.R

@@ -150,6 +150,8 @@ VPA <- function(x = 1, Data, expanded = FALSE, SR = c("BH", "Ricker"), vulnerabi
   }
 
   CAA_hist2[is.na(CAA_hist2) | CAA_hist2 < 1e-8] <- 1e-8
+  max_ind <- CAA_hist2[, ncol(CAA_hist2) - 1] == 1e-8
+  CAA_hist2[max_ind, ncol(CAA_hist2) - 1] <- CAA_hist2[max_ind, ncol(CAA_hist2)]
 
   Wa2 <- Wa[min_age:max_age2]
   Wa2[length(Wa2)] <- mean(Wa[max_age2:max_age])
@@ -157,6 +159,7 @@ VPA <- function(x = 1, Data, expanded = FALSE, SR = c("BH", "Ricker"), vulnerabi
   La2[length(Wa2)] <- mean(La[max_age2:max_age])
   mat_age2 <- mat_age[min_age:max_age2]
   mat_age2[length(mat_age2)] <- mean(mat_age[max_age2:max_age])
+  mat_age2 <- mat_age2/max(mat_age2)
 
   LH <- list(LAA = La2, WAA = Wa2, Linf = Linf, K = K, t0 = t0, a = a, b = b, A50 = A50, A95 = A95)
 
@@ -288,8 +291,8 @@ VPA <- function(x = 1, Data, expanded = FALSE, SR = c("BH", "Ricker"), vulnerabi
     info$vul_refpt <- apply(report$vul[(length(Year)-vul_pen[1]+1):length(Year), , drop = FALSE], 2, mean)
     info$vul_refpt <- info$vul_refpt/max(info$vul_refpt)
 
-    E <- report$SSB[1:(length(report$SSB)-min_age)]
-    R <- report$N[(min_age + 1):length(report$SSB), 1]
+    E <- report$E[1:(length(report$E)-min_age)]
+    R <- report$N[(min_age + 1):length(report$E), 1]
     ref_pt <- SCA_refpt_calc(E, R, data$weight, data$mat, data$M, info$vul_refpt, SR, fix_h, info$h)
 
     report <- c(report, ref_pt)

### 2019-04-12 R/report_SCA2.R

@@ -5,7 +5,7 @@ summary_SCA2 <- function(Assessment) {
   if(conv) current_status <- c(F_FMSY[length(F_FMSY)], B_BMSY[length(B_BMSY)], B_B0[length(B_B0)])
   else current_status <- rep(NA, 3)
   current_status <- data.frame(Value = current_status)
-  rownames(current_status) <- c("U/UMSY", "B/BMSY", "B/B0")
+  rownames(current_status) <- c("F/FMSY", "B/BMSY", "B/B0")
 
   Value <- c(info$data$M[1], info$data$max_age, info$LH$Linf, info$LH$K, info$LH$t0,
              info$LH$a * info$LH$Linf ^ info$LH$b, info$LH$A50, info$LH$A95)

### 2019-04-12 R/report_VPA.R

@@ -280,10 +280,10 @@ generate_plots_VPA <- function(Assessment, save_figure = FALSE, save_dir = tempd
     if(info$SR == "Ricker") expectedR <- Arec * SSB_plot * exp(-Brec * SSB_plot)
     estR <- R[as.numeric(names(R)) > Year[1]]
 
-    plot_SR(SSB_plot, expectedR, rec_dev = estR, ylab = paste0("Recruitment (age-", min(age), ")"))
+    plot_SR(SSB_plot, expectedR, R0, SSB0, rec_dev = estR, ylab = paste0("Recruitment (age-", min(age), ")"))
     if(save_figure) {
       create_png(filename = file.path(plot.dir, "assessment_stock_recruit.png"))
-      plot_SR(SSB_plot, expectedR, rec_dev = estR, ylab = paste0("Recruitment (age-", min(age), ")"))
+      plot_SR(SSB_plot, expectedR, R0, SSB0, rec_dev = estR, ylab = paste0("Recruitment (age-", min(age), ")"))
       dev.off()
       assess.file.caption <- rbind(assess.file.caption,
                                    c("assessment_stock_recruit.png", "Stock-recruitment relationship."))
@@ -301,14 +301,13 @@ generate_plots_VPA <- function(Assessment, save_figure = FALSE, save_dir = tempd
       }
     } else y_zoom <- NULL
 
-    plot_SR(SSB_plot, expectedR, rec_dev = estR, trajectory = TRUE, y_zoom = y_zoom, ylab = paste0("Recruitment (age-", min(age), ")"))
+    plot_SR(SSB_plot, expectedR, R0, SSB0, rec_dev = estR, trajectory = TRUE, y_zoom = y_zoom, ylab = paste0("Recruitment (age-", min(age), ")"))
     if(save_figure) {
       create_png(filename = file.path(plot.dir, "assessment_stock_recruit_trajectory.png"))
-      plot_SR(SSB_plot, expectedR, rec_dev = estR, trajectory = TRUE, y_zoom = y_zoom, ylab = paste0("Recruitment (age-", min(age), ")"))
+      plot_SR(SSB_plot, expectedR, R0, SSB0, rec_dev = estR, trajectory = TRUE, y_zoom = y_zoom, ylab = paste0("Recruitment (age-", min(age), ")"))
       dev.off()
       assess.file.caption <- rbind(assess.file.caption,
[…]

### 2019-04-12 inst/include/VPA.h

@@ -70,11 +70,13 @@
       if(a==max_age-1) {
         F(y-1,a-1) = Newton_VPA_F_plus(F(y,a), F_ratio, M(a-1), M(a), CAA_hist(y-1,a-1), CAA_hist(y-1,a), N(y,a), nitF);
       } else {
-        F(y-1,a-1) = Newton_VPA_F(F(y,a), M(a-1), CAA_hist(y-1,a-1), N(y,a), nitF);
+        F(y-1,a-1) = CppAD::CondExpGt(CAA_hist(y-1,a-1), Type(1e-4), Newton_VPA_F(F(y,a), M(a-1), CAA_hist(y-1,a-1), N(y,a), nitF), Type(1e-4));
+        //F(y-1,a-1) = Newton_VPA_F(F(y,a), M(a-1), CAA_hist(y-1,a-1), N(y,a), nitF);
       }
 
-      N(y-1,a-1) = (F(y-1,a-1) + M(a-1)) * CAA_hist(y-1,a-1);
-      N(y-1,a-1) /= (1 - exp(-F(y-1,a-1) - M(a-1))) * F(y-1,a-1);
+      N(y-1,a-1) = CppAD::CondExpGt(CAA_hist(y-1,a-1), Type(1e-4), (F(y-1,a-1) + M(a-1)) * CAA_hist(y-1,a-1)/(1 - exp(-F(y-1,a-1) - M(a-1)))/F(y-1,a-1), N(y,a) * exp(F(y-1,a-1) + M(a)));
+      //N(y-1,a-1) = (F(y-1,a-1) + M(a-1)) * CAA_hist(y-1,a-1);
+      //N(y-1,a-1) /= (1 - exp(-F(y-1,a-1) - M(a-1))) * F(y-1,a-1);
 
       if(a==max_age-1) {
         F(y-1,a) = F_ratio * F(y-1,a-1);

### 2019-09-17 inst/rmarkdown_templates/SS2OM.Rmd
Update SS2OM

---
title: "Operating model summary"
subtitle: "Comparison of OM and Stock Synthesis output"
date: "`r Sys.Date()`"
---
<style type="text/css">
h1 { /* Header 1 */
  font-size: 24px;
}
</style>

```{r setup, include = FALSE, echo = FALSE}
  knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE,
  fig.width = 6, fig.height = 4.5, out.width = "650px", comment = "#>")
```

# {.tabset}

## Spawning biomass

```{r, fig.cap = "Spawning biomass from the operating model (black lines and dots) and Stock Synthesis (red lines)."}
ylim_SSB <- c(0, 1.1 * max(c(SSB, Hist@TSdata$SSB)))
matplot(mainyrs, t(Hist@TSdata$SSB), xlab = "Year", ylab = "Spawning Biomass", ylim = ylim_SSB, pch = 1, col = "black", typ = "o")
lines(mainyrs, SSB, col = "red", lwd = 3)
abline(h = 0, col = "grey")
```

```{r, fig.cap = "Spawning depletion from the operating model (black lines and dots) and Stock Synthesis (red lines).."}
ylim_dep <- c(0, 1.1 * max(SSB/SSB0, Hist@TSdata$SSB/Hist@Ref$SSB0))
matplot(mainyrs, t(Hist@TSdata$SSB/Hist@Ref$SSB0), xlab = "Year", ylab = "Spawning Depletion", ylim = ylim_dep, pch = 1, col = "black", typ = "o")
lines(mainyrs, SSB/SSB0, col = "red", lwd = 3)
abline(h = 0, col = "grey")
```

## Catch

```{r, fig.cap = "Catch from the operating model (black lines and dots) and Stock Synthesis (red lines)."}
Catch_SS <- cbind(replist$timeseries[, 1:4], rowSums(replist$timeseries[, startsWith(names(replist$timeseries), "dead(B):")], na.rm = TRUE))
Catch_SS <- aggregate(Catch_SS[, 5], list(Year = Catch_SS$Yr), sum)
Catch_SS <- Catch_SS[vapply(Catch_SS$Year, "%in%", logical(1), mainyrs), 2]

ylim_cat <- c(0, 1.1 * max(Catch_SS, Hist@TSdata$Catch))
matplot(mainyrs, t(Hist@TSdata$Catch), xlab = "Year", ylab = "Catch", ylim = ylim_cat, pch = 1, col = "black", typ = "o")
lines(mainyrs, Catch_SS, col = "red", lwd = 3)
abline(h = 0, col = "grey")
```

```{r, fig.cap = "Relative catch from the operating model (black lines and dots) and Stock Synthesis (red lines)."}
ylim_cat <- c(0, 1.1 * max(Catch_SS/Catch_SS[length(Catch_SS)], Hist@TSdata$Catch/Hist@TSdata$Catch[, OM@nyears]))
matplot(mainyrs, t(Hist@TSdata$Catch/Hist@TSdata$Catch[, OM@nyears]), xlab = "Year", ylab = "Catch Relative to today", ylim = ylim_cat, pch = 1, col = "black", typ = "o")
lines(mainyrs, Catch_SS/Catch_SS[length(Catch_SS)], col = "red", lwd = 3)
abline(h = 0, col = "grey")
abline(h = 1, lty = 3)
```

## Abundance-at-age {.tabset}

### Recruitment

```{r, fig.cap = "Recruitment from the operating model (black lines and dots) and Stock Synthesis (red lines)."}
N_at_age <- melt(replist$natage[replist$natage$`Beg/Mid` == "B" & replist$natage$Seas == 1, -c(9:11)], list("Area", "Bio_Pattern", "Sex", "BirthSeas", "Platoon", "Morph", "Seas", "Yr"),
                 variable.name = "Age", value.name = "N")
N_at_age <- N_at_age[as.numeric(N_at_age$Age)-1 >= age_rec * ifelse(season_as_years, nseas, 1), ]  # Subset by age >= age_rec
N_at_age <- N_at_age[vapply(N_at_age$Yr, "%in%", logical(1), table = mainyrs), ]  # Subset by year
N_at_age <- N_at_age[vapply(N_at_age$Sex, "%in%", logical(1), table = gender), ] # Subset by gender

N_at_age <- summarise(group_by(N_at_age, Yr, Age), N = sum(N)) # Sum over area, morphs, platoons, sex, etc.

N_at_age <- acast(N_at_age, list("Yr", "Age"), value.var = "N")

ylim_rec <- c(0, 1.1 * max(Hist@AtAge$Nage[, 1, ], N_at_age[, 1]))
matplot(mainyrs, t(Hist@AtAge$Nage[, 1, ]), xlab = "Year", ylab = "Recruitment (age 1)", ylim = ylim_rec, pch = 1, col = "black", typ = "o")
lines(mainyrs, N_at_age[, 1], col = "red", lwd = 3)
abline(h = 0, col = "grey")
```

### Absolute abundance

```{r, fig.cap = "Annual abundance at age from the operating model (black lines and dots) and Stock Synthesis (red lines)."}
plot_composition(mainyrs, t(Hist@AtAge$Nage[1, , 1:nyears]), N_at_age, N = NULL, annual_ylab = "Numbers at age", annual_yscale = "raw")
```

### Relative abundance

```{r, fig.cap = "Annual proportions at age from the operating model (black lines and dots) and Stock Synthesis (red lines)."}
plot_composition(mainyrs, t(Hist@AtAge$Nage[1, , 1:nyears]), N_at_age, N = NULL, annual_ylab = "Proportions at age")
```

## Selectivity and F

```{r, fig.cap = "Annual selectivity at age from the operating model (black lines and dots) and Stock Synthesis (red lines)."}
plot_composition(mainyrs, t(Hist@AtAge$Select[1, , 1:nyears]), t(OM@cpars$V[1, , 1:nyears]), N = NULL, annual_ylab = "Selectivity", annual_yscale = "raw")
```

```{r, fig.cap = "Apical F from the operating model (black lines and dots) and Stock Synthesis (red lines)."}
Hist_F <- apply(Hist@AtAge$FM, c(1, 3), max)
ylim_F <- c(0, 1.1 * max(c(OM@cpars$Find, Hist_F)))
matplot(mainyrs, t(Hist_F), typ = "o", xlab = "Year", ylab = "Apical F", ylim = ylim_F, pch = 1, col = "black")
lines(mainyrs, OM@cpars$Find[1, ], col = "red", lwd = 3)
abline(h = 0, col = "grey")
```

## About

This report was generated on: `r Sys.time()`<br />
MSEtool version `r packageVersion("MSEtool")`<br />
DLMtool version `r packageVersion("DLMtool")`<br />
`r R.version.string`<br />

### 2019-09-17 R/SS2OM.R

@@ -12,13 +12,21 @@
 #' @param reps The number of stochastic replicates within each simulation in the operating model.
 #' @param maxF The maximum allowable F in the operating model.
 #' @param seed The random seed for the operating model.
-#' @param Obs The observation model (class Obs).
-#' @param Imp The implementation model (class Imp).
+#' @param interval The interval at which management procedures will update the management advice in \link[DLMtool]{runMSE}, e.g., 1 = annual updates.
+#' @param Obs The observation model (class Obs). This function only updates the catch and index observation error.
+#' @param Imp The implementation model (class Imp). This function does not update implementation parameters.
 #' @param import_mov Logical, whether to import movement matrix from the assessment.
-#' @param gender An integer index for the sex for which life history values are importing (1 = female, 2 = male).
+#' @param gender An integer that indexes the sex for importing life history parameters (1 = usually female, 2 = usually male, 1:2 = mean across both sexes).
+#' @param age_rec Integer for the age of recruitment. The default is 1 for DLMtool operating models. Generally, should not be changed.
+#' @param silent Whether to silence messages to the console.
 #' @param Name The name of the operating model
 #' @param Source Reference to assessment documentation e.g. a url
 #' @param Author Who did the assessment
+#' @param report Logical, if TRUE, the function will run \link[DLMtool]{runMSE} to generate historical data from the operating model
+#' to compare against SS output. A markdown report will be generated.
+#' @param filename If \code{report = TRUE}, character string for the name of the markdown and HTML files.
+#' @param dir If \code{report = TRUE}, the directory in which the markdown and HTML files will be saved.
+#' @param open_file If \code{report = TRUE}, whether the HTML document is opened after it is rendered.
 #' @param ... Arguments to pass to \link[r4ss]{SS_output}.
[…]

### 2019-09-17 R/define_S4_class_assess.R

@@ -169,6 +169,9 @@ if(getRversion() >= "2.15.1") {
 
   # For Awatea2OM - Quang assumes these variables are loaded in from .rda files
   utils::globalVariables(c("Bmcmc", "currentMCMC", "currentRes"))
+
+  # For dplyr functions in SS2OM
+  utils::globalVariables(c("Year", "Age", "M"))
 }
 
 

### 2019-09-17 R/report_auxiliary.R

@@ -642,9 +642,6 @@ plot_composition <- function(Year = 1:nrow(obs), obs, fit = NULL, plot_type = c(
                              N = rowSums(obs), CAL_bins = NULL, ages = NULL, ind = 1:nrow(obs),
                              annual_ylab = "Frequency", annual_yscale = c("proportions", "raw"),
                              bubble_adj = 5, fit_linewidth = 3, fit_color = "red") {
-  old_par <- par(no.readonly = TRUE)
-  on.exit(par(old_par))
-
   plot_type <- match.arg(plot_type)
   annual_yscale <- match.arg(annual_yscale)
   if(is.null(CAL_bins)) data_type <- "age" else data_type <- "length"
@@ -659,7 +656,7 @@ plot_composition <- function(Year = 1:nrow(obs), obs, fit = NULL, plot_type = c(
     data_val <- if(is.null(ages)) 1:ncol(obs) else ages
     data_lab <- "Age"
   }
-  N <- round(N, 1)
+  if(!is.null(N)) N <- round(N, 1)
 
   if(annual_yscale == "proportions") {
     obs_prob_all <- obs/rowSums(obs, na.rm = TRUE)
@@ -677,7 +674,7 @@ plot_composition <- function(Year = 1:nrow(obs), obs, fit = NULL, plot_type = c(
   Year <- Year[ind]
   obs <- obs[ind, , drop = FALSE]
   if(!is.null(fit)) fit <- fit[ind, , drop = FALSE]
-  N <- N[ind]
+  if(!is.null(N)) N <- N[ind]
 
   # Bubble plot (obs)
   if('bubble_data' %in% plot_type) {
@@ -736,7 +733,8 @@ plot_composition <- function(Year = 1:nrow(obs), obs, fit = NULL, plot_type = c(
 
   # Annual comps (obs vs. fitted if available)
   if('annual' %in% plot_type) {
-
+    old_par <- par(no.readonly = TRUE)
+    on.exit(par(old_par))
     par(mfcol = c(4, 4), mar = rep(0, 4), oma = c(5.1, 5.1, 2.1, 2.1))
     ylim <- c(0, 1.1 * max(obs_prob_all, fit_prob_all, na.rm = TRUE))
 
@@ -782,7 +780,7 @@ plot_composition <- function(Year = 1:nrow(obs), obs, fit = NULL, plot_type = c(
           fit.vec <- fit_prob[i, ]
           lines(data_val, fit.vec, lwd = fit_linewidth, col = fit_color)
[…]

### 2019-09-27 R/Misc_fun.R
Add r prior to SP

@@ -174,3 +174,29 @@ rescale_report <- function(var_div, var_mult, var_trans = NULL, fun_trans = NULL
   }
   invisible()
 }
+
+
+sample_steepness3 <- function(n, mu, cv, SR_type = c("BH", "Ricker")) {
+  if(n == 1) {
+    return(mu)
+  } else if(SR_type == "BH") {
+    sigma <- mu * cv
+    mu.beta.dist <- (mu - 0.2)/0.8
+    sigma.beta.dist <- sigma/0.8
+    beta.par <- derive_beta_par(mu.beta.dist, sigma.beta.dist)
+    h.transformed <- rbeta(n, beta.par[1], beta.par[2])
+    h <- 0.8 * h.transformed + 0.2
+    h[h > 0.99] <- 0.99
+    h[h < 0.2] <- 0.2
+    return(h)
+  } else {
+    sigma <- mu * cv
+    mu.lognorm.dist <- mconv(mu - 0.2)
+    sigma.lognorm.dist <- sigma
+
+    h.transformed <- trlnorm(n, mconv(mu.lognorm.dist, sigma.lognorm.dist), sdconv(mu.lognorm.dist, sigma.lognorm.dist))
+    h <- h.transformed + 0.2
+    h[h < 0.2] <- 0.2
+    return(h)
+  }
+}

### 2019-09-27 R/assess_SP.R

@@ -4,7 +4,8 @@
 #' and coded in TMB. The base model, \code{SP}, is conditioned on catch and estimates a predicted index.
 #' Continuous surplus production and fishing is modeled with sub-annual time steps which should approximate
 #' the behavior of ASPIC (Prager 1994). The Fox model, \code{SP_Fox}, fixes BMSY/K = 0.37 (1/e).
-#' The state-space version, \code{SP_SS} estimates annual deviates in biomass.
+#' The state-space version, \code{SP_SS} estimates annual deviates in biomass. An option allows for setting a
+#' prior for the intrinsic rate of increase.
 #' The function for the \code{spict} model (Pedersen and Berg, 2016) is available in \link[DLMtool]{DLMextra}.
 #'
 #' @param x An index for the objects in \code{Data} when running in \link[DLMtool]{runMSE}.
@@ -32,6 +33,11 @@
 #' Ignored if \code{n_seas} = 1.
 #' @param integrate Logical, whether the likelihood of the model integrates over the likelihood
 #' of the biomass deviations (thus, treating it as a state-space variable).
+#' @param r_prior Logical, whether a prior for the intrinsic rate of increase will be used in the model. See details.
+#' @param r_reps If \code{r_prior = TRUE}, the number of samples of natural mortality and steepness for calculating the
+#' mean and standard deviation of the r prior.
+#' @param SR_type If \code{r_prior = TRUE}, the stock-recruit relationship used to calculate unfished recruits per spawner.
+#' to calculate the r prior.
 #' @param silent Logical, passed to \code{\link[TMB]{MakeADFun}}, whether TMB
 #' will print trace information during optimization. Used for dignostics for model convergence.
 #' @param opt_hess Logical, whether the hessian function will be passed to \code{\link[stats]{nlminb}} during optimization
@@ -51,6 +57,11 @@
 #' For \code{SP_SS}, a start value can also be provided for \code{sigma} and \code{tau}, the standard deviation
 #' of the index and log-biomass deviates, respectively. Default for tau is 0.1. Deviations are estimated beginning in the year when index
[…]

### 2019-09-27 R/z_MPs.R

@@ -147,7 +147,8 @@ get_dependencies <- function(Assess, arg = list()) {
   return(dep)
 }
 
-dep_args <- list(fix_h = "Data@steep", fix_sigma = "Data@CV_Ind", fix_tau = "Data@sigmaR", fix_omega = "Data@CV_Cat")
+dep_args <- list(fix_h = "Data@steep", fix_sigma = "Data@CV_Ind", fix_tau = "Data@sigmaR", fix_omega = "Data@CV_Cat",
+                 r_prior = "Data@Mort, Data@CV_Mort, Data@steep, Data@CV_steep, Data@vbLinf, Data@vbK, Data@vbt0, Data@wla, Data@wlb, Data@MaxAge, Data@L50, Data@L95")
 
 
 

### 2019-09-27 inst/include/SP.h

@@ -9,6 +9,7 @@
   DATA_INTEGER(nstep);
   DATA_SCALAR(dt);
   DATA_INTEGER(nitF);
+  DATA_VECTOR(r_prior);
 
   PARAMETER(log_FMSY);
   PARAMETER(log_MSY);
@@ -27,7 +28,7 @@
 
   Type BMSY = MSY/FMSY;
   Type K = BMSY / n_term2;
-  Type r = MSY * n_term / K;
+  Type r = MSY * n_term / K; // r = FMSY * pow(n, 1/(1-n))
 
   vector<Type> B(ny+1);
   vector<Type> SP(ny);
@@ -35,8 +36,11 @@
   vector<Type> Ipred(ny);
   vector<Type> F(ny);
 
-  Type nll = 0.;
-  Type penalty = 0.;
+  Type nll = 0;
+  Type penalty = 0;
+  Type prior = 0;
+
+  if(r_prior(0) > 0) prior -= dnorm(r, r_prior(0), r_prior(1), true);
 
   Cpred.setZero();
 
@@ -55,7 +59,7 @@
     if(!R_IsNA(asDouble(I_hist(y)))) nll -= dnorm(log(I_hist(y)), log(Ipred(y)), sigma, true);
   }
 
-  nll += penalty;
+  nll += penalty + prior;
 
   Type F_FMSY_final = F(F.size()-1)/FMSY;
   Type B_BMSY_final = B(B.size()-1)/BMSY;
@@ -87,6 +91,7 @@
   REPORT(F);
   REPORT(nll);
   REPORT(penalty);
+  REPORT(prior);
 
   return nll;
 

### 2019-09-27 R/assess_SP.R
Minor fixes to r prior in SP

@@ -33,11 +33,11 @@
 #' Ignored if \code{n_seas} = 1.
 #' @param integrate Logical, whether the likelihood of the model integrates over the likelihood
 #' of the biomass deviations (thus, treating it as a state-space variable).
-#' @param r_prior Logical, whether a prior for the intrinsic rate of increase will be used in the model. See details.
-#' @param r_reps If \code{r_prior = TRUE}, the number of samples of natural mortality and steepness for calculating the
+#' @param use_r_prior Logical, whether a prior for the intrinsic rate of increase will be used in the model. See details.
+#' @param r_reps If \code{use_r_prior = TRUE}, the number of samples of natural mortality and steepness for calculating the
 #' mean and standard deviation of the r prior.
-#' @param SR_type If \code{r_prior = TRUE}, the stock-recruit relationship used to calculate unfished recruits per spawner.
-#' to calculate the r prior.
+#' @param SR_type If \code{use_r_prior = TRUE}, the stock-recruit relationship used to calculate unfished recruits per spawner at the origin
+#' of spwaning biomass approaches zero. Used for the r prior.
 #' @param silent Logical, passed to \code{\link[TMB]{MakeADFun}}, whether TMB
 #' will print trace information during optimization. Used for dignostics for model convergence.
 #' @param opt_hess Logical, whether the hessian function will be passed to \code{\link[stats]{nlminb}} during optimization
@@ -58,7 +58,11 @@
 #' of the index and log-biomass deviates, respectively. Default for tau is 0.1. Deviations are estimated beginning in the year when index
 #' data are available.
 #'
-#' If \code{r_prior = TRUE}, a prior is created using the Euler-Lotka method (Equation 15a of McAllister et al. 2001).
+#' If \code{use_r_prior = TRUE}, \code{SP} and \code{SP_SS} will use a prior for the intrinsic rate of increase in the objective function.
[…]

### 2019-10-23 R/SRA_scope.R
Update SRA_scope vignette

@@ -386,7 +386,7 @@ SRA_scope <- function(OM, Chist = NULL, Ehist = NULL, condition = c("catch", "ef
   message("Range of unfished recruitment (OM@cpars$R0): ", paste(round(range(OM@cpars$R0), 2), collapse = " - "))
 
   ### Depletion and init D
-  if(any(C_eq > 0) || any(E_eq) > 0) {
+  if(any(C_eq > 0) || any(E_eq > 0)) {
     initD <- vapply(res, function(x) x$E[1]/x$E0[1], numeric(1))[keep]
     message("Estimated range in initial spawning depletion: ", paste(round(range(initD), 2), collapse = " - "))
   }
@@ -670,7 +670,7 @@ SRA_scope_est <- function(x = 1, Catch = NULL, Effort = NULL, Index = NULL, cond
     } else {
       vul_par[3:4, ff] <- c(-20, logit(FleetPars$Vmaxlen[nyears, x]))
       map_vul_par[3, ff] <- NA
-      if(fix_dome) map_ful_par[4, ff] <- NA
+      if(fix_dome) map_vul_par[4, ff] <- NA
     }
   }
 

### 2019-10-23 R/define_S4_class_SRA.R

@@ -247,18 +247,32 @@ setMethod("plot", signature(x = "SRA", y = "missing"),
                                "```{r, fig.cap = \"Apical F from the operating model.\"}",
                                "Hist_F <- apply(Hist@AtAge$FM, c(1, 3), max)",
                                "matplot(Year, t(Hist_F), typ = \"l\", col = \"black\", xlab = \"Year\", ylab = \"OM Apical F\", ylim = c(0, 1.1 * max(Hist_F)))",
+                               "abline(h = 0, col = \"grey\")",
                                "```\n",
                                "",
                                "```{r, fig.cap = \"Spawning biomass (SSB) from the operating model.\"}",
                                "matplot(Year, t(Hist@TSdata$SSB), typ = \"l\", col = \"black\", xlab = \"Year\", ylab = \"OM SSB\", ylim = c(0, 1.1 * max(Hist@TSdata$SSB)))",
+                               "abline(h = 0, col = \"grey\")",
+                               "```\n",
+                               "",
+                               "```{r, fig.cap = \"Spawning biomass (SSB) relative to MSY from the operating model.\"}",
+                               "matplot(Year, t(Hist@TSdata$SSB/Hist@Ref$SSBMSY), typ = \"l\", col = \"black\", xlab = \"Year\", ylab = expression(OM~~SSB/SSB[MSY]), ylim = c(0, 1.1 * max(Hist@TSdata$SSB/Hist@Ref$SSBMSY)))",
+                               "abline(h = 0, col = \"grey\")",
+                               "```\n",
+                               "",
+                               "```{r, fig.cap = \"Spawning depletion from the operating model.\"}",
+                               "matplot(Year, t(Hist@TSdata$SSB/Hist@Ref$SSB0), typ = \"l\", col = \"black\", xlab = \"Year\", ylab = expression(OM~~SSB/SSB[0]), ylim = c(0, 1.1 * max(Hist@TSdata$SSB/Hist@Ref$SSB0)))",
+                               "abline(h = 0, col = \"grey\")",
                                "```\n",
                                "",
                                "```{r, fig.cap = \"Recruitment from the operating model.\"}",
                                "matplot(Year, t(Hist@TSdata$Rec), typ = \"l\", col = \"black\", xlab = \"Year\", ylab = \"OM Recruitment\", ylim = c(0, 1.1 * max(Hist@TSdata$Rec)))",
+                               "abline(h = 0, col = \"grey\")",
                                "```\n",
                                "",
[…]

### 2019-10-24 R/assess_DD.R
Add projection function, fixes #12

@@ -17,8 +17,8 @@
 #' @param fix_U_equilibrium Logical, whether the equilibrium harvest rate prior to the first year of the model is
 #' estimated. If TRUE, U_equilibruim is fixed to value provided in start (if provided), otherwise, equal to zero
 #' (assumes virgin conditions).
-#' @param fix_sigma Logical, whether the standard deviation of the catch is fixed. If \code{TRUE},
-#' sigma is fixed to value provided in \code{start} (if provided), otherwise, value based on \code{Data@@CV_Cat}.
+#' @param fix_omega Logical, whether the standard deviation of the catch is fixed. If \code{TRUE},
+#' omega is fixed to value provided in \code{start} (if provided), otherwise, value based on \code{Data@@CV_Cat}.
 #' @param fix_tau Logical, the standard deviation of the recruitment deviations is fixed. If \code{TRUE},
 #' tau is fixed to value provided in \code{start} (if provided), otherwise, equal to 1.
 #' @param integrate Logical, whether the likelihood of the model integrates over the likelihood
@@ -42,7 +42,7 @@
 #' To provide starting values for \code{DD_TMB}, a named list can be provided for \code{R0} (virgin recruitment),
 #' \code{h} (steepness), and \code{q} (catchability coefficient) via the \code{start} argument (see example).
 #'
-#' For \code{DD_SS}, additional start values can be provided for and \code{sigma} and \code{tau}, the standard
+#' For \code{DD_SS}, additional start values can be provided for and \code{omega} and \code{tau}, the standard
 #' deviation of the catch and recruitment variability, respectively.
 #' @note Similar to many other assessment
 #' models, the model depends on assumptions such as stationary productivity and
@@ -112,7 +112,8 @@ DD_TMB <- function(x = 1, Data, SR = c("BH", "Ricker"), rescale = "mean1", start
   I_hist <- Data@Ind[x, yind]
   E_hist <- C_hist/I_hist
   if(any(is.na(E_hist))) stop("Missing values in catch and index in Data object.")
[…]

### 2019-10-24 R/assess_SCA.R

@@ -382,7 +382,7 @@ SCA <- function(x = 1, Data, SR = c("BH", "Ricker"), vulnerability = c("logistic
 
   if(rescale != 1) {
     vars_div <- c("B", "E", "CAApred", "CN", "Cpred", "N", "VB", "R", "R_early", "VB0", "R0", "B0", "E0", "N0")
-    vars_mult <- "Brec"
+    vars_mult <- c("Brec", "q")
     var_trans <- c("R0", "q")
     fun_trans <- c("/", "*")
     fun_fixed <- c("log", NA)

### 2019-10-24 R/assess_SCA2.R

@@ -174,7 +174,7 @@ SCA2 <- function(x = 1, Data, SR = c("BH", "Ricker"), vulnerability = c("logisti
 
   if(rescale != 1) {
     vars_div <- c("meanR", "B", "E", "CAApred", "Cpred", "CN", "N", "VB", "R", "R_early")
-    vars_mult <- NULL
+    vars_mult <- "q"
     var_trans <- c("meanR", "q")
     fun_trans <- c("/", "*")
     fun_fixed <- c("log", NA)

### 2019-10-24 R/assess_SCA_Pope.R

@@ -225,7 +225,7 @@ SCA_Pope <- function(x = 1, Data, SR = c("BH", "Ricker"), vulnerability = c("log
 
   if(rescale != 1) {
     vars_div <- c("B", "E", "CAApred", "CN", "N", "VB", "R", "R_early", "VB0", "R0", "B0", "E0", "N0")
-    vars_mult <- "Brec"
+    vars_mult <- c("Brec", "q")
     var_trans <- c("R0", "q")
     fun_trans <- c("/", "*")
     fun_fixed <- c("log", NA)

### 2019-10-24 R/assess_SP.R

@@ -205,7 +205,7 @@ SP <- function(x = 1, Data, rescale = "mean1", start = NULL, fix_dep = TRUE, fix
 
   if(rescale != 1) {
     vars_div <- c("B", "BMSY", "K", "MSY", "Cpred", "SP")
-    vars_mult <- NULL
+    vars_mult <- "q"
     var_trans <- c("MSY", "K", "q")
     fun_trans <- c("/", "/", "*")
     fun_fixed <- c("log", NA, NA)
@@ -347,7 +347,7 @@ SP_SS <- function(x = 1, Data, rescale = "mean1", start = NULL, fix_dep = TRUE, 
 
   if(rescale != 1) {
     vars_div <- c("B", "BMSY", "K", "MSY", "Cpred", "SP")
-    vars_mult <- NULL
+    vars_mult <- "q"
     var_trans <- c("MSY", "K", "q")
     fun_trans <- c("/", "/", "*")
     fun_fixed <- c("log", NA, NA)

### 2019-10-24 R/assess_cDD.R

@@ -165,9 +165,9 @@ cDD <- function(x = 1, Data, SR = c("BH", "Ricker"), rescale = "mean1", start = 
 
   if(rescale != 1) {
     vars_div <- c("B0", "B", "Cpred", "N0", "N", "R0", "R", "Binf", "Rinf", "Ninf")
-    vars_mult <- c("Brec")
-    var_trans <- c("R0")
-    fun_trans <- c("/")
+    vars_mult <- c("Brec", "q")
+    var_trans <- c("R0", "q")
+    fun_trans <- c("/", "*")
     fun_fixed <- c("log")
     rescale_report(vars_div, vars_mult, var_trans, fun_trans, fun_fixed)
   }
@@ -323,9 +323,9 @@ cDD_SS <- function(x = 1, Data, SR = c("BH", "Ricker"), rescale = "mean1", start
 
   if(rescale != 1) {
     vars_div <- c("B0", "B", "Cpred", "N0", "N", "R0", "R", "Binf", "Rinf", "Ninf")
-    vars_mult <- c("Brec")
-    var_trans <- c("R0")
-    fun_trans <- c("/")
+    vars_mult <- c("Brec", "q")
+    var_trans <- c("R0", "q")
+    fun_trans <- c("/", "*")
     fun_fixed <- c("log")
     rescale_report(vars_div, vars_mult, var_trans, fun_trans, fun_fixed)
   }

### 2019-10-24 R/define_S4_class_SRA.R

@@ -7,7 +7,7 @@
 #' @docType class
 #'
 #' @slot OM An updated operating model, class \linkS4class{OM}.
-#' @slot SSB An matrix of estimated spawning biomass with \code{OM@@nsim} rows and \code{OM@@nyears+1} columns.
+#' @slot SSB A matrix of estimated spawning biomass with \code{OM@@nsim} rows and \code{OM@@nyears+1} columns.
 #' @slot NAA An array for the predicted numbers at age with dimension \code{OM@@nsim}, \code{OM@@nyears+1}, and \code{OM@@maxage}.
 #' @slot CAA An array for the predicted catch at age with dimension \code{OM@@nsim}, \code{OM@@nyears}, \code{OM@@maxage}, and nfleet.
 #' @slot CAL An array for the predicted catch at length with dimension \code{OM@@nsim}, \code{OM@@nyears}, length bins, and nfleet.

### 2019-10-24 R/define_S4_class_proj.R

@@ -1,0 +1,424 @@
+
+#' Class-\code{proj}
+#'
+#' An S4 class for the output from \link{projection}.
+#'
+#' @name proj-class
+#' @docType class
+#'
+#' @slot Model Name of the assessment model.
+#' @slot Name Name of Data object.
+#' @slot FMort A matrix of fishing mortality over \code{p_sim} rows and \code{p_years} columns.
+#' @slot B An matrix of biomass with \code{p_sim} rows and \code{p_years} columns.
+#' @slot SSB An matrix of spawning biomass with \code{p_sim} rows and \code{p_years} columns.
+#' @slot R A matrix of recruitment over \code{p_sim} rows and \code{p_years} columns.
+#' @slot N A matrix of abundance over \code{p_sim} rows and \code{p_years} columns.
+#' @slot Catch A matrix of observed catch over \code{p_sim} rows and \code{p_years} columns.
+#' @slot Index A matrix of observed index over \code{p_sim} rows and \code{p_years} columns.
+#' @slot C_at_age An array for catch-at-age with dimension \code{c(p_sim, p_years, maxage)}.
+#' @seealso \link{projection}
+#' @author Q. Huynh
+#' @export proj
+#' @exportClass proj
+proj <- setClass("proj", slots = c(Model = "character", Name = "character", FMort = "matrix",
+                                   B = "matrix", SSB = "matrix", R = "matrix", N = "matrix",
+                                   Catch = "matrix", Index = "matrix", C_at_age = "array"))
+
+
+#' Projections for assessment models
+#'
+#' This function takes an assessment model and runs a stochastic projection based on future F or catch.
+#'
+#' @param Assessment An object of class \linkS4class{Assessment}.
+#' @param constrain Whether to project on future F or catch. By default, projects on F.
+#' @param FMort The projection F, either of length 1 for constant F for the entirety of the projection or length p_years.
+#' @param Catch The projection catch, either of length 1 for constant catch for the entirety of the projection or length p_years.
[…]

### 2019-10-24 R/report_DD_TMB.R

@@ -13,10 +13,10 @@ summary_DD_TMB <- function(Assessment, state_space = FALSE) {
                   "Age of knife-edge selectivity",
                   "Weight at age k")
   rownam <- c("S0", "alpha", "rho", "k", "w_k")
-  if("log_sigma" %in% names(obj$env$map)) {
-    Value <- c(Value, TMB_report$sigma)
+  if("log_omega" %in% names(obj$env$map)) {
+    Value <- c(Value, TMB_report$omega)
     Description <- c(Description, "Catch SD (log-space)")
-    rownam <- c(rownam, "sigma")
+    rownam <- c(rownam, "omega")
   }
   if("log_tau" %in% names(obj$env$map)) {
     Value <- c(Value, TMB_report$tau)

### 2019-10-24 R/report_VPA.R

@@ -35,6 +35,7 @@ summary_VPA <- function(Assessment) {
                  derived_quantities = derived, model_estimates = model_estimates)
   return(output)
 }
+class(VPA) <- "Assessment"
 
 rmd_VPA <- function(Assessment) {
   ss <- rmd_summary("Virtual Population Analysis (VPA)")

### 2019-10-24 R/z_MPs.R

@@ -140,7 +140,7 @@ get_dependencies <- function(Assess, arg = list()) {
   more_deps <- dep_args[do.call(c, formals2[dep_match])]
 
   if(length(more_deps) > 0) {
-    if(Assess == "DD_SS" && any(names(more_deps) == "fix_sigma")) more_deps$fix_sigma <- "Data@CV_Cat"
+    #if(Assess == "DD_SS" && any(names(more_deps) == "fix_sigma")) more_deps$fix_sigma <- "Data@CV_Cat"
     more_deps <- paste(do.call(c, more_deps), collapse = ", ")
     dep <- paste(c(dep, more_deps), collapse = ", ")
   }

### 2019-10-24 inst/include/DD.hpp

@@ -99,13 +99,13 @@ Type DD(objective_function<Type> *obj) {
   }
 
   //--ARGUMENTS FOR NLL
-  Type sigma = calc_sigma(C_hist, Cpred);
+  Type omega = calc_sigma(C_hist, Cpred);
 
   // Objective function
   //creates storage for jnll and sets value to 0
   Type nll = 0.;
   for(int tt=0; tt<ny; tt++) {
-    if(C_hist(tt) > 0) nll -= dnorm(log(C_hist(tt)), log(Cpred(tt)), sigma, true);
+    if(C_hist(tt) > 0) nll -= dnorm(log(C_hist(tt)), log(Cpred(tt)), omega, true);
   }
 
   //Summing individual jnll and penalties
@@ -115,13 +115,14 @@ Type DD(objective_function<Type> *obj) {
   ADREPORT(R0);
   ADREPORT(h);
   ADREPORT(q);
-  ADREPORT(sigma);
-  REPORT(sigma);
+  ADREPORT(omega);
+  REPORT(omega);
   REPORT(nll);
   REPORT(Arec);
   REPORT(Brec);
   REPORT(Spr0);
   REPORT(Cpred);
+  REPORT(q);
   REPORT(B);
   REPORT(N);
   REPORT(R);

### 2019-10-24 inst/include/DD_SS.hpp

@@ -23,7 +23,7 @@ Type DD_SS(objective_function<Type> *obj) {
   PARAMETER(transformed_h);
   PARAMETER(log_q);
   PARAMETER(U_equilibrium);
-  PARAMETER(log_sigma);
+  PARAMETER(log_omega);
   PARAMETER(log_tau);
   PARAMETER_VECTOR(log_rec_dev);
 
@@ -34,7 +34,7 @@ Type DD_SS(objective_function<Type> *obj) {
   h += 0.2;
   Type R0 = exp(log_R0);
   Type q = exp(log_q);
-  Type sigma = exp(log_sigma);
+  Type omega = exp(log_omega);
   Type tau = exp(log_tau);
 
   //--DECLARING DERIVED VALUES
@@ -117,7 +117,7 @@ Type DD_SS(objective_function<Type> *obj) {
   nll_comp.setZero();
 
   for(int tt=0; tt<ny; tt++){
-    if(C_hist(tt) > 0) nll_comp(0) -= keep(tt) * dnorm(log(C_hist(tt)), log(Cpred(tt)), sigma, true);
+    if(C_hist(tt) > 0) nll_comp(0) -= keep(tt) * dnorm(log(C_hist(tt)), log(Cpred(tt)), omega, true);
     if(tt + k < ny) nll_comp(1) -= dnorm(log_rec_dev(tt), Type(0), tau, true);
   }
 
@@ -128,9 +128,9 @@ Type DD_SS(objective_function<Type> *obj) {
   ADREPORT(R0);
   ADREPORT(h);
   ADREPORT(q);
-  ADREPORT(sigma);
+  ADREPORT(omega);
   ADREPORT(tau);
-  REPORT(sigma);
+  REPORT(omega);
   REPORT(tau);
   REPORT(nll_comp);
   REPORT(nll);
@@ -138,6 +138,7 @@ Type DD_SS(objective_function<Type> *obj) {
   REPORT(Brec);
   REPORT(Spr0);
   REPORT(Cpred);
+  REPORT(q);
   REPORT(B);
   REPORT(N);
   REPORT(R);

### 2019-10-24 inst/include/SCA.hpp

@@ -73,20 +73,17 @@ Type SCA(objective_function<Type> *obj) {
   Type E0 = R0 * EPR0;
   Type VB0 = R0 * sum_VBPR(NPR_virgin, weight, vul);
 
-  Type Arec, Brec;
-
+  Type CR, Brec;
   if(SR_type == "BH") {
-    Arec = 4 * h;
-    Arec /= 1-h;
+    CR = 4 * h;
+    CR /= 1-h;
     Brec = 5*h - 1;
     Brec /= (1-h);
   } else {
-    Arec = pow(5*h, 1.25);
-    Brec = 1.25;
-    Brec *= log(5*h);
+    CR = pow(5*h, 1.25);
+    Brec = 1.25 * log(5*h);
   }
-  Type CR = Arec;
-  Arec /= EPR0;
+  Type Arec = CR/EPR0;
   Brec /= E0;
 
   ////// During time series year = 1, 2, ..., n_y
@@ -230,6 +227,7 @@ Type SCA(objective_function<Type> *obj) {
   REPORT(vul);
 
   REPORT(F);
+  REPORT(q);
 
   REPORT(N);
   REPORT(CN);

### 2019-10-24 inst/include/SCA2.hpp

@@ -172,6 +172,7 @@ Type SCA2(objective_function<Type> *obj) {
   REPORT(vul);
 
   REPORT(F);
+  REPORT(q);
 
   REPORT(N);
   REPORT(CN);

### 2019-10-24 inst/include/SCA_Pope.hpp

@@ -68,20 +68,17 @@ Type SCA_Pope(objective_function<Type> *obj) {
   Type E0 = R0 * EPR0;
   Type VB0 = R0 * sum_VBPR(NPR_virgin, weight, vul);
 
-  Type Arec, Brec;
-
+  Type CR, Brec;
   if(SR_type == "BH") {
-    Arec = 4 * h;
-    Arec /= 1-h;
+    CR = 4 * h;
+    CR /= 1-h;
     Brec = 5*h - 1;
     Brec /= (1-h);
   } else {
-    Arec = pow(5*h, 1.25);
-    Brec = 1.25;
-    Brec *= log(5*h);
+    CR = pow(5*h, 1.25);
+    Brec = 1.25 * log(5*h);
   }
-  Type CR = Arec;
-  Arec /= EPR0;
+  Type Arec = CR/EPR0;
   Brec /= E0;
 
   ////// During time series year = 1, 2, ..., n_y
@@ -219,6 +216,7 @@ Type SCA_Pope(objective_function<Type> *obj) {
 
   REPORT(vul_par);
   REPORT(vul);
+  REPORT(q);
 
   REPORT(N);
   REPORT(CN);

### 2019-10-24 inst/include/SP.hpp

@@ -85,6 +85,8 @@ Type SP(objective_function<Type> *obj) {
   REPORT(MSY);
   REPORT(dep);
   REPORT(n);
+  REPORT(n_term);
+  REPORT(q);
   REPORT(sigma);
   REPORT(r);
   REPORT(K);

### 2019-10-24 inst/include/SP_SS.hpp

@@ -94,6 +94,8 @@ Type SP_SS(objective_function<Type> *obj) {
   REPORT(MSY);
   REPORT(dep);
   REPORT(n);
+  REPORT(n_term);
+  REPORT(q);
   REPORT(sigma);
   REPORT(tau);
   REPORT(r);

### 2019-10-24 inst/include/cDD.hpp

@@ -123,6 +123,7 @@ Type cDD(objective_function<Type> *obj) {
   REPORT(BPR0);
   REPORT(Cpred);
   REPORT(Ipred);
+  REPORT(q);
   REPORT(B);
   REPORT(N);
   REPORT(R);

### 2019-10-24 inst/include/cDD_SS.hpp

@@ -129,12 +129,14 @@ Type cDD_SS(objective_function<Type> *obj) {
   ADREPORT(sigma);
   ADREPORT(tau);
   REPORT(sigma);
+  REPORT(tau);
   REPORT(nll);
   REPORT(Arec);
   REPORT(Brec);
   REPORT(BPR0);
   REPORT(Cpred);
   REPORT(Ipred);
+  REPORT(q);
   REPORT(B);
   REPORT(N);
   REPORT(R);

### 2020-01-23 R/SRA_scope.R
Set-up age-only model

@@ -396,7 +396,7 @@ SRA_scope <- function(OM, data = list(), condition = c("catch", "catch2", "effor
   Perr <- do.call(rbind, lapply(mod, make_Perr))
 
   make_early_Perr <- function(x) {
-    res <- x$report$R_eq * x$report$NPR_equilibrium / (x$report$R0 * x$report$NPR_unfished[[1]])
+    res <- x$report$R_eq * x$report$NPR_equilibrium / x$report$R0 / x$report$NPR_unfished[1, ]
     bias_corr <- ifelse(x$obj$env$data$est_early_rec_dev, exp(-0.5 * x$report$tau^2), 1)
     early_dev <- exp(x$report$log_early_rec_dev) * bias_corr
     out <- res[-1] * early_dev
@@ -619,6 +619,7 @@ SRA_scope_est <- function(x = 1, data, I_type, selectivity, s_selectivity, SR_ty
 
   if(is.null(dots$nit_F)) nit_F <- 3L else nit_F <- dots$nit_F
   if(is.null(dots$plusgroup)) plusgroup <- 1L else plusgroup <- as.integer(dots$plusgroup)
+  age_only_model <- StockPars$Len_age[x, , 1:(nyears+1)] %>% apply(2, function(x) all(x == 1:max_age)) %>% all()
   TMB_data_all <- list(condition = data$condition,
                        nll_C = as.integer((data$condition == "effort" & nfleet > 1) || data$condition == "catch"),
                        I_hist = data$Index, sigma_I = data$I_sd, CAA_hist = data$CAA, CAA_n = pmin(CAA_n, ESS[1]),
@@ -627,7 +628,8 @@ SRA_scope_est <- function(x = 1, data, I_type, selectivity, s_selectivity, SR_ty
                        sel_block = rbind(data$sel_block, data$sel_block[nyears, ]), nsel_block = data$nsel_block,
                        n_y = nyears, max_age = max_age, nfleet = nfleet, nsurvey = nsurvey,
                        M = t(StockPars$M_ageArray[x, , 1:nyears]), len_age = t(StockPars$Len_age[x, , 1:(nyears+1)]),
-                       Linf = StockPars$Linf[x], CV_LAA = StockPars$LenCV[x], wt = t(StockPars$Wt_age[x, , 1:(nyears+1)]),
+                       Linf = ifelse(age_only_model, max_age, StockPars$Linf[x]),
[…]

### 2020-01-23 R/define_S4_class_SRA.R

@@ -326,8 +326,10 @@ setMethod("plot", signature(x = "SRA", y = "missing"),
               CAA_bubble <- rmd_bubble("Year", "CAA_all", fig.cap = "Predicted catch-at-age (summed over all fleets).", bubble_adj = as.character(bubble_adj))
 
               CAL_all <- apply(report$CALpred, c(1, 2), sum)
-              CAL_bubble <- rmd_bubble("Year", "CAL_all", CAL_bins = "data_mean_fit$length_bin",
-                                       fig.cap = "Predicted catch-at-length (summed over all fleets).", bubble_adj = as.character(bubble_adj))
+              if(!all(is.na(CAL_all))) {
+                CAL_bubble <- rmd_bubble("Year", "CAL_all", CAL_bins = "data_mean_fit$length_bin",
+                                         fig.cap = "Predicted catch-at-length (summed over all fleets).", bubble_adj = as.character(bubble_adj))
+              } else CAL_bubble <- ""
 
               ts_output <- c(sel_matplot, F_matplot, rmd_SSB(), SSB_plot, rmd_SSB_SSB0(FALSE), rmd_R(), rmd_SRA_SR(),
                              rmd_residual("log_rec_dev", fig.cap = "Time series of recruitment deviations.", label = "log-Recruitment deviations"),
@@ -673,10 +675,12 @@ rmd_SRA_fleet_output <- function(ff, f_name) {
            paste0("if(any(data$CAL[, , ", ff, "] > 0, na.rm = TRUE)) {"),
            paste0("  MLobs <- (data$CAL[, , ", ff, "] %*% length_bin)/rowSums(data$CAL[, , ", ff, "], na.rm = TRUE)"),
            paste0("} else if(any(data$ML[, ", ff, "] > 0, na.rm = TRUE)) MLobs <- data$ML[, ", ff, "] else MLobs <- NA"),
-           "ylim <- c(0.9, 1.1) * range(c(MLpred, MLobs), na.rm = TRUE)",
-           "matplot(Year, MLpred, type = \"l\", col = scenario$col, lty = scenario$lty, lwd = scenario$lwd, xlab = \"Year\", ylab = \"Mean length\", ylim = ylim)",
-           "if(!all(is.na(MLobs))) lines(Year, MLobs, col = \"black\", typ = \"o\")",
[…]

### 2020-01-23 inst/include/SP.hpp

@@ -48,7 +48,7 @@ Type SP(objective_function<Type> *obj) {
 
   Type BMSY = MSY/FMSY;
   Type K = BMSY / n_term2;
-  Type r = MSY * n_term / K; // r = FMSY * pow(n, 1/(1-n))
+  Type r = MSY * n_term / K; // r = FMSY * n
 
   vector<Type> B(ny+1);
   vector<Type> SP(ny);

### 2020-01-23 inst/include/SRA_scope.hpp

@@ -242,7 +242,7 @@ Type SRA_scope(objective_function<Type> *obj) {
       if(est_rec_dev(y+1)) R(y+1) *= exp(log_rec_dev(y+1) - 0.5 * tau * tau);
     }
     N(y+1,0) = R(y+1);
-    ALK(y) = generate_ALK(length_bin, len_age, CV_LAA, max_age, nlbin, bin_width, y);
+    if(Type(max_age) != Linf) ALK(y) = generate_ALK(length_bin, len_age, CV_LAA, max_age, nlbin, bin_width, y);
 
     if(condition == "catch") {
       for(int ff=0;ff<nfleet;ff++) {
@@ -272,17 +272,20 @@ Type SRA_scope(objective_function<Type> *obj) {
         CN(y,ff) += CAApred(y,a,ff);
         Cpred(y,ff) += CAApred(y,a,ff) * wt(y,a);
 
-        for(int len=0;len<nlbin;len++) {
-          CALpred(y,len,ff) += CAApred(y,a,ff) * ALK(y)(a,len);
-          mlen_pred(y,ff) += CAApred(y,a,ff) * ALK(y)(a,len) * length_bin(len);
+        if(Type(max_age) != Linf) {
+          for(int len=0;len<nlbin;len++) {
+            CALpred(y,len,ff) += CAApred(y,a,ff) * ALK(y)(a,len);
+            mlen_pred(y,ff) += CAApred(y,a,ff) * ALK(y)(a,len) * length_bin(len);
+          }
         }
+
         VB(y+1,ff) += vul(y+1,a,ff) * N(y+1,a) * wt(y+1,a);
       }
 
       B(y+1) += N(y+1,a) * wt(y+1,a);
       E(y+1) += N(y+1,a) * wt(y+1,a) * mat(y+1,a);
     }
-    for(int ff=0;ff<nfleet;ff++) mlen_pred(y,ff) /= CN(y,ff);
+    if(Type(max_age) != Linf) for(int ff=0;ff<nfleet;ff++) mlen_pred(y,ff) /= CN(y,ff);
   }
 
   // Calculate nuisance parameters and likelihood
@@ -310,7 +313,7 @@ Type SRA_scope(objective_function<Type> *obj) {
         s_CN(y,sur) += s_CAApred(y,a,sur);
 
         if(I_basis(sur)) B_sur(y,sur) += s_CAApred(y,a,sur) * wt(y,a);
[…]

### 2020-02-06 R/MSEtool-package.R
Update SRA vignette

@@ -23,6 +23,10 @@
 #' \item \href{../doc/VPA.html}{Virtual population analysis} (\link{VPA})
 #' }
 #'
+#'
+#' The \code{\link{SRA_scope}} model can be used to condition DLMtool operating models from real data. Information can be found
+#' \href{../doc/SRA_scope.html}{here}.
+#'
 #' MSEtool also contains \link{multiMSE}, a platform for multi-stock and multi-fleet operating models based on components
 #' from DLMtool. An overview of multiMSE is available in the \href{../doc/multiMSE.html}{multiMSE vignette}.
 #'

### 2020-02-06 R/report_SCA.R

@@ -193,7 +193,7 @@ retrospective_SCA <- function(Assessment, nyr, SCA2 = FALSE) {
     params_ret$log_rec_dev <- rep(0, n_y_ret)
     params_ret$logF <- info$params$logF[1:n_y_ret]
 
-    if(data$yindF + 1 > n_y_ret) {
+    if(info$data$yindF + 1 > n_y_ret) {
       data_ret$yindF <- as.integer(0.5 * data_ret$n_y)
       params_ret$logF[data_ret$yindF + 1] <- info$params$logF[info$data$yindF + 1]
     }

### 2020-02-08 R/define_S4_class_SRA.R
Set dpi for word markdown

@@ -149,7 +149,7 @@ setMethod("plot", signature(x = "SRA", y = "missing"),
                         "",
                         "```{r setup, include = FALSE, echo = FALSE}",
                         "  knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE,",
-                        "  fig.width = 6, fig.height = 4.5, out.width = \"650px\", comment = \"#>\")",
+                        "  fig.width = 6, fig.height = 4.5, dpi = 600, out.width = \"650px\", comment = \"#>\")",
                         "```\n")
 
             ####### Updated historical OM parameters
@@ -984,7 +984,7 @@ compare_SRA <- function(..., compare = TRUE, filename = "compare_SRA", dir = tem
               "",
               "```{r setup, include = FALSE, echo = FALSE}",
               "  knitr::opts_chunk$set(collapse = TRUE, echo = FALSE, message = FALSE,",
-              "  fig.width = 6, fig.height = 4.5, out.width = \"650px\", comment = \"#>\")",
+              "  fig.width = 6, fig.height = 4.5, dpi = 600, out.width = \"650px\", comment = \"#>\")",
               "```\n")
 
 

### 2020-04-08 R/SRA_scope.R
Update SRA vignette/help

@@ -16,9 +16,9 @@
 #' @param condition String to indicate whether the SRA model is conditioned on "catch" (where F is estimated), "catch2" (where F is solved internally using Newton's method),
 #' or "effort".
 #' @param selectivity A character vector of length nfleet to indicate \code{"logistic"} or \code{"dome"} selectivity for each fleet in \code{Chist}.
+#' If there is time-varying selectivity, this is a characte vector of length nsel_block (see Data section below).
 #' @param s_selectivity A vector of length nsurvey to indicate \code{"logistic"} or \code{"dome"} selectivity for each survey in \code{Index}. Use a number
-#' for an age-specific index.
-#' Only used if any of the corresponding entries of \code{data$I_type = "est"} or if a number is specified here.
+#' for an age-specific index. Only used if any of the corresponding entries of \code{data$I_type = "est"} or if a number is specified here.
 #' @param LWT A named list of likelihood weights for the SRA model. See details.
 #' @param comp_like A string indicating either "multinomial" (default) or "lognormal" distributions for the composition data.
 #' @param ESS If \code{comp_like = "multinomial"}, a numeric vector of length two to cap the maximum effective samples size of the age and length compositions,
@@ -34,6 +34,10 @@
 #' @param control A named list of arguments (e.g, max. iterations, etc.) for optimization, to be passed to \code{\link[stats]{nlminb}}.
 #' @param ... Other arguments to pass in for starting values of parameters and fixing parameters. See details.
 #' @return An object of class \linkS4class{SRA}, including the updated operating model object.
+#'
+#' @section Vignette:
+#' Two vignettes for the SRA model provide a general \href{../doc/SRA_scope.html}{overview} and more details on
+#' \href{../doc/SRA_scope2.html}{selectivity} settings.
 #'
 #' @section Data:
 #' One of indices, age compositions, or length compositions should be provided in addition to the historical catch or effort. Not all arguments
@@ -71,6 +75,9 @@
 #' \item I_units - Optional, an integer vector to indicate whether indices are biomass based (1) or abundance-based (0). By default, all are biomass-based.
[…]

### 2020-04-08 R/define_S4_class_SRA.R

@@ -12,10 +12,51 @@
 #' @slot CAA An array for the predicted catch at age with dimension \code{OM@@nsim}, \code{OM@@nyears}, \code{OM@@maxage}, and nfleet.
 #' @slot CAL An array for the predicted catch at length with dimension \code{OM@@nsim}, \code{OM@@nyears}, length bins, and nfleet.
 #' @slot conv A logical vector of length \code{OM@@nsim} indicating convergence of the SRA scoping model in the i-th simulation.
-#' @slot Misc A list of length \code{OM@@nsim} with more output from the fitted SRA scoping model.
-#' @slot mean_fit A list of output from fit to mean values of life history parameters in the operating model.
+#' @slot Misc A list of length \code{OM@@nsim} with more output from the fitted SRA scoping model. Within each simulation, items of interest include:
+#'
+#' \itemize{
+#' \item B - total biomass - vector of length nyears+1
+#' \item E0 - annual unfished spawning biomass - vector of length nyears
+#' \item E0_SR - unfished spawning biomass for the stock-recruit relationship - numeric
+#' \item Arec - alpha parameter of the stock-recruit relationship - numeric
+#' \item Brec - beta parameter of the stock-recruit relationship - numeric
+#' \item R - recruitment - vector of length nyears+1
+#' \item R_early - recruitment for the cohorts in first year of the model - vector maxage-1
+#' \item VB - vulnerable biomass - matrix of nyears x nfleet
+#' \item N - abundance at age - matrix of nyears+1 x maxage
+#' \item F - apical fishing mortality - matrix of nyears x nfleet
+#' \item F_at_age - fishing mortality at age - array of nyears x maxage x nfleet
+#' \item F_equilibrium - equilibrium fishing mortality prior to first year - vector of length nfleet
+#' \item M - natural mortality - matrix of nyears x maxage
+#' \item Z - total mortality - matrix of nyears x maxage
+#' \item q - survey catchability - vector of length nsurvey
+#' \item s_vul - survey selectivity at age - array of dim nyears+1, maxage, nsurvey
+#' \item s_vul_len - corresponding survey selectivity at length - matrix of nbins x nsurvey
+#' \item Ipred - predicted index values - matrix of nyears x nsurvey
+#' \item s_CAApred - predicted survey catch at age - array of dim nyears, maxage, nsurvey
+#' \item vul - fleet selectivity at age - array of dim nyears+1, maxage, nfleet (or nsel_block)
[…]

### 2020-04-10 R/SRA_scope.R
Update vignette

@@ -35,8 +35,13 @@
 #' @param ... Other arguments to pass in for starting values of parameters and fixing parameters. See details.
 #'
 #' @section Vignette:
-#' Two vignettes for the SRA model provide a general \href{../doc/SRA_scope.html}{overview} and more details on
-#' \href{../doc/SRA_scope2.html}{selectivity} settings.
+#' Three vignettes are available for the SRA model:
+#'
+#' \itemize{
+#' \item \href{../doc/SRA_scope.html}{General overview of approach}
+#' \item \href{../doc/SRA_scope_eq.html}{Mathematical description}
+#' \item \href{../doc/SRA_scope_sel.html}{Setup of selectivity settings} (useful for more data-rich cases)
+#' }
 #'
 #' @section Data:
 #' One of indices, age compositions, or length compositions should be provided in addition to the historical catch or effort. Not all arguments
@@ -75,7 +80,7 @@
 #' \item age_error - Optional, a square matrix of maxage rows and columns to specify ageing error. The aa-th column assigns a proportion of the true age in the
 #' a-th row to observed age. Thus, all \code{rowSums(age_error)} should be 1. Default is an identity matrix (no ageing error).
 #' \item sel_block - Optional, for time-varying fleet selectivity (in time blocks), a matrix of nyears x nfleet that assigns a selectivity function to a fleet-year combination.
-#' See \href{../doc/SRA_scope2.html}{selectivity} vignette.
+#' See \href{../doc/SRA_scope_sel.html}{selectivity} vignette.
 #' \item nsel_block - Optional, the number of selectivity blocks.
 #' }
 #'
@@ -535,31 +540,7 @@ SRA_scope <- function(OM, data = list(), condition = c("catch", "catch2", "effor
   return(output)
 }
 
-#### Parameters
-# R0
-# sel
-# rec_devs
 
-#### Data list
-# Cat vector by year
-# Index matrix by year and fleet
-# Index_type vector by fleet
-# CAA matrix by year, age
-# CAL matrix by year, length bin
-# mean length vector by year
-
-#### Life history
-# M array by nsim, age, year
-# mat_age array by nsim, age
-# Length at age array by nsim, age
-# Weight at age arraty by nsim, length
-# CV_LAA vector by nsim
-# h steepness
-
[…]

### 2020-05-06 R/MSEtool-package.R
Update vignette titles

@@ -11,7 +11,7 @@
 #' @author Tom Carruthers \email{t.carruthers@@oceans.ubc.ca}
 #' @author Adrian Hordyk \email{a.hordyk@@oceans.ubc.ca}
 #' @section How to use MSEtool:
-#' The main features of MSEtool are the assessment models and the ability to make management procedures by combining
+#' The main features of MSEtool are the assessment models and the ability to make model-based management procedures by combining
 #' assessment models with harvest control rules. Such MPs can be used and tested in management strategy evaluation
 #' with DLMtool operating models. An overview of these features is available in the \href{../doc/MSEtool.html}{MSEtool vignette}.
 #'
@@ -20,6 +20,7 @@
 #' \item \href{../doc/Surplus_production.html}{Surplus production} (\link{SP}, \link{SP_SS}, \link{SP_Fox}, and \code{spict})
 #' \item \href{../doc/Delay_difference.html}{Delay difference} (\link{DD}, \link{cDD}, \link{DD_SS}, and \link{cDD_SS})
 #' \item \href{../doc/SCA.html}{Statistical catch-at-age} (\link{SCA}, \link{SCA2}, and \link{SCA_Pope})
+#' \item Simple Stock Synthesis (\link{SSS} which implements \link{SCA_Pope} with fixed depletion assumption)
 #' \item \href{../doc/VPA.html}{Virtual population analysis} (\link{VPA})
 #' }
 #'

### 2020-06-16 R/SRA_scope.R
Update vignette

@@ -292,7 +292,7 @@ SRA_scope <- function(OM, data = list(), condition = c("catch", "catch2", "effor
     }
 
     report_internal_fn <- function(x, samps, obj, conv) {
-      report <- obj$report(samps[x, ]) %>% SRA_posthoc_adjust(obj)
+      report <- obj$report(samps[x, ]) %>% SRA_posthoc_adjust(obj, par = samps[x, ])
       report$conv <- conv
       return(report)
     }
@@ -978,13 +978,12 @@ all_identical_sims_fn <- function(StockPars, FleetPars, ObsPars, data, dots) {
 }
 
 
-SRA_dynamic_SSB0 <- function(obj) {
+SRA_dynamic_SSB0 <- function(obj, par = obj$env$last.par.best) {
 
   if(obj$env$data$condition == "catch") {
 
-    par_F0 <- obj$env$last.par.best
-    par_F0[names(par_F0) == "log_F" | names(par_F0) == "log_F_equilibrium"] <- log(1e-8)
-    out <- obj$report(par_F0)$E
+    par[names(par) == "log_F" | names(par) == "log_F_equilibrium"] <- log(1e-8)
+    out <- obj$report(par)$E
 
   } else if(obj$env$data$condition == "catch2") {
 
@@ -993,20 +992,19 @@ SRA_dynamic_SSB0 <- function(obj) {
 
     obj2 <- MakeADFun(data = new_args$data, parameters = new_args$params, map = new_args$map, random = obj$env$random,
                       DLL = "MSEtool", silent = TRUE)
-    out <- obj2$report(obj$env$last.par.best)$E
+    out <- obj2$report(par)$E
 
   } else {
 
-    par_F0 <- obj$env$last.par.best
-    par_F0[names(par_F0) == "log_q_effort"] <- log(1e-8)
-    out <- obj$report(par_F0)$E
+    par[names(par) == "log_q_effort"] <- log(1e-8)
+    out <- obj$report(par)$E
 
   }
 
   return(out)
 }
 
-SRA_posthoc_adjust <- function(report, obj) {
+SRA_posthoc_adjust <- function(report, obj, par = obj$env$last.par.best) {
   data <- obj$env$data
   report$F_at_age <- report$Z - data$M
   report$NPR_unfished <- do.call(rbind, report$NPR_unfished)
[…]
